---
title: "<br>Altcoin trading based on MA and EMA crosses"
date: "`r Sys.Date()`"
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")
library("kableExtra")

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

URL="https://web-api.coinmarketcap.com/v1/cryptocurrency/listings/latest?limit=200"
download.file(URL,destfile="cmcdat.txt")
cmcdat <- fromJSON("cmcdat.txt")
cmcdf <- data.frame(cmcdat$data[,1:4],cmcdat$data$quote$USD)
cmcdf <- cmcdf[order(-cmcdf$percent_change_7d),]
cmcdf <- subset(cmcdf,percent_change_7d > 0)
dim(cmcdf)

```

```{r,functions1}

fetch <- function(myid) {
  date0 <- Sys.Date()-100
  date1 <- Sys.Date()-1
  myid <- as.character(myid)
  URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?id=",myid,"&convert=USD&interval=daily&time_start=",date0,"&time_end=",date1,sep="")
  if ( file.exists("tmpdat.txt") ) { unlink("tmpdat.txt") }
  Sys.sleep(1)
  download.file(URL,destfile="tmpdat.txt")
  dat <- fromJSON("tmpdat.txt")
  unlink("tmpdat.txt")
  price <- dat$data$quotes
  price <- data.frame(price$time_close,price$quote$USD$close)
  colnames(price) <- c("date","Close")
  price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)
  return(price)
}

## MA CROSS
ma_cross <- function(price,name) {
  n1=13
  n2=20
  ma1 <- tail(SMA(Cl(price),n=n1),1)
  ma2 <- tail(SMA(Cl(price),n=n2),1)
  if (ma1 > ma2) {
    ma1 <- SMA(Cl(price),n=n1)
    ma2 <- SMA(Cl(price),n=n2)
    SMA_HEADER=paste(name, n1 , "and", n2 ,"period MA cross")
    plot(price$Close~as.Date(price$date),type="l",log="y",
    xlab="Date",ylab="price (USD)",main=SMA_HEADER)
    grid()
    lines(ma1~ as.Date(price$date) ,col="red")
    lines(ma2~ as.Date(price$date) ,col="blue")
    return(1)
  } else {
    return(0)
  }
}

## MA
ma <- function(price,name) {
  n1=15
  n2=22
  n3=34
  n4=43
  ma2 <- tail(SMA(Cl(price),n=n2),1)
  current_price <- tail(price$Close,1)
  if (current_price > ma2) {
    ma1 <- SMA(Cl(price),n=n1)
    ma2 <- SMA(Cl(price),n=n2)
    ma3 <- SMA(Cl(price),n=n3)
    ma4 <- SMA(Cl(price),n=n4)
    SMA_HEADER=paste(name, "15,22,34,43 period MA cross")
    plot(price$Close~as.Date(price$date),type="l",log="y",
    xlab="Date",ylab="price (USD)",main=SMA_HEADER)
    grid()
    lines(ma1~ as.Date(price$date) ,col="orange")
    lines(ma2~ as.Date(price$date) ,col="red")
    lines(ma3~ as.Date(price$date) ,col="purple")
    lines(ma4~ as.Date(price$date) ,col="blue")
    return(1)
  } else {
    return(0)
  }
}

coinupdate <- function(myids) {
  if (length(myids)==1) {
    myid <- myids
    name=cmcdf[which(cmcdf$id==myid),2]
    dat <- fetch(myids)
    res1 <- ma_cross(dat,name)
    res2 <- ma(dat,name)
    res <- c("id"=myid,"macross"=res1,"ma"=res2)
  } else {
    res <- sapply(myids, function(myid) {
      name=cmcdf[which(cmcdf$id==myid),2]
      dat <- fetch(myid)
      res1 <- ma_cross(dat,name)
      res2 <- ma(dat,name)
      c("id"=myid,"macross"=res1,"ma"=res2)
    })
  }
  return(res)
}

```

```{r,analyze1}

myids <- cmcdf$id

coindata <- coinupdate(myids)

```

Here are the top uptrending coins.

```{r, table1}

coindata <- t(coindata)

df <- merge(coindata,cmcdf,by="id")

subset(df,macross==1 & ma==1) %>% kbl() %>% kable_styling()

```

Here is the full table.

```{r, table2}

df %>% kbl() %>% kable_styling()

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
