---
title: "<br>Trading based on MA and EMA crosses"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}

sma <- function(price,n,fee) {
ma <- SMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}


```

## Exponential moving average function


```{r,ema}

ema <- function(price,n,fee) {
ma <- EMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

```

## Backtesting

```{r,backtest}

fee=0.005

myrange <- 3:200

res <- sapply(myrange,function(n) { sma(price,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_sma <- res

res <- sapply(myrange,function(n) { ema(price,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_ema <- res

```

## Plots

```{r,plots,fig.height=8,fig.width=8}

plot(x=myrange, y=res_sma$returns, pch=19,
  xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_sma$returns, pch=19,
  xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_ema$returns, pch=19,
  xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_ema$returns, pch=19,
  xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(res_sma$mean,res_sma$returns, main="SMA",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$mean,res_sma$returns , cex=1, labels=rownames(res_sma))
abline(h=hodl,col="red")

plot(res_ema$mean,res_ema$returns, main="EMA",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$mean,res_ema$returns , cex=1, labels=rownames(res_ema))
abline(h=hodl,col="red")

plot(res_sma$trades,res_sma$returns, main="SMA",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$trades,res_sma$returns , cex=1, labels=rownames(res_sma))
abline(h=hodl,col="red")

plot(res_ema$trades,res_ema$returns, main="SMA",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$trades,res_ema$returns , cex=1, labels=rownames(res_ema))
abline(h=hodl,col="red")


```

## SMA Ensembl

```{r,ensembl_sma,fig.height=10,fig.width=10}

res_sma$ln <- log(res_sma$mean)
price2 <- price[200:nrow(price),]
mas <- sapply(3:200,function(n) {SMA(price$Close,n) } ) 
mas2 <- mas[200:nrow(mas),]
hodl2 <- price2[nrow(price2),2] / price2[1,2]

price2$sig <- sapply(1:nrow(price2),function(d) {
  p=price2[d,2]
  x <- as.numeric(p>mas2[d,])
  y <- sum(x * res_sma$ln)
  return(y)
} )

par(mfrow=c(2,1))
plot(price2$Close~as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$sig~as.Date(price2$date),type="l",
  xlab="Date",ylab="sig",main="SMA signal")
grid()

```

## GRS

```{r,ens,fig.height=10,fig.width=10}

mamx <- sapply(3:200,function(x) { SMA(price$Close,n=x) } )
colnames(mamx) <- 3:200
mamx2 <- apply(mamx,2,  function(x) { price$Close > x } )
mamx2 <- apply(mamx2,2,  function(x) { as.numeric(x) } )
mamx2 <- (mamx2*2)-1
grs <- t(apply(mamx2,1,function(x) { x * res_sma$ln } ) )
grs_rs <- rowSums(grs)

par(mfrow=c(2,1))
plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(rowSums(grs) ~as.Date(price$date),type="l",
  xlab="Date",ylab="sig",main="SMA signal")
grid()


```

Trading signal 

```{r,sig_ema1}

entry=0
exit=0

sig_sma <- function(price,grs_rs,entry=0,exit=0,fee=0.005) {
price2 <- price
price2$higher <- grs_rs > 0
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

```

# Backtest Signal EMA

```{r,backtest_sig2}

res <- t(sapply(3:200, function(n) { sig_ema(price2,n,fee) }  ))
res <- as.data.frame(res)
colnames(res) <- c("returns","trades")
rownames(res) <- 3:200
res$mean <- res$returns^(1/res$trades)
res_sig_ema <- res
res

```

## Conclusion

Although returns were higher using the tuned SMA signal, the mean ROI was lower, at 105%, meaning many more trades need to be executed to achieve marginally higher returns.
Still, this is a good proof of concept at how to combine multiple indicators into a single measure.
In the subsequent analysis it will be possible to combine DMI, TSI, MAs on daily and weekly timeframes into a single indicator

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
