---
title: "Trading based on MA and EMA multiples"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

n=number of periods

m=multiple

```{r,sma}


sma <- function(price,n,m,fee) {
ma <- m * SMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( nrow(price2) == 0 ) { return(c(0,0)) }
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}


```

## Exponential moving average function


```{r,ema}

ema <- function(price,n,m,fee) {
ma <- m * EMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( nrow(price2) == 0 ) { return(c(0,0)) }
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

```

## Backtesting

```{r,backtest}

fee=0.001

myrange <- 3:700
res <- sapply(myrange,function(n) { sma(price,n,1,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_sma1 <- res

m=1.2
myrange <- 3:700
res <- sapply(myrange,function(n) { sma(price,n,m,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_sma2 <- res
subset(res_sma2,returns>100)

res <- sapply(myrange,function(n) { ema(price,n,1,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_ema <- res
res_ema

m=0.8
res <- sapply(myrange,function(n) { ema(price,n,m,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_ema2 <- res
subset(res_ema2,returns>100)

```

## Conclusion

Not successful

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
