---
title: "Trading based on MA and EMA crosses"
date: "`r Sys.Date()`"
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")
library("reshape2")
library("gplots")
library("parallel")

NUMCORES=detectCores()

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=weekly&time_start=2013-06-01&time_end=",mydate,sep="")

if ( ! file.exists("btcdat.txt") ) {

  download.file(URL,destfile="btcdat.txt")

}

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```


## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}


n1=12
n2=26
sig=9

macd_sma <- function(price,n,fee,sig) {
n1=myparams[n,1]
n2=myparams[n,2]
if (n1 < n2) { 
  macd1 <- MACD(Cl(price),nFast=n1,nSlow=n2,nSig=sig,maType="SMA")
  price2 <- cbind(price,macd1)
  price2$higher <- price2$macd > price2$signal
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(prod(trades$diff))
} else {
  return(1)
}
}

```

## Exponential moving average function


```{r,ema}

macd_ema <- function(price,n,fee,sig) {
n1=myparams[n,1]
n2=myparams[n,2]
if (n1 < n2) {
  macd1 <- MACD(Cl(price),nFast=n1,nSlow=n2,nSig=sig,maType="EMA")
  price2 <- cbind(price,macd1)
  price2$higher <- price2$macd > price2$signal
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(prod(trades$diff))
} else {
  return(1)
}
}

```

## Backtesting

```{r,backtest}

fee=0.005

myparams <- expand.grid(4:100,4:100)

# SMA approach
bigres <- sapply(4:50,function(sig) {
  res_sma <- mclapply(1:nrow(myparams),function(n) {
    macd_sma(price,n,fee,sig) 
  } , mc.cores=NUMCORES)
  mas1 <- as.data.frame(myparams,stringsAsFactors=FALSE)
  mas1$returns <- unlist(res_sma)
  head(mas1[order(-mas1$returns),],1)[,3]
})

bigres <- c(rep(1,3), bigres )
plot(bigres,main="MACD with SMAs")
grid()

sig=27
res_sma <- mclapply(1:nrow(myparams),function(n) {
  macd_sma(price,n,fee,sig)
} , mc.cores=NUMCORES)
mas1 <- as.data.frame(myparams,stringsAsFactors=FALSE)
mas1$returns <- unlist(res_sma)
head(mas1[order(-mas1$returns),])


# EMA approach
bigres <- sapply(4:50,function(sig) {
  res_ema <- mclapply(1:nrow(myparams),function(n) {
    macd_ema(price,n,fee,sig)
  } , mc.cores=NUMCORES)
  mas1 <- as.data.frame(myparams,stringsAsFactors=FALSE)
  mas1$returns <- unlist(res_ema)
  head(mas1[order(-mas1$returns),],1)[,3]
})

bigres <- c(rep(1,3), bigres )
plot(bigres,main="MACD with EMAs")
grid()

sig=24
res_sma <- mclapply(1:nrow(myparams),function(n) {
  macd_sma(price,n,fee,sig)
} , mc.cores=NUMCORES)
mas1 <- as.data.frame(myparams,stringsAsFactors=FALSE)
mas1$returns <- unlist(res_sma)
head(mas1[order(-mas1$returns),])

```

## Conclusion

Here I looked at the using the MACD for entry and exit signals. 
This is not that much better than using the simplest MA and is on a shorter timeframe and requires many more trades.
For this reason I'm not going to explore this any more.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
