---
title: "<br>Trading BTC based on MA and EMA crosses"
date: "`r Sys.Date()`"
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,price$quote$USD$volume,stringsAsFactors=FALSE)
colnames(price) <- c("date","close","Volume")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

price <- subset(price,Volume>0)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}

sma <- function(price,n,fee) {
ma <- SMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

ema <- function(price,n,fee) {
ma <- EMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

dema <- function(price,n,fee) {
ma <- DEMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

wma <- function(price,n,fee) {
ma <- WMA(x=Cl(price),volume=price$Volume,n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

evwma <- function(price,n,fee) {
ma <- EVWMA(price=Cl(price),volume=price$Volume,n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

zlema <- function(price,n,fee) {
ma <- ZLEMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

vwap <- function(price,n,fee) {
ma <- VWAP(price=Cl(price),volume=price$Volume,n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

hma <- function(price,n,fee) {
ma <- HMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

alma <- function(price,n,fee) {
ma <- ALMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}


T3ma <- function(price,n,fee) {
T3 <- function(x, n=10, v=1) DEMA(DEMA(DEMA(x,n,v),n,v),n,v)
ma <- T3(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

```

## Backtesting

```{r,backtest}

fee=0.005
myrange=3:200

res_sma <- t(sapply(myrange,function(n) { sma(price,n,fee) } ))
rownames(res_sma) <- myrange
res_ema <- t(sapply(3:200,function(n) { ema(price,n,fee) } ) )
rownames(res_ema) <- myrange
res_dema <- t(sapply(3:200,function(n) { dema(price,n,fee) } ) )
rownames(res_dema) <- myrange
res_wma <- t(sapply(3:200,function(n) { wma(price,n,fee) } ) )
rownames(res_wma) <- myrange
res_evwma <- t(sapply(3:200,function(n) { evwma(price,n,fee) } ) )
rownames(res_evwma) <- myrange
res_zlema <- t(sapply(3:200,function(n) { zlema(price,n,fee) } ) )
rownames(res_zlema) <- myrange
res_vwap <- t(sapply(3:200,function(n) { vwap(price,n,fee) } ) )
rownames(res_vwap) <- myrange
res_hma <- t(sapply(3:200,function(n) { hma(price,n,fee) } ) )
rownames(res_hma) <- myrange
res_alma <- t(sapply(3:200,function(n) { alma(price,n,fee) } ) )
rownames(res_alma) <- myrange
res_t3 <- t(sapply(3:200,function(n) { T3ma(price,n,fee) } ) )
rownames(res_t3) <- myrange

```

## Plots

```{r,plots1,fig.height=8,fig.width=8}

plot(rownames(res_sma),res_sma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_sma),res_sma[,1],pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_ema),res_ema[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_dema),res_dema[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="DEMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_dema),res_dema[,1],pch=19, xlim=c(10,60), xlab="moving average interval (days)",ylab="fold return",main="DEMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_wma),res_wma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="WMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_evwma),res_evwma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="EVWMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_zlema),res_zlema[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="ZLEMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_vwap),res_vwap[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="VWAP")
abline(h=hodl,col="red")
grid()

plot(rownames(res_hma),res_hma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="HMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_hma),res_hma[,1],pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="HMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_alma),res_alma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="ALMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_alma),res_alma[,1],pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="ALMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_t3),res_t3[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="T3ma")
abline(h=hodl,col="red")
grid()

plot(rownames(res_t3),res_t3[,1],pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="T3ma")
abline(h=hodl,col="red")
grid()

```

```{r,plots2,fig.height=8,fig.width=8}

plot(res_sma[,2],res_sma[,1],pch=19,cex=4,col="gray",main="SMA",xlab="n trades",ylab="fold return USD")
text(res_sma[,2],res_sma[,1],labels=rownames(res_sma),cex=1.5)
abline(h=hodl,col="red")

plot(res_ema[,2],res_ema[,1],pch=19,cex=4,col="gray",main="EMA",xlab="n trades",ylab="fold return USD")
text(res_ema[,2],res_ema[,1],labels=rownames(res_ema),cex=1.5)
abline(h=hodl,col="red")

plot(res_dema[,2],res_dema[,1],pch=19,cex=4,col="gray",main="DEMA",xlab="n trades",ylab="fold return USD")
text(res_dema[,2],res_dema[,1],labels=rownames(res_dema),cex=1.5)
abline(h=hodl,col="red")

plot(res_wma[,2],res_wma[,1],pch=19,cex=4,col="gray",main="WMA",xlab="n trades",ylab="fold return USD")
text(res_wma[,2],res_wma[,1],labels=rownames(res_wma),cex=1.5)
abline(h=hodl,col="red")

plot(res_evwma[,2],res_evwma[,1],pch=19,cex=4,col="gray",main="EVWMA",xlab="n trades",ylab="fold return USD")
text(res_evwma[,2],res_evwma[,1],labels=rownames(res_evwma),cex=1.5)
abline(h=hodl,col="red")

plot(res_zlema[,2],res_zlema[,1],pch=19,cex=4,col="gray",main="ZLEMA",xlab="n trades",ylab="fold return USD")
text(res_zlema[,2],res_zlema[,1],labels=rownames(res_zlema),cex=1.5)
abline(h=hodl,col="red")

plot(res_vwap[,2],res_vwap[,1],pch=19,cex=4,col="gray",main="VWAP",xlab="n trades",ylab="fold return USD")
text(res_vwap[,2],res_vwap[,1],labels=rownames(res_vwap),cex=1.5)
abline(h=hodl,col="red")

plot(res_hma[,2],res_hma[,1],pch=19,cex=4,col="gray",main="HMA",xlab="n trades",ylab="fold return USD")
text(res_hma[,2],res_hma[,1],labels=rownames(res_hma),cex=1.5)
abline(h=hodl,col="red")

plot(res_alma[,2],res_alma[,1],pch=19,cex=4,col="gray",main="ALMA",xlab="n trades",ylab="fold return USD")
text(res_alma[,2],res_alma[,1],labels=rownames(res_alma),cex=1.5)
abline(h=hodl,col="red")

plot(res_t3[,2],res_t3[,1],pch=19,cex=4,col="gray",main="T3",xlab="n trades",ylab="fold return USD")
text(res_t3[,2],res_t3[,1],labels=rownames(res_t3),cex=1.5)
abline(h=hodl,col="red")

```

## Restart analysis

Omitting the volume weighted ones because they aren't very good.
Also this allows us to examine more 2013 data where there isn't a good volume indicator.

```{r,redo,fig.height=8,fig.width=8}

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,price$quote$USD$volume,stringsAsFactors=FALSE)
colnames(price) <- c("date","close","Volume")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

myend <- price[c(1,nrow(price)),2][2]
mystart <- price[c(1,nrow(price)),2][1]
hodl <- (myend - mystart) / mystart
hodl

res_sma <- t(sapply(myrange,function(n) { sma(price,n,fee) } ))
rownames(res_sma) <- myrange
res_ema <- t(sapply(3:200,function(n) { ema(price,n,fee) } ) )
rownames(res_ema) <- myrange
res_dema <- t(sapply(3:200,function(n) { dema(price,n,fee) } ) )
rownames(res_dema) <- myrange
res_wma <- t(sapply(3:200,function(n) { wma(price,n,fee) } ) )
rownames(res_wma) <- myrange
#res_evwma <- t(sapply(3:200,function(n) { evwma(price,n,fee) } ) )
#rownames(res_evwma) <- myrange
res_zlema <- t(sapply(3:200,function(n) { zlema(price,n,fee) } ) )
rownames(res_zlema) <- myrange
#res_vwap <- t(sapply(3:200,function(n) { vwap(price,n,fee) } ) )
#rownames(res_vwap) <- myrange
res_hma <- t(sapply(3:200,function(n) { hma(price,n,fee) } ) )
rownames(res_hma) <- myrange
res_alma <- t(sapply(3:200,function(n) { alma(price,n,fee) } ) )
rownames(res_alma) <- myrange
res_t3 <- t(sapply(3:200,function(n) { T3ma(price,n,fee) } ) )
rownames(res_t3) <- myrange

plot(rownames(res_sma),res_sma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_sma),res_sma[,1],pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_ema),res_ema[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_dema),res_dema[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="DEMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_dema),res_dema[,1],pch=19,xlim=c(1,60),xlab="moving average interval (days)",ylab="fold return",main="DEMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_wma),res_wma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="WMA")
abline(h=hodl,col="red")
grid()

#plot(rownames(res_evwma),res_evwma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="EVWMA")
#abline(h=hodl,col="red")
#grid()

plot(rownames(res_zlema),res_zlema[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="ZLEMA")
abline(h=hodl,col="red")
grid()

#plot(rownames(res_vwap),res_vwap[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="VWAP")
#abline(h=hodl,col="red")
#grid()

plot(rownames(res_hma),res_hma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="HMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_hma),res_hma[,1],pch=19,xlim=c(25,75),xlab="moving average interval (days)",ylab="fold return",main="HMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_alma),res_alma[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="ALMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_alma),res_alma[,1],pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="ALMA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_t3),res_t3[,1],pch=19,xlab="moving average interval (days)",ylab="fold return",main="T3MA")
abline(h=hodl,col="red")
grid()

plot(rownames(res_t3),res_t3[,1],pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="T3MA")
abline(h=hodl,col="red")
grid()

plot(res_sma[,2],res_sma[,1],pch=19,cex=4,col="gray",main="SMA",xlab="n trades",ylab="fold return USD")
text(res_sma[,2],res_sma[,1],labels=rownames(res_sma),cex=1.5)
abline(h=hodl,col="red")

plot(res_ema[,2],res_ema[,1],pch=19,cex=4,col="gray",main="EMA",xlab="n trades",ylab="fold return USD")
text(res_ema[,2],res_ema[,1],labels=rownames(res_ema),cex=1.5)
abline(h=hodl,col="red")

plot(res_dema[,2],res_dema[,1],pch=19,cex=4,col="gray",main="DEMA",xlab="n trades",ylab="fold return USD")
text(res_dema[,2],res_dema[,1],labels=rownames(res_dema),cex=1.5)
abline(h=hodl,col="red")

plot(res_wma[,2],res_wma[,1],pch=19,cex=4,col="gray",main="WMA",xlab="n trades",ylab="fold return USD")
text(res_wma[,2],res_wma[,1],labels=rownames(res_wma),cex=1.5)
abline(h=hodl,col="red")

#plot(res_evwma[,2],res_evwma[,1],pch=19,cex=4,col="gray",main="EVWMA",xlab="n trades",ylab="fold return USD")
#text(res_evwma[,2],res_evwma[,1],labels=rownames(res_evwma),cex=1.5)
#abline(h=hodl,col="red")

plot(res_zlema[,2],res_zlema[,1],pch=19,cex=4,col="gray",main="ZLEMA",xlab="n trades",ylab="fold return USD")
text(res_zlema[,2],res_zlema[,1],labels=rownames(res_zlema),cex=1.5)
abline(h=hodl,col="red")

#plot(res_vwap[,2],res_vwap[,1],pch=19,cex=4,col="gray",main="VWAP",xlab="n trades",ylab="fold return USD")
#text(res_vwap[,2],res_vwap[,1],labels=rownames(res_vwap),cex=1.5)
#abline(h=hodl,col="red")

plot(res_hma[,2],res_hma[,1],pch=19,cex=4,col="gray",main="HMA",xlab="n trades",ylab="fold return USD")
text(res_hma[,2],res_hma[,1],labels=rownames(res_hma),cex=1.5)
abline(h=hodl,col="red")

plot(res_alma[,2],res_alma[,1],pch=19,cex=4,col="gray",main="ALMA",xlab="n trades",ylab="fold return USD")
text(res_alma[,2],res_alma[,1],labels=rownames(res_alma),cex=1.5)
abline(h=hodl,col="red")

plot(res_t3[,2],res_t3[,1],pch=19,cex=4,col="gray",main="T3MA",xlab="n trades",ylab="fold return USD")
text(res_t3[,2],res_t3[,1],labels=rownames(res_t3),cex=1.5)
abline(h=hodl,col="red")

```

## Optimisation

```{r,best_sma,fig.width=10,fig.height=8}

best_sma = as.numeric(rownames(res_sma)[which(res_sma==max(res_sma[,1]))])
message(paste("Best SMA period:",best_sma))
ma <- SMA(Cl(price),n=best_sma)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]  
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
sma_trades <- cbind(buy,sell)
sma_trades$diff <- ( ( sma_trades$sell_price - sma_trades$buy_price ) / sma_trades$buy_price ) +1 - (2 * fee )
message(paste("Number of trades:",nrow(sma_trades)))
message(paste("Total ROI:",prod(sma_trades$diff)))
message(paste("Returns over hodl:",prod(sma_trades$diff)/hodl))
SMA_HEADER=paste(best_sma,"day moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```


```{r,best_ema,fig.width=10,fig.height=8}

best_ema = as.numeric(rownames(res_ema)[which(res_ema==max(res_ema[,1]))])
message(paste("Best EMA period:",best_ema))
ma <- EMA(Cl(price),n=best_ema)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
ema_trades <- cbind(buy,sell)
ema_trades$diff <- ( ( ema_trades$sell_price - ema_trades$buy_price ) / ema_trades$buy_price ) +1 - ( 2 * fee )
message(paste("Number of trades:",nrow(ema_trades)))
message(paste("Total ROI:",prod(ema_trades$diff)))
message(paste("Returns over hodl:",prod(ema_trades$diff)/hodl))
EMA_HEADER=paste(best_ema,"day exponential moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```

```{r,best_dema,fig.width=10,fig.height=8}

best_dema = as.numeric(rownames(res_dema)[which(res_dema==max(res_dema[,1]))])
message(paste("Best DEMA period:",best_dema))
ma <- DEMA(Cl(price),n=best_dema)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
dema_trades <- cbind(buy,sell)
dema_trades$diff <- ( ( dema_trades$sell_price - dema_trades$buy_price ) / dema_trades$buy_price ) +1 - ( 2 * fee )
message(paste("Number of trades:",nrow(dema_trades)))
message(paste("Total ROI:",prod(dema_trades$diff)))
message(paste("Returns over hodl:",prod(dema_trades$diff)/hodl))
DEMA_HEADER=paste(best_dema,"day double exponential moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=DEMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```

```{r,best_hma,fig.width=10,fig.height=8}

best_hma = as.numeric(rownames(res_hma)[which(res_hma==max(res_hma[,1]))])
message(paste("Best HMA period:",best_hma))
ma <- HMA(Cl(price),n=best_hma)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
hma_trades <- cbind(buy,sell)
hma_trades$diff <- ( ( hma_trades$sell_price - hma_trades$buy_price ) / hma_trades$buy_price ) +1 - ( 2 * fee )
message(paste("Number of trades:",nrow(hma_trades)))
message(paste("Total ROI:",prod(hma_trades$diff)))
message(paste("Returns over hodl:",prod(hma_trades$diff)/hodl))
HMA_HEADER=paste(best_hma,"day Hull moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=HMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```

## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_sma_fees <- sapply(fee_levels,function(x) { sma(price,best_sma,fee=x) } )
plot(fee_levels,res_sma_fees[1,] ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

res_ema_fees <- sapply(fee_levels,function(x) { ema(price,best_ema,fee=x) } )
plot(fee_levels,res_ema_fees [1,],pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades$year <- sapply(strsplit(sma_trades[,3],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,5] } )
sma_yearly <- sma_yearly[which(lapply(sma_yearly,length)>0)]
vioplot(sma_yearly)
barplot(sapply(sma_yearly,prod))
abline(h=1,col="red",lty=2)

ema_trades$year <- sapply(strsplit(ema_trades[,3],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,5] } )
ema_yearly <- ema_yearly[which(lapply(ema_yearly,length)>0)]
vioplot(ema_yearly)
barplot(sapply(ema_yearly,prod))
abline(h=1,col="red",lty=2)

```


## Conclusion

Trading on moving averages is an easy way to increase returns on crypto volatility.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
