---
title: "Trading BTC based on MA and EMA crosses"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/MDZanalytics/cryptoblog/blob/main/ma_cross.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/MDZanalytics/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.


XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

Trading off simple moving averages (SMAs), exponenial moving averages (EMAs) and crosses between these 
lines are popular methods for timing entry and exit of positions.
In this blog post I will be back-testing simple moving averages to find the one which performed the best for
BTC's price history.

The function is modified so as to place greater emphasis on recent performance similar to the exponential
moving average (EMA).

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("reshape2")
  library("gplots")
  library("parallel")
  library("kableExtra")
})

NUMCORES=detectCores()/2

Sys.Date()

```

Reminder: this is not financial advice.

## Get data

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)

```

## Calculate hodl returns

It's important to calculate the returns for simply holding the asset rather than actively trading.

The number shown here is the fold return on investment.
BTC has appreciated a lot since 2013!

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Moving average cross function

Here are the functions for trading based on SMA and EMA crosses, buying when line 1 breaks through line2
and sell when it falls below.
All trades are executed at the end of the daily session (UTC midnight).

offset is the exponent for time 0. For example if it is 0.5 and return is 3x, the day1 trade is 3 ^ 0.5 = 1.7

Start with offset=0.5

multiplier is how much to emphasise the latest period over the earliest period.

Start with multiplier=1.

Trades in oldest time are worth 0.5x.
Very recent trades are worth 1.5x.
Trades in the middle of the time are worth 1x.

```{r,sma}

smacross <- function(price,n,fee,offset=0.5 , multiplier=1) {
n1=mas[n,1]
n2=mas[n,2]
if (n1 != n2) { 
  ma1 <- SMA(Cl(price),n=n1)
  ma2 <- SMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  time0 <- as.numeric(as.Date(price[1,1]))
  time1 <- as.numeric(as.Date(price[nrow(price),1]))
  timed <- time1-time0
  times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
  times <- rowMeans(times)
  times <- times - time0
  times <- times/timed * multiplier
  times <- times + offset
  trades$times <- times
  trades$adj_diff <- trades$diff ^ trades$times
  return(c(prod(trades$diff), prod(trades$adj_diff), nrow(trades)))
} else {
  return(c(1,1,0))
}
}

emacross <- function(price,n,fee,offset=0.5 , multiplier=1) {
n1=mas[n,1]
n2=mas[n,2]
if (n1 != n2) {
  ma1 <- EMA(Cl(price),n=n1)
  ma2 <- EMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  time0 <- as.numeric(as.Date(price[1,1]))
  time1 <- as.numeric(as.Date(price[nrow(price),1]))
  timed <- time1-time0
  times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
  times <- rowMeans(times)
  times <- times - time0
  times <- times/timed * multiplier
  times <- times + offset
  trades$times <- times
  trades$adj_diff <- trades$diff ^ trades$times
  return(c(prod(trades$diff), prod(trades$adj_diff), nrow(trades)))
} else {
  return(c(1,1,0))
}
}

```

## Backtesting

Now let's go back in time to test the performance of the different MA cross strategies on the daily
timeframe, using MA intervals from 4 to 200 days.

Let's assume a 0.1% trading fee, as per binance current rate.

First looking at SMA crosses.
Below, I show the 20 highest returning indicators.

```{r,backtestsma}

fee=0.005

mas <- expand.grid(5:200,5:200)
mas <- mas[which( mas$Var2 < mas$Var1) ,]

res_sma <- mclapply(1:nrow(mas),function(n) {
 smacross(price,n,fee) 
} , mc.cores=NUMCORES)

res_sma <- do.call(rbind,res_sma)
mas1 <- as.data.frame(mas,stringsAsFactors=FALSE)
res_sma <- cbind(mas1,res_sma)
colnames(res_sma) <- c("MA1","MA2","returns","returns_adjusted","trades")
res_sma$mean <- res_sma$returns^(1/res_sma$trades)
res_sma$mean_adjusted <- res_sma$returns_adjusted^(1/res_sma$trades)

best_sma <- res_sma[order(-res_sma$mean_adjusted),]
head(best_sma,20) %>% kbl(caption="Best ROI SMA strategies") %>%  kable_styling()

best_sma <- head(res_sma[order(-res_sma$returns_adjusted),],20)
best_sma %>% kbl(caption="Highest returns SMA strategies") %>%  kable_styling()

```

Now let's take a look at the charts for SMA crosses.
The darker the colour, the higher the returns.

```{r,plots1,fig.height=8,fig.width=8}

w <- as.matrix(acast(res_sma, MA1~MA2, value.var="returns_adjusted"))

colfunc <- colorRampPalette(c("white", "yellow" , "orange" , "red" , "darkred"))

colfunc <- colorRampPalette(c("white", "red"))

heatmap.2(w,scale="none",trace="none",dendrogram="none", Rowv=FALSE, Colv=FALSE,
  col=colfunc(25), main="SMA crosses", xlab="short period", ylab="long period")

heatmap.2(w[7:43,1:20],scale="none",trace="none",dendrogram="none", Rowv=FALSE,
  Colv=FALSE, col=colfunc(25) , main="SMA crosses",
  xlab="short period", ylab="long period")

```

Now we'll do the backtesting for EMA crosses.
Here are the best 20 EMA cross strategies.

```{r,backtestema}

res_ema <- mclapply(1:nrow(mas),function(n) {
 emacross(price,n,fee)
} , mc.cores=NUMCORES)
res_ema <- do.call(rbind,res_ema)
mas1 <- as.data.frame(mas,stringsAsFactors=FALSE)
res_ema <- cbind(mas1,res_ema)
colnames(res_ema) <- c("MA1","MA2","returns","returns_adjusted","trades")
res_ema$mean <- res_ema$returns^(1/res_ema$trades)
res_ema$mean_adjusted <- res_ema$returns_adjusted^(1/res_ema$trades)

best_ema <- res_ema[order(-res_ema$mean_adjusted),]    
head(best_ema,20) %>% kbl(caption="Best ROI EMA strategies") %>%  kable_styling()

best_ema <- head(res_ema[order(-res_ema$returns_adjusted),],20)
best_ema %>% kbl(caption="Highest returns EMA strategies") %>%  kable_styling()

```

Now let's take a look at the charts for EMA crosses.

```{r,plots2,fig.height=8,fig.width=8}

w <- as.matrix(acast(res_ema, MA1~MA2, value.var="returns_adjusted"))

colfunc <- colorRampPalette(c("white", "yellow" , "orange" , "red" , "darkred"))

colfunc <- colorRampPalette(c("white", "red"))

heatmap.2(w,scale="none",trace="none",dendrogram="none", Rowv=FALSE, Colv=FALSE,
  col=colfunc(25), main="EMA crosses", xlab="short period", ylab="long period")

heatmap.2(w[7:50,1:30],scale="none",trace="none",dendrogram="none", Rowv=FALSE,
  Colv=FALSE, col=colfunc(25) , main="EMA crosses",
  xlab="short period", ylab="long period")

```

## SMA cross optimisation

Now let's take a closer look at the best SMA cross strategy.

```{r,best_sma,fig.width=10,fig.height=8}

n1=best_sma[1,1]
n2=best_sma[1,2]
message(paste("Best short SMA period:",n2))
message(paste("Best long SMA period:",n1))
if (n1 != n2) {
  ma1 <- SMA(Cl(price),n=n1)
  ma2 <- SMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  time0 <- as.numeric(as.Date(price2[1,1]))
  time1 <- as.numeric(as.Date(price2[nrow(price2),1]))
  timed <- time1-time0
  times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
  times <- rowMeans(times)
  times <- times - time0
  times <- times/timed
  times <- times + 0.5
  trades$times <- times
  trades$adj_diff <- trades$diff ^ trades$times
  returns <- prod(trades$adj_diff)
} else {
  returns <- 1
}
sma_trades <- trades
message(paste("Number of trades:",nrow(sma_trades)))
message(paste("Total ROI:",prod(sma_trades$adj_diff)))
message(paste("Returns over hodl:",prod(sma_trades$adj_diff)/hodl))

```

Here is a break down of all the trades executed for the best SMA cross strategy.

```{r,best_sma2,fig.width=10,fig.height=8}

sma_trades %>% kbl() %>%  kable_styling()

```

Here is a chart showing the SMA cross indicator.

```{r,best_sma3,fig.width=10,fig.height=8}
SMA_HEADER=paste(n2 , "and", n1 ,"period SMA cross")
plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma1~ as.Date(price$date) ,col="red")
lines(ma2~ as.Date(price$date) ,col="blue")

```

### SMA cross recent action

```{r,sma_recent}

plot(tail(price$Close,200)~tail(as.Date(price$date),200),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(tail(ma1,200) ~ tail(as.Date(price$date),200) ,col="red")
lines(tail(ma2,200) ~ tail(as.Date(price$date),200) ,col="blue")

```

## EMA cross optimisation

Same thing for the best EMA cross strategy.

```{r,best_ema,fig.width=10,fig.height=8}

n1=best_ema[1,1]
n2=best_ema[1,2]
message(paste("Best short EMA period:",n2))
message(paste("Best long EMA period:",n1))
if (n1 != n2) {
  ma1 <- EMA(Cl(price),n=n1)
  ma2 <- EMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  time0 <- as.numeric(as.Date(price2[1,1]))
  time1 <- as.numeric(as.Date(price2[nrow(price2),1]))
  timed <- time1-time0
  times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
  times <- rowMeans(times)
  times <- times - time0
  times <- times/timed
  times <- times + 0.5
  trades$times <- times
  trades$adj_diff <- trades$diff ^ trades$times
  returns <- prod(trades$diff)
} else {
  returns <- 1
}
ema_trades <- trades
message(paste("Number of trades:",nrow(ema_trades)))
message(paste("Total ROI:",prod(ema_trades$adj_diff)))
message(paste("Returns over hodl:",prod(ema_trades$adj_diff)/hodl))

```

Here is a break down of all the trades executed for the best EMA cross strategy.

```{r,best_ema2,fig.width=10,fig.height=8}

ema_trades %>% kbl() %>%  kable_styling()

```

Here is a chart showing the EMA cross indicator.

```{r,best_ema3,fig.width=10,fig.height=8}

EMA_HEADER=paste(n2 , "and", n1 ,"period EMA cross")
plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(ma1~ as.Date(price$date) ,col="red")
lines(ma2~ as.Date(price$date) ,col="blue")

```

### EMA cross recent action

```{r,ema_recent}

plot(tail(price$Close,200)~tail(as.Date(price$date),200),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(tail(ma1,200) ~ tail(as.Date(price$date),200) ,col="red")
lines(tail(ma2,200) ~ tail(as.Date(price$date),200) ,col="blue")

```

## Effect of fees

It is important to factor in the effect of fees because higher exchange fees will erode profits.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_sma_fees <- sapply(fee_levels,function(x) {
  smacross(price,rownames(best_sma)[1],fee=x)
} )
plot(fee_levels,res_sma_fees[2,] ,pch=19,xlab="fee proportion",ylab="fold return",
  main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

res_ema_fees <- sapply(fee_levels,function(x) {
  emacross(price,rownames(best_ema)[1],fee=x)
} )
plot(fee_levels,res_ema_fees[2,] ,pch=19,xlab="fee proportion",ylab="fold return",
  main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

It is good to see whether the strategy gives consistent returns over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades$year <- sapply(strsplit(sma_trades[,"sell_date"],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,"adj_diff"] } )
sma_yearly <- sma_yearly[which(lapply(sma_yearly,length)>0)]
vioplot(sma_yearly,ylab="trade ROI distribution",main="BTC daily SMA cross strategy")

yr <- sapply(sma_yearly,prod)
barplot(yr,ylab="yearly fold returns",main="BTC daily SMA cross strategy")
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

ema_trades$year <- sapply(strsplit(ema_trades[,"sell_date"],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,"adj_diff"] } )
ema_yearly <- ema_yearly[which(lapply(ema_yearly,length)>0)]
vioplot(ema_yearly,ylab="trade ROI distribution",main="BTC daily EMA cross strategy")

yr <- sapply(ema_yearly,prod)
barplot(yr,ylab="yearly fold returns",main="BTC daily EMA cross strategy")
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```


## Conclusion

Best SMA cross strategy (as at last script edit) was 12/21d, with returns roughly double the best single SMA strategy (43d), achieved with fewer trades.
Best EMA cross strategy was 27/48d and got returns 2.9 times hodl, but with only 15 trades.
In conclusion, these strategies give better returns than single MA strategies.

If you would like to receive trade alerts on these strategies, let me know by email info@mdz-analytics.com.
These alerts are sent at 30 mins before the close of the daily session and will assist you in sticking to your desired strategy.
Subscribing to alerts costs 0.0005 BTC per year.

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,sessioninfo}

sessionInfo()

```

</details>
