---
title: "Temporal changes to performance of BTC based on MAs signals over time."
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/ma_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.


XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

Trading off simple moving averages (SMA), exponenial moving averages (EMA) and crosses between these is 
a popular method for timing entry and exit of positions.
In this blog post I will be exploring MA strategies on the daily timeframe to find the ones which 
perform best.

It is important to understand whether backtested indicators have good performance going forward.
Therefore the price data will be split into two parts, period 0 and period 1.
We'll see whether indicators based on early price data yield good performance later on.
If it turns out that the performance of indicators changes dramatically, then more investigation 
on these changing trends is required, so we can remain ahead of the curve.

This report is time-sensitive - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
})

Sys.Date()

```

Reminder: this is not financial advice.

## Get data

Historical data are from http://coinmarketcap.com/


```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

Split price action into two parts.

```{r,split}

midpoint <- round(nrow(price)/2)
price0 <- price[1:midpoint,]
price1 <- price[midpoint:nrow(price),]

price0[c(1:5,(nrow(price0)-5):nrow(price0)),]  %>% kbl(caption="period0 range") %>%  kable_paper("hover", full_width = F)
price1[c(1:5,(nrow(price1)-5):nrow(price1)),]  %>% kbl(caption="period1 range") %>%  kable_paper("hover", full_width = F)

```

## Calculate hodl returns

It's important to calculate the returns for simply holding the asset rather than actively trading.

The first number is the ROI since 2013.
The second number is for period0 and third number is for period1.

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

myend <- price0[c(1,nrow(price0)),2][2]
mystart <- price0[c(1,nrow(price0)),2][1]
hodl0 <- (myend - mystart) / mystart
hodl0

myend <- price1[c(1,nrow(price1)),2][2]
mystart <- price1[c(1,nrow(price1)),2][1]
hodl1 <- (myend - mystart) / mystart
hodl1

```

## Moving average function

Here are the functions for trading based on SMA and EMAs, buying when the price breaks through the
MA line and sell when it falls below the MA line. All trades are executed at the end of the daily session (UTC midnight).

```{r,functions}

sma <- function(price,n,fee) {
ma <- SMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

ema <- function(price,n,fee) {
ma <- EMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

```

## Backtesting

Let's look back in time and trial SMA and EMA strategies with an interval from 3 to 200 days.
This is performed separately for period0 and period1.

Let's assume a 0.1% trading fee, as per binance current rate.


```{r,backtest}

fee=0.001

myrange <- 3:200

# sma old
res <- sapply(myrange,function(n) { sma(price0,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns0","trades0")
res <- as.data.frame(res)
res$mean0 <- res$returns0^(1/res$trades0)
res_sma0 <- res
res_sma0$ma0 <- rownames(res_sma0)
res_sma0$xhodl0 <- res_sma0$returns0 / hodl0
# sma new
res <- sapply(myrange,function(n) { sma(price1,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns1","trades1")
res <- as.data.frame(res)
res$mean1 <- res$returns1^(1/res$trades1)
res_sma1 <- res
res_sma1$ma1 <- rownames(res_sma1)
res_sma1$xhodl1 <- res_sma1$returns1 / hodl1
# join
res_sma <- cbind(res_sma0,res_sma1)

res_sma %>% kbl(caption="SMA trading strategies in period 0 and 1") %>%  kable_paper("hover", full_width = F)

# ema old
res <- sapply(myrange,function(n) { ema(price0,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns0","trades0")
res <- as.data.frame(res)
res$mean0 <- res$returns0^(1/res$trades0)
res_ema0 <- res
res_ema0$ma0 <- rownames(res_ema0)
res_ema0$xhodl0 <- res_ema0$returns0 / hodl0
# ema new
res <- sapply(myrange,function(n) { ema(price1,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns1","trades1")
res <- as.data.frame(res)
res$mean1 <- res$returns1^(1/res$trades1)
res_ema1 <- res
res_ema1$ma1 <- rownames(res_ema1)
res_ema1$xhodl1 <- res_ema1$returns1 / hodl1
# join
res_ema <- cbind(res_ema0,res_ema1)

res_ema %>% kbl(caption="EMA trading strategies in period 0 and 1") %>%  kable_paper("hover", full_width = F)

```

## Plots

Now let's look at which of these approaches works best.

```{r,plots1,fig.height=8,fig.width=8}

# SMA
par(mfrow=c(2,1))
plot(res_sma$ma0,res_sma$returns0,
  pch=19,cex=2,col="gray",
  xlab="moving average interval (days)",ylab="fold return",main="SMA period0")
text(res_sma$ma0,res_sma$returns0,labels=res_sma$ma0,pch=19,cex=1,col="black")
grid()
abline(h=hodl0,col="red")

plot(res_sma$ma1,res_sma$returns1,
  pch=19,cex=2,col="gray", 
  xlab="moving average interval (days)",ylab="fold return",main="SMA period1")
text(res_sma$ma1,res_sma$returns1,labels=res_sma$ma1,pch=19,cex=1,col="black")
grid()
abline(h=hodl1,col="red")

par(mfrow=c(2,1))
plot(res_sma$ma0,res_sma$returns0,
  pch=19,cex=2,col="gray", xlim=c(0,50),
  xlab="moving average interval (days)",ylab="fold return",main="SMA period0")
text(res_sma$ma0,res_sma$returns0,labels=res_sma$ma0,pch=19,cex=1,col="black")
grid()
abline(h=hodl0,col="red")

plot(res_sma$ma1,res_sma$returns1,
  pch=19,cex=2,col="gray", xlim=c(0,50),
  xlab="moving average interval (days)",ylab="fold return",main="SMA period1")
text(res_sma$ma1,res_sma$returns1,labels=res_sma$ma1,pch=19,cex=1,col="black")
grid()
abline(h=hodl1,col="red")

# EMA
par(mfrow=c(2,1))
plot(res_ema$ma0,res_ema$returns0,
  pch=19,cex=2,col="gray",
  xlab="moving average interval (days)",ylab="fold return",main="EMA period0")
text(res_ema$ma0,res_ema$returns0,labels=res_ema$ma0,pch=19,cex=1,col="black")
grid()
abline(h=hodl0,col="red")

plot(res_ema$ma1,res_ema$returns1,
  pch=19,cex=2,col="gray",
  xlab="moving average interval (days)",ylab="fold return",main="EMA period1")
text(res_ema$ma1,res_ema$returns1,labels=res_ema$ma1,pch=19,cex=1,col="black")
grid()
abline(h=hodl1,col="red")

par(mfrow=c(2,1))
plot(res_ema$ma0,res_ema$returns0,
  pch=19,cex=2,col="gray", xlim=c(0,50),
  xlab="moving average interval (days)",ylab="fold return",main="EMA period0")
text(res_ema$ma0,res_ema$returns0,labels=res_ema$ma0,pch=19,cex=1,col="black")
grid()
abline(h=hodl0,col="red")

plot(res_ema$ma1,res_ema$returns1,
  pch=19,cex=2,col="gray", xlim=c(0,50),
  xlab="moving average interval (days)",ylab="fold return",main="EMA period1")
text(res_ema$ma1,res_ema$returns1,labels=res_ema$ma1,pch=19,cex=1,col="black")
grid()
abline(h=hodl1,col="red")

```

Some of those gains look very high, but the time frames are quite short.
Could it mean that there are a lot of trades each with very small gains?

```{r,plots2,fig.height=8,fig.width=8}

par(mfrow=c(2,1))
plot(res_sma$mean0,res_sma$returns0, main="SMA period0",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$mean0,res_sma$returns0 , cex=1, labels=res_sma$ma0)
abline(h=hodl0,col="red")

plot(res_sma$mean1,res_sma$returns1, main="SMA period1",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$mean1,res_sma$returns1 , cex=1, labels=res_sma$ma1)
abline(h=hodl1,col="red")

head(res_sma[order(-res_sma$returns0),],10) %>% kbl(caption="Highest total ROI SMA strategy") %>%  kable_styling()

head(res_sma[order(-res_sma$mean0),],10) %>% kbl(caption="Highest Mean ROI SMA strategy for period1") %>%  kable_styling()

par(mfrow=c(2,1))
plot(res_ema$mean0,res_ema$returns0, main="EMA period0",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$mean0,res_ema$returns0 , cex=1, labels=res_ema$ma0)
abline(h=hodl0,col="red")

plot(res_ema$mean1,res_ema$returns1, main="EMA period1",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$mean1,res_ema$returns1 , cex=1, labels=res_ema$ma1)
abline(h=hodl1,col="red")

head(res_ema[order(-res_ema$returns0),],10) %>% kbl(caption="Highest total ROI EMA strategy") %>%  kable_styling()

head(res_ema[order(-res_ema$mean0),],10) %>% kbl(caption="Highest Mean ROI EMA strategy for period1") %>%  kable_styling()

```

Now we'll have a look at how many trades are required for these approaches.
Ideally we don't want to be making trades every few days.

```{r,plots3,fig.height=8,fig.width=8}

# SMA
par(mfrow=c(2,1))
plot(res_sma$trades0,res_sma$returns0, main="SMA period0",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$trades0,res_sma$returns0 , cex=1, labels=res_sma$ma0)
abline(h=hodl0,col="red")

plot(res_sma$trades1,res_sma$returns1, main="SMA period1",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$trades1,res_sma$returns1 , cex=1, labels=res_sma$ma1)
abline(h=hodl1,col="red")

# EMA
par(mfrow=c(2,1))
plot(res_ema$trades0,res_ema$returns0, main="EMA period0",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$trades0,res_ema$returns0 , cex=1, labels=res_ema$ma0)
abline(h=hodl0,col="red")

plot(res_ema$trades1,res_ema$returns1, main="EMA period1",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$trades1,res_ema$returns1 , cex=1, labels=res_ema$ma1)
abline(h=hodl1,col="red")

```

## Focus on indicators with good recent performance

Doing this will focus more on indicators which are giving consistent gains over time,
and bias for indicators with good recent performance.

```{r,av1}

res_sma$xhodl_mean <- (res_sma$xhodl0 + res_sma$xhodl1 + res_sma$xhodl1) / 3

head(res_sma[order(-res_sma$xhodl_mean),],10) %>% 
  kbl(caption="Highest mean ROI SMA strategy for period0 and period1") %>%  kable_styling()

res_ema$xhodl_mean <- (res_ema$xhodl0 + res_ema$xhodl1 + res_ema$xhodl1) / 3

head(res_ema[order(-res_ema$xhodl_mean),],10) %>% 
  kbl(caption="Highest mean ROI EMA strategy for period0 and period1") %>%  kable_styling()


```

## Conclusion

Quite a bit to unpack here.
The first one is that SMA based signals give better returns than EMA, at least in terms of mean ROI.
Just compare the mean ROI of SMA38d vs EMA14d.

The best SMA for period0 was the 42D but in period1 this same indicator gave results only slightly better than hodling.
In period1, the best SMA was the 39D, but in period0 it gave results only slightly better than hodling.

This work suggests that indicator performance does change quite a bit over the years.
The indicators based on "old" price data don't perform the best, but they stil perform better than hodl.
But this is good because with further work, we will be able to see the changing trends over time.
We'll be able to use indicators that have good recent performance and be able to guess what indicators will 
perform better in future.

If you would like to receive trade alerts based on these strategies, let me know by email mark.ziemann{αt}gmail.com.
These alerts are sent at 30 mins before the close of the daily session and will assist you in sticking to your desired strategy.
Subscribing to alerts costs 0.0005 BTC per year.

If you would like help with backtesting, I can provide coding assistance for a modest price.

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,sessioninfo}

sessionInfo()

```

</details>
