---
title: "Trading BTC based on the Stochastic indicator"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/stoch_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.


XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

Here, we're using the stochastic as a buy/sell indicator for trading BTC.

The backtest function is modified so as to place greater emphasis on recent performance similar to the exponential
moving average (EMA).

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library("parallel")
})

Sys.Date()

CORES=detectCores()/2

```

Reminder: this is not financial advice.

## Get data

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```
## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),4][2] 
mystart <- price[c(1,nrow(price)),4][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Basic look at the indicator

Default indicator.
Slightly too fast.

```{r,sto1,fig.height=10,fig.width=8}

price2 <- tail(price,200)
sto <- stoch(HLC(price2), nFastK = 14, nFastD = 3, nSlowD = 3, bounded = TRUE, smooth=1)

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(sto[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="stoch",main="stoch fast")
lines(as.Date(price2$date), sto[,2]  , col="red"  )
grid()

fee = 0.001
offset=0.5
multiplier=1

sto <- stoch(HLC(price), nFastK=14 , nFastD=3 , nSlowD=3 , bounded = TRUE, smooth=1)
price2 <- price
price2 <- cbind(price2,sto)
price2$higher <- price2$fastK > price2$fastD
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

time0 <- as.numeric(as.Date(price[1,1]))
time1 <- as.numeric(as.Date(price[nrow(price),1]))
timed <- time1-time0
times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
times <- rowMeans(times)
times <- times - time0
times <- times/timed * multiplier
times <- times + offset
trades$times <- times
trades$adj_diff <- trades$diff ^ trades$times

basic_mean <- mean(trades$adj_diff)
basic_returns <- prod(trades$adj_diff)

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$adj_diff)))
message(paste("Returns over hodl:",prod(trades$adj_diff)/hodl))

```

Now with the slower settings.

```{r,sto2,fig.height=10,fig.width=8}

price2 <- tail(price,200)
sto <- stoch(HLC(price2), nFastK = 21, nFastD = 14, nSlowD = 21, bounded = TRUE, smooth=7)

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(sto[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="stoch",main="stoch slow")
lines(as.Date(price2$date), sto[,2]  , col="red"  )
grid()

par(mfrow=c(1,1))

sto <- stoch(HLC(price), nFastK=21 , nFastD=14 , nSlowD=21 , bounded = TRUE, smooth=7)
price2 <- price
price2 <- cbind(price2,sto)
price2$higher <- price2$fastK > price2$fastD
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
time0 <- as.numeric(as.Date(price[1,1]))
time1 <- as.numeric(as.Date(price[nrow(price),1]))
timed <- time1-time0
times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
times <- rowMeans(times)
times <- times - time0
times <- times/timed * multiplier
times <- times + offset
trades$times <- times
trades$adj_diff <- trades$diff ^ trades$times

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$adj_diff)))
message(paste("Returns over hodl:",prod(trades$adj_diff)/hodl))

```

## Stoch trading function

Assuming 0.1% trading fee, as per binance current rate.

offset is the exponent for time 0. For example if it is 0.5 and return is 3x, the day1 trade is 3 ^ 0.5 = 1.7

Start with offset=0.5

multiplier is how much to emphasise the latest period over the earliest period.

Start with multiplier=1.

Trades in oldest time are worth 0.5x.
Very recent trades are worth 1.5x.
Trades in the middle of the time are worth 1x.

```{r,stofunc1}

mystoch <- function(price,nFastK=14,nFastD=3,nSlowD=3,smooth=1,fee, offset=0.5 , multiplier=1) {
  sto <- stoch(HLC(price), nFastK , nFastD , nSlowD , bounded = TRUE, smooth=smooth)
  price2 <- price
  price2 <- cbind(price2,sto)
  price2$higher <- price2$fastK > price2$fastD
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  time0 <- as.numeric(as.Date(price[1,1]))
  time1 <- as.numeric(as.Date(price[nrow(price),1]))
  timed <- time1-time0
  times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
  times <- rowMeans(times)
  times <- times - time0
  times <- times/timed * multiplier
  times <- times + offset
  trades$times <- times
  trades$adj_diff <- trades$diff ^ trades$times
  return(c(prod(trades$diff), prod(trades$adj_diff), nrow(trades)))
}

```

## Backtesting

Trying a range of values for nfastk,nfastd and nslowd

The slowD will be set to the same as fast D.

```{r,backtest1}

fee=0.001

nFastK <- seq(2,50,5)
nFastD <- seq(2,80,5)
smooth <- seq(1,80,5)
params <- expand.grid(nFastK,nFastD,smooth)
colnames(params) <- c("nFastK","nFastD","smooth")
res <- mclapply(X=1:nrow(params) , FUN=function(i) {
  nFastK=as.numeric(params[i,1])
  nFastD=as.numeric(params[i,2])
  nSlowD=as.numeric(params[i,2])
  smooth=as.numeric(params[i,3])
  mystoch(price,nFastK=nFastK,nFastD=nFastD,nSlowD=nSlowD,smooth=smooth,fee=fee)
} , mc.cores = CORES)
res <- as.data.frame(do.call(rbind, res))
colnames(res) <- c("returns","returns_adjusted","trades")
res2 <- cbind(params,res)
res2$mean <- res2$returns^(1/res2$trades)
res2$mean_adjusted <- res2$returns_adjusted^(1/res2$trades)

```

Strategies with best mean trade ROI.
These don't occur frequently, but when they do, it gives a solid return.

```{r,res1}

head(res2[order(-res2$mean),],20) %>% kbl(caption="highest mean ROI strategies (unadjusted)") %>%  kable_styling("hover", full_width = F)

head(res2[order(-res2$mean_adjusted),],20) %>% kbl(caption="highest mean ROI strategies (adjusted)") %>%  kable_styling("hover", full_width = F)


```

Strategies with the highest returns - these are too frequent and each trade has only small returns.

```{r,res2}

head(res2[order(-res2$returns),],20) %>% kbl(caption="highest total ROI strategies (unadjusted)") %>%  kable_styling("hover", full_width = F)

head(res2[order(-res2$returns_adjusted),],20) %>% kbl(caption="highest total ROI strategies (adjusted)") %>%  kable_styling("hover", full_width = F)

```

Simply exclude any strategies with more than 200 trades and then look at the ones with the highest
returns.

```{r,res3}

res3 <- subset(res2,trades<200)

head(res3[order(-res3$returns),],20) %>% kbl(caption="highest total ROI strategies with <200 trades (unadjusted)") %>%  kable_styling("hover", full_width = F)

head(res3[order(-res3$returns_adjusted),],20) %>% kbl(caption="highest total ROI strategies with <200 trades (adjusted)") %>%  kable_styling("hover", full_width = F)

```

Now let's have a look at a chart of the total returns and mean ROI of all the strategies tested.
An optimal strategy would be towards the upper right side of the chart.
The default strategy is shown in red.
In the second chart, I'm looking at the strategies with <200 trades since 2013.
Any strategy with mean ROI > 1.1 is considered good. 

```{r,plots1,fig.height=8,fig.width=8}

plot(res2$mean , res2$returns, log="y",xlab="mean ROI", ylab="total returns")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()
mtext("unadjusted")

plot(res2$mean_adjusted , res2$returns_adjusted, log="y",xlab="mean ROI", ylab="total returns")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()
mtext("unadjusted")

plot(res3$mean , res3$returns, log="y",xlab="mean ROI", ylab="total returns, <200 trades")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()
mtext("unadjusted")

plot(res3$mean_adjusted , res3$returns_adjusted, log="y",xlab="mean ROI", ylab="total returns, <200 trades")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()
mtext("adjusted")

```

Here's the table of results for the strategies with mean trade ROI of >108% and highest overall ROI.

```{r,res4}

res5 <- subset(res2, mean > 1.08 )
res5 <- res5[order(-res5$returns_adjusted),]
res5 <- head(res5,20)
res5 %>% kbl(caption="Best total ROI strategies with mean ROI>1.08 (adjusted)") %>%  kable_stying("hover", full_width = F)

```

Now that I know which ranges of values work well, I will zoom in and do a more comprehensive
backtesting for the following values nfastk(32:52),nfastd/nslowd(42:82) and smooth (1:5).

```{r,backtest2}

fee=0.001

nFastK <- seq(min(res5$nFastK),max(res5$nFastK),1)
nFastD <- seq(min(res5$nFastD),max(res5$nFastD),1)
smooth <- seq(min(res5$smooth),max(res5$smooth),1)
params <- expand.grid(nFastK,nFastD,smooth)
colnames(params) <- c("nFastK","nFastD","smooth")
res <- mclapply(X=1:nrow(params) , FUN=function(i) {
  nFastK=as.numeric(params[i,1])
  nFastD=as.numeric(params[i,2])
  nSlowD=as.numeric(params[i,2])
  smooth=as.numeric(params[i,3])
  mystoch(price,nFastK=nFastK,nFastD=nFastD,nSlowD=nSlowD,smooth=smooth,fee=fee)
} , mc.cores = CORES)
res <- as.data.frame(do.call(rbind, res))
colnames(res) <- c("returns","returns_adjusted","trades")
res2 <- cbind(params,res)
res2$mean <- res2$returns^(1/res2$trades)
res2$mean_adjusted <- res2$returns_adjusted^(1/res2$trades)

```

Here are the highest returning strategies with fewer than 200 trades.

```{r,res6}

res3 <- subset(res2,trades<200)
res3 <- res3[order(-res3$returns_adjusted),]
head(res3,10) %>% kbl(caption="highest returning strategies with fewer than 200 trades(adjusted)") %>%  kable_styling("hover", full_width = F)

```

## Optimisation

This one looks to be the best: nFastK=7, nSlowD/nFastD=45, smooth=17

I have tested the best slowD and varying it did not make any improvement.

Below I show the results for this strategy, charts and each individual trade.

```{r,best1,fig.width=10,fig.height=8}

best_nFastK = res3[1,1]
best_nFastD = res3[1,2]
best_smooth = res3[1,3]

message(paste("Best nFastK:",best_nFastK))
message(paste("Best nFastD:",best_nFastD))
message(paste("Best smooth:",best_smooth))

sto <- stoch(HLC(price), nFastK=best_nFastK , nFastD=best_nFastD , nSlowD=best_nFastD , 
  bounded = TRUE, smooth=best_smooth)
price2 <- price
price2 <- cbind(price2,sto)
price2$higher <- price2$fastK > price2$fastD
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
time0 <- as.numeric(as.Date(price[1,1]))
time1 <- as.numeric(as.Date(price[nrow(price),1]))
timed <- time1-time0
times <- cbind(as.numeric(as.Date(trades$buy_date)),as.numeric(as.Date(trades$sell_date)))
times <- rowMeans(times)
times <- times - time0
times <- times/timed * multiplier
times <- times + offset
trades$times <- times
trades$adj_diff <- trades$diff ^ trades$times

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$adj_diff)))
message(paste("Returns over hodl:",prod(trades$adj_diff)/hodl))

HEADER="STOCH"
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(sto[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="stoch",main="Stoch")
lines(as.Date(price$date), sto[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

trades %>% kbl() %>%  kable_styling("hover", full_width = F)

```

## Recent action

```{r,recentaction}

HEADER="STOCH"
par(mfrow=c(2,1))
plot(tail(price$close,200)~tail(as.Date(price$date),200),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tail(sto[,1],200) ~ tail(as.Date(price$date),200),type="l",col="blue",
  xlab="Date",ylab="stoch",main="Stoch")
lines(as.Date(price$date), sto[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Effect of fees

It is important to factor in the effect of fees because higher exchange fees will erode profits.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_fees <- sapply(fee_levels,function(x) {
  mystoch(price,nFastK=best_nFastK,nFastD=best_nFastD,nSlowD=best_nFastD,smooth=best_smooth,fee=x)
} )

plot(fee_levels,res_fees[2,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext("higher time frame")
abline(h=hodl,col="red")
grid()

```

## Trades by year

It is good to see whether the strategy gives consistent returns over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

trades$year <- sapply(strsplit(trades[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades,year==y)[,7] } )
yearly <- yearly[which(lapply(yearly,length)>0)]
vioplot(yearly)
yr <- sapply(yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```

## Conclusion

The stochastic gave moderate returns with 2.6x hodl which was less than the SMA strategy (~6x hodl).

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,sessioninfo}

sessionInfo()

```

</details>
