---
title: "Daily trade alerts for BTC and ETH"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()+1`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/alerts.Rmd

If you find this resource useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

In the backtesting analysis, I have established three different strategies for timing entry/exits.
Click on the hyperlinks below to see how these signals were identified.

1. Simple moving averages: <a href="https://mdz-analytics.com/crypto/ma_trading.html">BTC</a> 43 day; 
<a href="https://mdz-analytics.com/crypto/ma_trading_eth.html">ETH</a> 34 day.

2. Directional movement index (DMI): <a href="https://mdz-analytics.com/crypto/dmi_trading.html">BTC</a> 26 day;
<a href="https://mdz-analytics.com/crypto/dmi_trading_eth.html">ETH</a> 19 day.

3. True strength index (TSI): <a href="https://mdz-analytics.com/crypto/tsi_trading.html">BTC</a> n1=99, n2=15, n.signal=5;
<a href="https://mdz-analytics.com/crypto/tsi_trading.html">ETH</a> n1=99, n2=3, n.signal=13.

This script will examine the price action just before the daily close and then investigate whether these trading signals are flashing.
Note that alerts for TSI are not ready yet.
Alerts will be sent by pushbullet.

If you would like to receive these signals in real-time, you can subscribe for 0.0005 BTC per year.
If you need some help to design your own trading signals, I am happy to help for a fee.
email me at mark.ziemann{αt}gmail.com for any enquiries.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("RPushbullet")
  library("kableExtra")
})

TESTING = FALSE

```

## Get BTC data

Obtaining BTC price data (daily) for the last 200 days.

```{r,btc1}

datebefore <- Sys.Date()-200
mydate <- Sys.Date()
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=",datebefore,"&time_end=",Sys.Date(),sep="")
download.file(URL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Now gather houly price data for the last 23 hours.

```{r,btc2}

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=hourly&time_period=hourly&time_start=",Sys.Date()-1,"&time_end=",Sys.Date(),sep="")
download.file(URL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
price2 <- btcdat$data$quotes
price2 <- data.frame(price2$time_close, price2$quote$USD$high, price2$quote$USD$low, price2$quote$USD$close,stringsAsFactors=FALSE)
colnames(price2) <- c("date","high","low","close")
price2$date <- sapply(strsplit(as.character(price2$date),"T"),"[[",1)
tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)

high <- max(price2[,2])
low <- min(price2[,3])
close <- price2[nrow(price2),4]

df <- data.frame(date=Sys.Date(),high=high,low=low,close=close,stringsAsFactors=FALSE)
price <- rbind(price,df)
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

if ( TESTING == TRUE) {
  price[nrow(price),"Close"] <- price[nrow(price),"Close"] * 2
}

```

## BTC on the 43 MA

Now to determine whether BTC has crossed the 43-day SMA.

```{r, btc_sma}

price$ma <- SMA(Cl(price),n=43)
price$signal <- price$close > price$ma
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)
today_signal <- price[nrow(price),"signal"]
previous_signal <- price[(nrow(price)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    btc_signal="BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) { 
    btc_signal="SELL"
  }
} else {
btc_signal <- "NONE"
}

message(paste("BTC SMA signal:",btc_signal))

```

```{r, btc_sma_plot,fig.height=8,fig.width=8}

plot(price$close~as.Date(price$date),type="l",
  xlab="Date",ylab="price (USD)",main="BTC w 43D MA")
grid()
lines(price$ma~ as.Date(price$date) ,col="red")

```

Send a push notification.

```{r,btcpush}

if ( btc_signal != "NONE") {
  MYNOTE <- paste("The BTC 43D MA signal is", btc_signal, 
    ". Visit https://blinder.cc/crypto/alerts.html for the details")
  pbPost("note", "Crypto Alert", MYNOTE)
}

```

## DMI BTC

Directional movement indicator is a highly profitable approach to identify trend changes.
DMI using a 26 day interval gives returns 4x better than the 43D SMA approach.

```{r,dmi_btc,fig.height=10,fig.width=8}

dmi.adx <- ADX(price[,c("high","low","close")],n=26)
price <- cbind(price,dmi.adx)
price$signal <- dmi.adx[,1] > dmi.adx[,2]
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

today_signal <- price[nrow(price),"signal"]
previous_signal <- price[(nrow(price)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    btc_signal="DMI BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    btc_signal="DMI SELL"
  }
} else {
  btc_signal <- "NONE"
}

message(paste("BTC DMI signal:",btc_signal))

```

```{r,dmi_btc_plot,fig.height=10,fig.width=8}

price <- as.data.frame(price)
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",
  xlab="Date",ylab="price (USD)",main="BTC USD price")
grid()
plot(price$DIp~as.Date(price$date),type="l", col="blue",
  xlab="Date",ylab="price (USD)",main="BTC 26d DMI")
grid()
lines(price$DIn ~ as.Date(price$date) ,col="red")

```

Send a push notification.

```{r,btcpush2}

if ( btc_signal != "NONE") {
  MYNOTE <- paste("BTC is flashing a ", btc_signal,
   ". Visit https://blinder.cc/crypto/alerts.html for the details")
  pbPost("note", "Crypto Alert", MYNOTE)
}

```

## Get ETH data

Obtaining ETH historical data (daily).

```{r,eth1}

datebefore <- Sys.Date()-200
mydate <- Sys.Date()
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=",datebefore,"&time_end=",Sys.Date(),sep="")
download.file(URL,destfile="ethdat.txt")
ethdat <- fromJSON("ethdat.txt")
price <- ethdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Now gather houly price data for the last 23 hours.

```{r,eth2}

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=hourly&time_period=hourly&time_start=",Sys.Date()-1,"&time_end=",Sys.Date(),sep="")
download.file(URL,destfile="ethdat.txt")
ethdat <- fromJSON("ethdat.txt")
price2 <- ethdat$data$quotes
price2 <- data.frame(price2$time_close, price2$quote$USD$high, price2$quote$USD$low, price2$quote$USD$close,stringsAsFactors=FALSE)
colnames(price2) <- c("date","high","low","close")
price2$date <- sapply(strsplit(as.character(price2$date),"T"),"[[",1)
tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)

high <- max(price2[,2])
low <- min(price2[,3])
close <- price2[nrow(price2),4]

df <- data.frame(date=Sys.Date(),high=high,low=low,close=close,stringsAsFactors=FALSE)
price <- rbind(price,df)
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

if ( TESTING == TRUE) {
  price[nrow(price),"Close"] <- price[nrow(price),"Close"] * 2
}

```

## ETH on the 34 MA

Now to determine whether ETH has crossed the 34-day SMA.

```{r, eth_sma}

price$ma <- SMA(Cl(price),n=34)
price$signal <- price$close > price$ma
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)
today_signal <- price[nrow(price),"signal"]
previous_signal <- price[(nrow(price)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    eth_signal="BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    eth_signal="SELL"
  }
} else {
  eth_signal <- "NONE"
}

message(paste("ETH SMA signal:",eth_signal))

```

```{r, eth_sma_plot,fig.height=8,fig.width=8}

plot(price$close~as.Date(price$date),type="l",
  xlab="Date",ylab="price (USD)",main="ETH w 34D MA")
grid()
lines(price$ma~ as.Date(price$date) ,col="red")

```

Send a push notification.

```{r,ethpush}

if ( eth_signal != "NONE") {
  MYNOTE <- paste("The ETH 34D MA signal is", eth_signal,
   ". Visit https://blinder.cc/crypto/alerts.html for the details")
  pbPost("note", "Crypto Alert", MYNOTE)
}

```

## DMI ETH

For ETH, the 19 day interval DMI gave a good return.

```{r,dmi_eth,fig.height=10,fig.width=8}

dmi.adx <- ADX(price[,c("high","low","close")],n=19)
price <- cbind(price,dmi.adx)
price$signal <- dmi.adx[,1] > dmi.adx[,2]
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

today_signal <- price[nrow(price),"signal"]
previous_signal <- price[(nrow(price)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    eth_signal="DMI BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    eth_signal="DMI SELL"
  }
} else {
  eth_signal <- "NONE"
}

message(paste("ETH SMA signal:",eth_signal))

```

```{r, eth_dmi_plot,fig.height=10,fig.width=8}

price <- as.data.frame(price)
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",
  xlab="Date",ylab="price (USD)",main="ETH USD price")
grid()
plot(price$DIp~as.Date(price$date),type="l", col="blue",
  xlab="Date",ylab="price (USD)",main="ETH 19d DMI")
grid()
lines(price$DIn ~ as.Date(price$date) ,col="red")

```

Send a push notification.

```{r,ethpush2}

if ( eth_signal != "NONE") {
  MYNOTE <- paste("ETH is flashing a ", eth_signal,
   ". Visit https://blinder.cc/crypto/alerts.html for the details")
  pbPost("note", "Crypto Alert", MYNOTE)
}

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
