---
title: "Trading BTC based on CCI"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/cci_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: bc1qgsecwc3e7r0g7zch8wum04a8hk74tfy488m2nv

XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

In this blog post I will be trilaing CCI commodity channel index for entry and exits for BTC.

Traditionally traders used this to identify buy signals when the index goes above the -100 level and exit when it goes below the +100 level.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("parallel")
  library("kableExtra")
})

knitr::opts_chunk$set(fig.width=8, fig.height=8) 

```

Reminder: this is not financial advice.

## Get data

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$high,price$quote$USD$low,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)

```

## Calculate hodl returns

It's important to calculate the returns for simply holding the asset rather than actively trading.

The number shown here is the fold return on investment.
BTC has appreciated a lot since 2013!

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## CCI chart analysis

The textbook application is to enter after CCI goes above -100 and exit when CCI goes below +100

```{r,cci1}

n=20
c=0.015
cci <- CCI(price[,c("high","low","close")], n = 20, c = 0.015 )

par(mfrow=c(2,1))
plot(price$close~ as.Date(price$date),type="l",log="y",xlab="",ylab="USD price")
grid()
plot(cci ~ as.Date(price$date),type="l",xlab="",ylab="CCI")
grid()
abline(h=-100,col="red")
abline(h=+100,col="red")

```

Here is the function for CCI, where we'll change the entry and exit level.

```{r,cci2}

ccifun <- function(price,entry=-100,exit=100,n=20,fee=0.001) {
  cci <- CCI(price[,c("high","low","close")], n = n, c = 0.015 )
  cciy <- c(NA,cci)[1:nrow(price)]
  price2 <- data.frame(price,cci,cciy,stringsAsFactors=FALSE)
  price2$sell <- as.numeric( price2$cci < exit & price2$cciy > exit )
  price2$buy <- as.numeric( price2$cci > entry & price2$cciy < entry )
  price2 <- subset(price2, buy == 1 | sell == 1)
  price2$y <- c(NA,price2$buy)[1:nrow(price2)]
  price2 <- price2[which(price2$buy != price2$y),]

  if (nrow(price2)>0) {
    buy <- subset(price2,buy==1)[c("date","close")]
    colnames(buy) <- c("buy_date","buy_price")
    sell <- subset(price2,sell==1)[c("date","close")]
    colnames(sell) <- c("sell_date","sell_price")
    if( nrow(buy) != nrow(sell) ) {
      MIN=min(nrow(buy),nrow(sell))
      buy <- buy[1:MIN,]
      sell <- sell[1:MIN,]
    }
    trades <- cbind(buy,sell)
    trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
    return(c(prod(trades$diff),nrow(trades)))
  } else {
    return(c(1,0))
  }
}


```

## Backtesting

Now let's go back in time to test the performance of the DMI indicator using different
parameters:

* entry at -200:0

* exit at 0:+200

* n=10:20

Let's assume a 0.1% trading fee, as per binance current rate.

Here is a table of results.

```{r,backtest1}

fee=0.001
n=20

ccifun(price, entry=-100,  exit=100, n=20)

mymx <- expand.grid(seq(-100,0,5),seq(0,100,5),seq(5,50,3))

mylist <- as.list(as.data.frame(t(mymx)))

cci_res <- mclapply(X=mylist,mc.cores=4,FUN=function(x) { 
  ccifun(price,entry=x[1],exit=x[2],n=x[3],fee=0.001)
})

cci_res <- do.call(rbind,cci_res)
cci_res <- cbind(mymx,cci_res)
colnames(cci_res) <- c("entry","exit","n","returns","trades")
cci_res2$mean <- cci_res2$returns^(1/cci_res2$trades)

head(cci_res[order(-cci_res[,"returns"]),],20) %>%
  kbl(caption="Highest returns CCI strategies") %>%
  kable_styling()

```

Now do a "fine" parameter sweep in a narrow range.

```{r,backtest2}

mymx <- expand.grid(seq(-20,10,1),seq(0,25,1),5:50)

mylist <- as.list(as.data.frame(t(mymx)))

cci_res2 <- mclapply(X=mylist,mc.cores=2,FUN=function(x) {
  ccifun(price,entry=x[1],exit=x[2],n=x[3],fee=0.001)
})

cci_res2 <- do.call(rbind,cci_res2)
cci_res2 <- cbind(mymx,cci_res2)
colnames(cci_res2) <- c("entry","exit","n","returns","trades")
cci_res2$mean <- cci_res2$returns^(1/cci_res2$trades)

head(cci_res2[order(-cci_res2[,"returns"]),],20) %>% 
  kbl(caption="Highest returns CCI strategies") %>%
  kable_styling()

cci_res3 <- subset(cci_res2, trades < 200 )
head(cci_res3[order(-cci_res3[,"returns"]),],20) %>%
  kbl(caption="Highest returns CCI strategies with 200 trades or less") %>%
  kable_styling()

head(cci_res3[order(-cci_res3[,"mean"]),],20) %>%
  kbl(caption="Highest mean ROI CCI strategies") %>%
  kable_styling()

```

## Optimised strategy

Entry: 1

Exit: 10

n=19

```{r,opti1}

n=19
entry=1
exit=10

cci <- CCI(price[,c("high","low","close")], n = n, c = 0.015 )
cciy <- c(NA,cci)[1:nrow(price)]
price2 <- data.frame(price,cci,cciy,stringsAsFactors=FALSE)
price2$sell <- as.numeric( price2$cci < exit & price2$cciy > exit )
price2$buy <- as.numeric( price2$cci > entry & price2$cciy < entry )
price2 <- subset(price2, buy == 1 | sell == 1)
price2$y <- c(NA,price2$buy)[1:nrow(price2)]
price2 <- price2[which(price2$buy != price2$y),]

buy <- subset(price2,buy==1)[c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,sell==1)[c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

trades %>%
  kbl(caption="Trades for the optimised strategy") %>%
  kable_styling()

```

## Effect of fees

It is important to factor in the effect of fees because higher exchange fees will erode profits.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_fees <- sapply(fee_levels,function(x) { ccifun(price, entry=1,  exit=10, n=19, fee=x) } )
plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

It is good to see whether the strategy gives consistent returns over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

trades$year <- sapply(strsplit(trades[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades,year==y)[,5] } )
yearly <- yearly[which(lapply(yearly,length)>0)]
vioplot(yearly)
yr <- sapply(yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```
## Conclusion

This could potentially be an interesting strategy.
There were sure a lot of false signals.
And the returns are still a bit lower than the TSI approach while still better than the DMI approach.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
