---
title: "Using correlations to predict market direction"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/tsi_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: bc1qgsecwc3e7r0g7zch8wum04a8hk74tfy488m2nv

XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

In this analysis, I will be seeing whether correlation analysis of recent and historical price action
can predict the price in the near future.
I will be using the daily timeframe for approx 100 days of recent action and comparing this profile with
historic price action.
Each historic profile will be associated with a price outcome.
These price outcomes can be quantified by the sum of the price outcomes * correlation coefficient.

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library("parallel")
})

Sys.Date()

```

Reminder: this is not financial advice.

## Get data

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)
price0 <- price

```

## Calculate hodl returns

It's important to calculate the returns for simply holding the asset rather than actively trading.

The number shown here is the fold return on investment.
BTC has appreciated a lot since 2013!

```{r,hodl}

myend <- price[c(1,nrow(price)),4][2] 
mystart <- price[c(1,nrow(price)),4][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Break the data into chunks

nd = num days of correlation data to use

fd = num days for future price prediction

nc = num cors to include

nsims = number of historical simulations

```{r,chunks,fig.height=10,fig.width=8}

nd=100
fd=100
nc=100

recent <- tail(price,nd)
nr<-nrow(price)
price2 <- price0[1:(nr-round(nd/2)),]
chunks <- lapply( 1:(nrow(price2)-fd-nd) , function(x) { price2[x:(x+fd+nd),] } )
names(chunks) <- 1:length(chunks)
profiles <- lapply(chunks , function(x) { x[1:nd,"close"]  } )
results <- lapply(chunks , function(x) { x[nd+fd,"close"]/x[nd,"close"] } )
cors <- lapply( profiles, function(x) {cor(x,recent$close)} )
cors <- unlist(cors)
cors <- cors[order(cors)]
cors <- tail(cors,100)
results <- results[match(names(cors) ,names(results))]
results <- unlist(results)
prod <- results * cors
boxplot(prod)
abline(h=1)
mean(prod)
median(prod)

```

## Backtest it for forecast accuracy

This is working now but unable to predict anything useful.

```{r,backtest,fig.height=10,fig.width=8}

ndrange <- c(10,20,50,100,150,200)
fdrange <- c(7,14,21,28)
ncrange <- c(5,10,20,50)

pricecor <- function(nd,fd,nc) {

nsims=1000

mysims <- head(sample(x=1:(nrow(price)-nd-fd)),nsims)
mysims <- mysims[which(mysims > (nd+fd+(round(nd/2))))]

res <- mclapply(mysims, function(y) {

  price0 <- head(price,y)
  recent <- tail(price0,(nd+fd))

  end <- tail(recent,1)$close
  recent <- head(recent,nd)
  current <- tail(recent,1)$close  
  actual <- end / current

  nr<-nrow(price0)
  price2 <- price0[1:(nr-round(nd/2)),]
  chunks <- lapply( 1:(nrow(price2)-fd-nd) , function(x) { price2[x:(x+fd+nd),] } )
  names(chunks) <- 1:length(chunks)
  profiles <- lapply(chunks , function(x) { x[1:nd,"close"]  } )
  results <- lapply(chunks , function(x) { x[nd+fd,"close"]/x[nd,"close"] } )
  cors <- lapply( profiles, function(x) {cor(x,recent$close)} )
  cors <- unlist(cors)
  cors <- cors[order(cors)]
  cors <- tail(cors,nc)
  results <- results[match(names(cors) ,names(results))]
  results <- unlist(results)
  prod <- results * cors
  return( c("pred"=median(prod),"act"=actual) )
},mc.cores=8)

res <- do.call(rbind,res)
res <- as.data.frame(res)

ret <- ( nrow(subset(res,pred>1 & act > 1)) + nrow(subset(res,pred < 1 & act < 1)) ) / nrow(res)

return( c( nd, fd, nc, ret ) )

}

params <- expand.grid(ndrange,fdrange,ncrange)

bt <- mclapply( 1:nrow(params), function(i) {
  nd <- params[i,1]
  fd <- params[i,2]
  nc <- params[i,3]
  pricecor( nd, fd, nc )
} , mc.cores=8 )


```

Here are the strategies which had the highest mean return per trade.
These opportunities don't arise often but could be potential opportunities for leveraged trades.

```{r,backtest_res1}

head(res2[order(-res2$mean),],20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Here are the strategies which had the highest overall returns.
This involves more frequent trading and the returns on each trade were smaller.
This is less optimal if trades are executed manually.

```{r,backtest_res2}

head(res2[order(-res2$returns),],20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Now look at the strategies with mean ROI>1.1 and highest overall returns.

```{r,backtest,res3}

res3 <- subset(res2,mean>=1.1)
res3 <- res3[order(-res3$returns),]
head(res3,20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```


## Plots

Another approach to selecting "good" parameters could be to begin with the default values and then
aim to increase returns and decrease the number of trades executed.

The red dot in the chart below indicates the TSI with default interval values.
Data points to the upper right side of that dot are strategies with better performance.

```{r,plots1,fig.height=8,fig.width=8}

plot(res2$mean , res2$returns, log="y",xlab="mean ROI", ylab="total returns")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()

plot(res3$mean , res3$returns, log="y",xlab="mean ROI", ylab="total returns")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()

```

## Backtest part 2

Now a fine search of the best parameter space.

```{r,backtest2}

fee=0.001

n1r <- seq(min(head(res3,20)$n1),max(head(res3,20)$n1),1)
n2r <- seq(min(head(res3,20)$n2),max(head(res3,20)$n2),1)
nsr <- seq(min(head(res3,20)$ns),max(head(res3,20)$ns),1)
params <- expand.grid(n1r,n2r,nsr)
colnames(params) <- c("n1","n2","nsignal")
params <- params[which( params$n1 > params$n2) ,]

res <- mclapply(X=1:nrow(params) , FUN=function(i) {
  n1=as.numeric(params[i,1])
  n2=as.numeric(params[i,2])
  ns=as.numeric(params[i,3])
  mytsi(price,n1=n1,n2=n2,ns=ns,fee=fee)
} , mc.cores = detectCores())
res <- as.data.frame(do.call(rbind, res))
colnames(res) <- c("returns","trades")
res2 <- cbind(params,res)
res2$returnspertrade <- res2$returns/res2$trades
res2$mean <- res2$returns^(1/res2$trades)

basic <- mytsi(price,n1=25,n2=13,ns=7,fee=fee)
basic_returns <- basic[1]
basic_mean <- basic[1]^(1/basic[2])

```

Here are the strategies which had the highest mean return per trade.
These opportunities don't arise often but could be potential opportunities for leveraged trades.

```{r,backtest_res4}

head(res2[order(-res2$mean),],20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Here are the strategies which had the highest overall returns.
This involves more frequent trading and the returns on each trade were smaller.
This is less optimal if trades are executed manually.

```{r,backtest_res5}

head(res2[order(-res2$returns),],20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Now look at the strategies with mean ROI>1.1 and highest overall returns.

```{r,backtest,res6}

res3 <- subset(res2,mean>=1.1)
res3 <- res3[order(-res3$returns),]
head(res3,20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

## Optimised strategy

Here I use the optimised values for n1, n2 and n.signal and we can have a look at all the trades 
executed over the years.

```{r,best1,fig.width=10,fig.height=8}

best_n1 = res3[1,1]
best_n2 = res3[1,2]
best_ns = res3[1,3]

message(paste("Selected n1 period:",best_n1))
message(paste("Selected n2 period:",best_n2))
message(paste("Selected n.signal period:",best_ns))

tsi <- TSI(price$close , n.first = best_n1, n.second = best_n2, n.signal = best_ns )
price2 <- price
price2 <- cbind(price2,tsi)
price2$higher <- price2$tsi > price2$signal
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

trades %>% kbl() %>%  kable_styling()

HEADER="TSI"
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tsi[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="TSI",main="TSI")
lines(as.Date(price$date), tsi[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Effect of fees

It is important to factor in the effect of fees because higher exchange fees will erode profits.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_fees <- sapply(fee_levels,function(x) {
  mytsi(price,n1=best_n1,n2=best_n2,ns=best_ns,fee=x)
} )

plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
abline(h=hodl,col="red")
grid()

```

## Trades by year

It is good to see whether the strategy gives consistent returns over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

trades$year <- sapply(strsplit(trades[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades,year==y)[,5] } )
yearly <- yearly[which(lapply(yearly,length)>0)]
vioplot(yearly)

yr <- sapply(yearly,prod)
barplot(yr,ylab="fold return per year")
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```

## Conclusion

There are many approaches to momentum trading.
Here using the TSI we have a strategy which scored 9x hodl returns which is superior to the best directional movement indicator
and best moving average strategy (3.4x hodl) with roughly the same number of trades.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
