---
title: "<br>Trading based on day of week"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

```

## Get data

Obtaining BTC historical data (daily).

Monday=1, Sunday=7

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

# day of the week
price$day <- c(7,1:6)
```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Day of the week

Monday=1

Sunday=7

```{r,sma}

b=7
s=5

dow <- function(price,b,s,fee) {
  ss <- subset(price,day==b | day==s)
  if (ss[1,3] == s) { ss <- ss[2:nrow(ss),] }
  buy <- subset(ss,day==b)[,2]
  sell <- subset(ss,day==s)[,2]
  MIN <- min( length(buy),length(sell))
  buy <- buy[1:MIN]
  sell <- sell[1:MIN]
  trades <- data.frame(buy,sell)
  trades$diff <- ( ( trades$sell - trades$buy ) / trades$buy ) +1 - ( 2 * fee )
  return(c(b,s,prod(trades$diff),nrow(trades)))
}

params <- expand.grid(1:7,1:7)
colnames(params) <- c("buy","sell")
params <- params[which(params$buy != params$sell),]

res <- t(sapply(1:nrow(params),function(i) {
  b <- params[i,1]
  s <- params[i,2]
  dow(price,b,s,fee) 
} ))

colnames(res) <- c("buy","sell","ROI","trades")
res

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
