---
title: "Trading based on Directional Movement Index (DMI)"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

```

## Get data

Obtaining BTC historical data (weekly).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=weekly&time_start=2013-06-01&time_end=",mydate,sep="")

if ( ! file.exists("btcdat_weekly.txt") ) {

  download.file(URL,destfile="btcdat_weekly.txt")

}

btcdat <- fromJSON("btcdat_weekly.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),4][2] 
mystart <- price[c(1,nrow(price)),4][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Basic look at the indicator

Default n = 14.

```{r,dmi1,fig.height=10,fig.width=8}

par(mfrow=c(2,1))
price2 <- tail(price,730)
dmi.adx <- ADX(price2[,c("high","low","close")])
par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(dmi.adx[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="DMI",main="DMI",ylim=c(0,100))
lines(as.Date(price2$date), dmi.adx[,2]  , col="red"  )
grid()

price2 <- tail(price,150)
dmi.adx <- ADX(price2[,c("high","low","close")],n=3)
par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(dmi.adx[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="DMI",main="DMI",ylim=c(0,100))
lines(as.Date(price2$date), dmi.adx[,2]  , col="red"  )
grid()

par(mfrow=c(1,1))


```

## DMI trading function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}


dmi <- function(price,n,fee) {
  dmi.adx <- ADX(price[,c("high","low","close")],n=n)
  price2 <- price
  price2 <- cbind(price2,dmi.adx)
  price2$higher <- price2$DIp > price2$DIn
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(c(prod(trades$diff),nrow(trades)))
}

```


## Backtesting

```{r,backtest}

fee=0.002

myrange <- 3:200
res_dmi <- t(sapply(myrange,function(n) { dmi(price,n,fee) } ))
rownames(res_dmi) <- myrange
colnames(res_dmi) <- c("returns","n trades")

```

## Plots

```{r,plots,fig.height=8,fig.width=8}

plot(res_dmi[,2],res_dmi[,1], cex=2, pch=19,col="gray",
  xlab="num trades",ylab="fold returns")
text(res_dmi[,2],res_dmi[,1],labels=rownames(res_dmi),cex=1.5)
abline(h=hodl,col="red")
grid()

res_dmi <- c(0,0,res_dmi[,1])

plot(res_dmi,pch=19,xlab="n interval (weeks)",ylab="fold return",main="DMI")
abline(h=hodl,col="red")
grid()

plot(res_dmi+1,pch=19,xlim=c(1,20),xlab="n interval (weeks)",ylab="fold return",main="DMI")
abline(h=hodl,col="red")
grid()

```

## Optimisation

```{r,best_sma,fig.width=10,fig.height=8}

best_n = 6
best_n

dmi.adx <- ADX(price[,c("high","low","close")],n=best_n)
price2 <- price
price2 <- cbind(price2,dmi.adx)
price2$higher <- price2$DIp > price2$DIn
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
nrow(trades)
trades
prod(trades$diff)


HEADER=paste(best_n,"DMI")
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(dmi.adx[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="DMI",main="DMI",ylim=c(0,100))
lines(as.Date(price$date), dmi.adx[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))


```

## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_fees <- sapply(fee_levels,function(x) { dmi(price,best_n,fee=x) } )
plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(HEADER)
abline(h=hodl,col="red")
grid()


```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

trades$year <- sapply(strsplit(trades[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades,year==y)[,5] } )
yearly <- yearly[which(lapply(yearly,length)>0)]
vioplot(yearly)
barplot(sapply(yearly,prod))
abline(h=1,col="red",lty=2)

```




## Conclusion

Hodling BTC over this period gave a 378 fold return.
Here we are looking at using the DMI as a buy/sell indicator.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
