---
title: "Trading BTC based on Directional Movement Index (DMI)"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/dmi_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

There are many indicators of market momentum, one of them is the directional movement 
indicator (DMI). 
For this post, I'll be using a simple implementation of the DMI, by buying when the positive
momentum is greater than the negative momentum indicator. 
Then sell when the negative indicator is larger than the positive indicator.
What I am examining here is different intervals for the calculation of the DMI, typically 
14 day is used, however I will be seeing whether other intervals perform better.

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
})

Sys.Date()

```

Reminder: this is not financial advice.

## Get data

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

It's important to calculate the returns for simply holding the asset rather than actively trading.

The number shown here is the fold return on investment.
BTC has appreciated a lot since 2013!

```{r,hodl}

myend <- price[c(1,nrow(price)),4][2] 
mystart <- price[c(1,nrow(price)),4][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Basic look at the DMI indicator

Here we can see the DMI over the past 730 days, using the default interval of 14d.
The default moving average algorithm is the Welles Wilder EMA.

```{r,dmi1,fig.height=10,fig.width=8}

par(mfrow=c(2,1))
price2 <- tail(price,730)
dmi.adx <- ADX(price2[,c("high","low","close")])
par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(dmi.adx[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="DMI",main="DMI",ylim=c(0,60))
lines(as.Date(price2$date), dmi.adx[,2]  , col="red"  )
grid()

par(mfrow=c(1,1))

```

## DMI trading function

Here is the function for trading based on DMI crosses, buying when positive momentum indicator 
goes above the negative momentum indicator line, and sell when it falls below.

All trades are executed at the end of the daily session (UTC midnight).

```{r,dmi_func}

dmi <- function(price,n,fee) {
  dmi.adx <- ADX(price[,c("high","low","close")],n=n)
  price2 <- price
  price2 <- cbind(price2,dmi.adx)
  price2$higher <- price2$DIp > price2$DIn
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(c(prod(trades$diff),nrow(trades)))
}

```

## Backtesting

Now let's go back in time to test the performance of the DMI indicator using different 
intervals from 4 days to 200 days.

Let's assume a 0.1% trading fee, as per binance current rate.

Here is a table of results.

```{r,backtest}

fee=0.001
myrange <- 4:200
res <- t(sapply(myrange,function(n) { dmi(price,n,fee) } ))
rownames(res) <- myrange
colnames(res) <- c("returns","trades")
res <-  as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res <- cbind(rownames(res),res,stringsAsFactors=FALSE)
colnames(res)[1] <- "period"
res %>% kbl() %>%  kable_paper("hover", full_width = F)

```

## Plots

Let's examine some of the trends in the data with some charts.

```{r,plots,fig.height=8,fig.width=8}

plot(res$trades,res$returns,log="y", cex=2, pch=19,col="gray",
  xlab="num trades",ylab="fold returns")
text(res$trades,res$returns,labels=res$period,cex=1.5)
abline(h=hodl,col="red")
grid()

plot(res$period,res$returns,pch=19,xlab="n interval (days)",ylab="fold return",main="DMI")
abline(h=hodl,col="red")
grid()

plot(res$period,res$returns,pch=19,xlim=c(1,50),xlab="n interval (days)",ylab="fold return",
  main="DMI",log="y")
abline(h=hodl,col="red")
grid()

plot(res$mean,res$returns,xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res$mean,res$returns , cex=1, labels=res$period)
abline(h=hodl,col="red")

plot(res$mean,res$returns,log="y",xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res$mean,res$returns , cex=1, labels=res$period)
abline(h=hodl,col="red")

plot(res$mean,res$returns,log="y",xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray",
  ylim=c(100,1E4))
grid()
text(res$mean,res$returns , cex=1, labels=res$period)
abline(h=hodl,col="red")

plot(res[,3],res[,2],pch=19,cex=2,col="gray",main="DMI returns per trade",xlab="n trades",ylab="fold return USD")
text(res[,3],res[,2],labels=res$period,pch=19,cex=1)
abline(h=hodl,col="red")


```

## 129 day DMI

The 129 day interval gives a high probability of high returns but occurs rarely.

Below is a table of the trades made.

```{r,best_sma1,fig.width=10,fig.height=8}

best_n = 129
message(paste("Selected DMI period:",best_n))

dmi.adx <- ADX(price[,c("high","low","close")],n=best_n)
price2 <- price
price2 <- cbind(price2,dmi.adx)
price2$higher <- price2$DIp > price2$DIn
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

trades %>% kbl() %>%  kable_styling()

HEADER=paste(best_n,"DMI")
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(dmi.adx[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="DMI",main="DMI 129d",ylim=c(0,60))
lines(as.Date(price$date), dmi.adx[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## 8 day DMI

This indicator gives a very high return, but trades are executed very frequently and the mean ROI is less than 1.1.
This requires a lot of work for manual execution, and may not be sustainable.

```{r,best_sma3,fig.width=10,fig.height=8}

best_n = 8
message(paste("Selected DMI period:",best_n))

dmi.adx <- ADX(price[,c("high","low","close")],n=best_n)
price2 <- price
price2 <- cbind(price2,dmi.adx)
price2$higher <- price2$DIp > price2$DIn
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

trades %>% kbl() %>%  kable_styling()

HEADER=paste(best_n,"DMI")
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(dmi.adx[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="DMI 8d",main="DMI",ylim=c(0,60))
lines(as.Date(price$date), dmi.adx[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## 32 day DMI

This is good indicator because the frequency is lower, but the gains are still very good.

```{r,best_sma2,fig.width=10,fig.height=8}

best_n = 32
message(paste("Selected DMI period:",best_n))

dmi.adx <- ADX(price[,c("high","low","close")],n=best_n)
price2 <- price
price2 <- cbind(price2,dmi.adx)
price2$higher <- price2$DIp > price2$DIn
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

trades %>% kbl() %>%  kable_styling()

HEADER=paste(best_n,"DMI")
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(dmi.adx[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="DMI",main="DMI 32d",ylim=c(0,60))
lines(as.Date(price$date), dmi.adx[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Effect of fees

It is important to factor in the effect of fees because higher exchange fees will erode profits.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_fees <- sapply(fee_levels,function(x) { dmi(price,best_n,fee=x) } )
plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(HEADER)
abline(h=hodl,col="red")
grid()


```

## Trades by year

It is good to see whether the strategy gives consistent returns over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

trades$year <- sapply(strsplit(trades[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades,year==y)[,5] } )
vioplot(yearly)
yr <- sapply(yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```

## Conclusion

For the DMI, adjusting the interval length reveals a trade-off.
Shorter timeframes give high returns, however require frequent trades, every few days.
This may not be sustainable in the long term if trades are executed manually.
Longer timeframes give a lower return overall, but the signals have a lower false positive rate.
The 8-day DMI gave returns of 9.8x the hodl ROI, which is significantly better than the moving average 
strategies which yielded returns around 3-6x the hodl ROI.

If you would like to receive trade alerts on these strategies, let me know by email mark.ziemann{αt}gmail.com.
These alerts are sent at 30 mins before the close of the daily session and will assist you in sticking to your desired strategy.
Subscribing to alerts costs 0.0005 BTC per year.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
