---
title: "<br>Trading ETH based on an ensemble of indicators"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
---

## Intro

In this report I will be looking at whether a combination of three indicators works better than
each on their own. 
The theory is that there is a chance that combining them will reduce false positives.

The three indicators I will be using are:

1. 34 day simple moving average (35x hodl)

2. Money flow indicator with a signal line; n1=17d, n2=250d (0.78x hodl).
The returns are probably closer to 2.2x hodl as the indicator only began signaling at the end of 2016.

3. RSIon the 21d period.
Enter when RSI goes above 50 and exit when RSI dips below 50.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("parallel")
  library("kableExtra")
})

knitr::opts_chunk$set(fig.width=8, fig.height=8) 

```

## Get data

Obtaining ETH historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="ethdat.txt")

dat <- fromJSON("ethdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,
  price$quote$USD$volume, stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close","volume")
price$date <- as.Date(sapply(strsplit(as.character(price$date),"T"),"[[",1))
#price <- subset(price,volume>0)

```

## Calculate hodl returns

This is the benchmark we use for judging the performance of an indicator. 
Any strategy with total returns under this value is considered bad.

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## MFI analysis

Money flow indicator with a signal line; n1=17d, n2=250d (0.78x hodl).
See the MFI backtesting report to understand how these values were decided.

```{r,mfi}

n1_best = 17
n2_best = 250

price2 <- price
mfi1 <- EMA(MFI(HLC=price[,2:4],volume=price$volume,n=n1_best),n=n1_best)
mfi2 <- EMA(MFI(HLC=price[,2:4],volume=price$volume,n=n2_best),n=n2_best)
price2$mfi <- mfi1 > mfi2

```

## SMA analysis

The 34 day SMA is optimal for ETH, as we know from the mdz-analytics.com website.
It performs slightly better than the optimal EMA, or MA crosses.

```{r,sma}

best_sma = 34
ma <- SMA(Cl(price),n=best_sma)
price2$sma <- price2$close > ma

```

## RSI analysis

The idea here is to enter when  RSI recovers the 50 level.

Previously we showed that a period of 21 days gave good performance with an entry at 49 and exit at 54.

```{r,rsi}

n_rsi <- 21

rsi <- RSI(Cl(price),n=n_rsi)

price2$rsi <- rsi>50

```

## Ensemble trading

Now that the indicators have each been calculated, we can open a trade when X indicators are positive and sell when only Y are positive.

In order to decide on X and Y, we should look at the data first.

```{r,ensembl1}

price2[is.na(price2)] <- 0
price2$sum <- price2$mfi + price2$sma + price2$rsi

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$sum ~ as.Date(price2$date),type="l",
  xlab="Date",ylab="positive indicators",main="Indicator set")
grid()

par(mfrow=c(1,1))

plot(price2$close ~ as.Date(price2$date),type="p",log="y", pch=19,cex=0.5,
  col=price2$sum+1,
  xlab="Date",ylab="price (USD)",main="USD price coloured by num indicators")
  mtext("black=0,red=1,green=2,blue=3")
grid()

```

This is looking good. 
Note that for the period in 2016, the MFI could not be calculated, so the rally is depicted in green.

```{r,backtest1}

ensembl <- function(price2,threshold,fee) {
  price2$higher <- price2$sum >= threshold
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(c(prod(trades$diff),nrow(trades)))
}

res <- t( sapply(1:3,function(i) {
  ensembl(price2,threshold=i,fee=0.002)
}) )

rownames(res) <- c(1,2,3)
xhodl <- t(t(as.numeric(res[, 1]/hodl)))
res <- cbind(res,xhodl)
colnames(res) <- c("totalROI","n_trades","xhodl")
res %>% kbl(caption="backtesting results with minimum of 1,2,3 postive indicators") %>%  kable_paper("hover", full_width = F)

```

## Conclusion

Results were excellent for the ensembl with a minimum threshold of 2 positive indicators.
The reason why the threshold of 3 was so low could be due to MFI indicator not being calculated for the 2016 period.
Only the ETH Stoch indicator performed better, with 105x hodl.

These indicators can be added to your TradingView, with notifications.
Alternatively, I can set up pushbullet notifications for the ensembl-based entry and exit signals.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
