---
title: "Trading BTC based on MA and EMAs"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/ma_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

The Heikin Ashi candlestick has an alternate formula and yields charts where  the momentum is easier to read.

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library(anytime)
})


Rcpp::cppFunction('NumericMatrix RawHeikinAshi(NumericMatrix x, CharacterVector n) {
// assumes OHLC matrix input
int nrow = x.nrow(), ncol = 4, Op=0, Hi=1, Lo=2, Cl=3;
NumericMatrix ha(nrow,ncol);
for (int i = 0; i < nrow; i++) {
    ha(i, Cl) = (x(i,Op) + x(i,Hi) + x(i,Lo) + x(i,Cl)) / 4.0;
    ha(i, Op) = (i > 0) ? ((ha(i - 1, Op) + ha(i - 1, Cl)) / 2.0) : x(i, Op);
    ha(i, Hi) = std::max(x(i, Hi), std::max(ha(i, Op), ha(i, Cl)));
    ha(i, Lo) = std::min(x(i, Lo), std::min(ha(i, Op), ha(i, Cl)));
}
colnames(ha) = n;
return ha;
}')

HAOHLC <- function(x) {
    x <- OHLC(try.xts(x))
    r <- RawHeikinAshi(x, paste0("ha.", colnames(x)))
    return(reclass(r, x))
}

Sys.Date()

```

Reminder: this is not financial advice.

## Get data

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$quote$USD$open , price$quote$USD$high, price$quote$USD$low, price$quote$USD$close, price$quote$USD$timestamp )
colnames(price) <- c("open","high","low","close","date")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

rownames(price) <- sapply(strsplit(as.character(price$date),"T"),"[[",1)
price$date <- NULL

# delete 1 day to start on Monday
weekdays(as.Date(rownames(head(price))[1]))
first_monday <- which(weekdays(as.Date(rownames(price)))=="Monday")[1]
price <- price[first_monday:nrow(price),]
weekdays(as.Date(rownames(head(price))[1]))

```

## Create traditional and heikin-ashi candlestick charts

```{r,candlesticks,fig.height=8,fig.width=8}

# traditional candlesticks
btc_trad <- xts(price,order.by=as.Date(rownames(price)))
chartSeries(btc_trad,log.scale=TRUE)

# heikin-ashi
ha <- HAOHLC(price)
btc_ha <- xts(ha,order.by=as.Date(rownames(ha)))
chartSeries(btc_ha,log.scale=TRUE)

price2 <- tail(price,200)

# traditional candlesticks
btc_trad <- xts(price2,order.by=as.Date(rownames(price2)))
chartSeries(btc_trad,log.scale=TRUE)

# heikin-ashi
ha <- HAOHLC(price2)
btc_ha <- xts(ha,order.by=as.Date(rownames(ha)))
chartSeries(btc_ha,log.scale=TRUE)

```

## Transform to 3 day chart

This approach can be used to convert smaller timeframe charts to higher timeframes.

In this example from daily to 3day.

```{r,transform}

timeframe <- function(price,n) {
  idx <- seq(1,nrow(price),n)
  pricel <- lapply(idx,function(i) {
    slice <- price[i:(i+(n-1)),]
    open <- slice[1,"open"]
    high <- max(slice$high)
    low <- min(slice$low)
    close <- slice[nrow(slice),"close"]
    return(c(open,high,low,close))
  })

  if ( nrow(price) %% n != 0 ) {
    remainder <- n - (nrow(price) %% n )
    slice <- tail(price,remainder)
    open <- slice[1,"open"]
    high <- max(slice$high)
    low <- min(slice$low)
    close <- slice[nrow(slice),"close"]
    pricel[[length(pricel)]] <- c(open,high,low,close)
  } 

  price2 <- do.call(rbind,pricel)
  rownames(price2) <- rownames(price)[idx]
  colnames(price2) <- c("open","high","low","close")
  return(price2)
}




```

## Generate 3day chart


```{r,Nday_candles}

ohlcnd <- timeframe(price,7)

# traditional candlesticks
btc_trad <- xts(ohlcnd,order.by=as.Date(rownames(ohlcnd)))
chartSeries(btc_trad,log.scale=TRUE)

# heikin-ashi
ha <- HAOHLC(ohlcnd)
btc_ha <- xts(ha,order.by=as.Date(rownames(ha)))
chartSeries(btc_ha,log.scale=TRUE)

ohlcnd_tail <- tail(ohlcnd,100)

# traditional candlesticks
btc_trad <- xts(ohlcnd_tail,order.by=as.Date(rownames(ohlcnd_tail)))
chartSeries(btc_trad,log.scale=TRUE)

# heikin-ashi
ha <- HAOHLC(ohlcnd_tail)
btc_ha <- xts(ha,order.by=as.Date(rownames(ha)))
chartSeries(btc_ha,log.scale=TRUE)


```

## Conclusion

The best SMAs were 11 and 43 days, and the returns were approx 3x better than simply hodling.
The best EMAs was 7 and 14 days.
Based on minimising the number of trades for only a small reduction of returns, it appears 
the 43d SMA is preferred.
Another reason why I like the 43 day is that the range from 40-44 are all giving good 
returns, this gives more confidence to this range as being profitable in future.
The fee rate strongly affects the fold return, so it is definitely worthwhile to seek 
exchanges with lower fees.
Profitability varies strongly by year with bull market most profitable and bear market least 
profitable, meaning it is probably a good idea to sit funds in fiat for 12 months following
 a cycle top.

If you would like to receive trade alerts on these strategies, let me know by email mark.ziemann{αt}gmail.com.
These alerts are sent at 30 mins before the close of the daily session and will assist you in sticking to your desired strategy.
Subscribing to alerts costs 0.0005 BTC per year.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
