---
title: "Trading based on MA and EMA crosses"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

if ( ! file.exists("btcdat.txt") ) {

  download.file(URL,destfile="btcdat.txt")

}

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl


```


## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma,fig.width=8,fig.height=8}


n=43
fee=0.002

ma <- SMA(Cl(price),n=n)

price2 <- price

price2$ma <- ma

price2$higher <- price2$Close > price2$ma

yesterday <- price2$higher

yesterday <- c(NA,yesterday)

yesterday <- yesterday[1:(length(yesterday)-1)]

price2$yesterday <- yesterday

price2$trade <- apply(price2,1,function(x) {
  as.numeric(as.logical(x[4])) - as.numeric(as.logical(x[5]))
} )

ma_v <- unlist(sapply(n:nrow(price2) , function(x) {
  ( price2$ma[x] - price2$ma[x-1] ) / price2$ma[x]     
}))

ma_a <- unlist(sapply(n:nrow(price2) , function(x) {
  ( price2$ma_v[x] - price2$ma_v[x-1] ) / price2$ma_v[x] 
}))

price2$ma_v <- c( rep(NA, n -1 ) , ma_v )

price2$ma_a <- c( rep(NA, n -1 ) , ma_a )

price2$smoothed_accelerator <- SMA(price2$ma_a,n)

HEADER=paste("moving average")

par(mfrow=c(3,1))

plot(price2$Close~as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="moving average")
grid()
lines( ma  ~ as.Date(price2$date) ,col="red")

plot(price2$ma_v~as.Date(price2$date),type="l",
  ylim=c(-0.02,0.03),xlab="Date",ylab="change in ma", main="change in moving average")
grid()

plot(price2$smoothed_accelerator ~as.Date(price2$date),type="l", 
  ylim=c(-2,5), xlab="Date",ylab="change in change in ma", main="smoothed change in change in moving average")
grid()


```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
