---
title: "Trading based on MA and EMA crosses"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")
library("reshape2")
library("gplots")
library("parallel")

NUMCORES=detectCores()

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

if ( ! file.exists("btcdat.txt") ) {

  download.file(URL,destfile="btcdat.txt")

}

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```


## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}

smacross <- function(price,n,fee) {
n1=mas[n,1]
n2=mas[n,2]
if (n1 != n2) { 
  ma1 <- SMA(Cl(price),n=n1)
  ma2 <- SMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- apply(price2,1,function(x) {
    as.numeric(as.logical(x[5])) - as.numeric(as.logical(x[6]))
  } )
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(prod(trades$diff))
} else {
  return(1)
}
}

```

## Exponential moving average function


```{r,ema}

emacross <- function(price,n,fee) {
n1=mas[n,1]
n2=mas[n,2]
if (n1 != n2) {
  ma1 <- EMA(Cl(price),n=n1)
  ma2 <- EMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- apply(price2,1,function(x) {
    as.numeric(as.logical(x[5])) - as.numeric(as.logical(x[6]))
  } )
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(prod(trades$diff))
} else {
  return(1)
}
}

```

## Backtesting

```{r,backtest}

fee=0.001

mas <- expand.grid(4:200,4:200)

res_sma <- mclapply(1:nrow(mas),function(n) {
 smacross(price,n,fee) 
} , mc.cores=NUMCORES)
mas1 <- as.data.frame(mas)
mas1$returns <- unlist(res_sma)
head(mas1[order(-mas1$returns),],20)

res_ema <- mclapply(1:nrow(mas),function(n) {
  emacross(price,n,fee) 
} ,  mc.cores=NUMCORES)
mas2 <- as.data.frame(mas)
mas2$returns <- unlist(res_ema)
head(mas2[order(-mas2$returns),],20)

```

## Plots

```{r,plots,fig.height=8,fig.width=8}

w <- as.matrix(acast(mas1, Var1~Var2, value.var="returns"))

w[1:10,1:5]

colfunc <- colorRampPalette(c("white", "yellow" , "orange" , "red" , "darkred"))

colfunc <- colorRampPalette(c("white", "red"))

heatmap.2(w,scale="none",trace="none",dendrogram="none", Rowv=FALSE, Colv=FALSE, 
  col=colfunc(25), main="SMA crosses", xlab="short period", ylab="long period")

heatmap.2(w[1:50,1:50],scale="none",trace="none",dendrogram="none", Rowv=FALSE, 
  Colv=FALSE, col=colfunc(25) , main="SMA crosses",
  xlab="short period", ylab="long period")

heatmap.2(w[1:26,1:20],scale="none",trace="none",dendrogram="none", Rowv=FALSE, 
  Colv=FALSE, col=colfunc(25), main="SMA crosses",
  xlab="short period", ylab="long period")

```

## Optimisation

```{r,best_sma,fig.width=10,fig.height=8}

best_sma = rownames(head(mas1[order(-mas1$returns),],1))
mas[which(rownames(mas1) == best_sma),]
n1=mas[which(rownames(mas1) == best_sma),1]
n2=mas[which(rownames(mas1) == best_sma),2]

if (n1 != n2) {
  ma1 <- SMA(Cl(price),n=n1)
  ma2 <- SMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- apply(price2,1,function(x) {
    as.numeric(as.logical(x[5])) - as.numeric(as.logical(x[6]))
  } )
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  returns <- prod(trades$diff)
} else {
  returns <- 1
}
returns
nrow(trades)
sma_trades <- trades
SMA_HEADER=paste(n1 , "and", n2 ,"period MA cross")

plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma1~ as.Date(price$date) ,col="red")
lines(ma2~ as.Date(price$date) ,col="blue")

```

Same thing for EMA crosses.

```{r,best_ema,fig.width=10,fig.height=8}

best_ema = rownames(head(mas2[order(-mas2$returns),],1))
mas[which(rownames(mas2) == best_ema),]
n1=mas[which(rownames(mas2) == best_ema),1]
n2=mas[which(rownames(mas2) == best_ema),2]

if (n1 != n2) {
  ma1 <- EMA(Cl(price),n=n1)
  ma2 <- EMA(Cl(price),n=n2)
  price2 <- price
  price2$ma1 <- ma1
  price2$ma2 <- ma2
  price2$higher <- price2$ma2 > price2$ma1
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- apply(price2,1,function(x) {
    as.numeric(as.logical(x[5])) - as.numeric(as.logical(x[6]))
  } )
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  returns <- prod(trades$diff)
} else {
  returns <- 1
}
returns
nrow(trades)
ema_trades <- trades
EMA_HEADER=paste(n1 , "and", n2 ,"period EMA cross")

plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(ma1~ as.Date(price$date) ,col="red")
lines(ma2~ as.Date(price$date) ,col="blue")

```

## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_sma_fees <- sapply(fee_levels,function(x) { 
  smacross(price,best_sma,fee=x)
} )
plot(fee_levels,res_sma_fees ,pch=19,xlab="fee proportion",ylab="fold return",
  main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

res_ema_fees <- sapply(fee_levels,function(x) { 
  emacross(price,best_ema,fee=x)
} )
plot(fee_levels,res_ema_fees ,pch=19,xlab="fee proportion",ylab="fold return",
  main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades$year <- sapply(strsplit(sma_trades[,"sell_date"],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,"diff"] } )
vioplot(sma_yearly)
barplot(sapply(sma_yearly,prod))
abline(h=1,col="red",lty=2)

ema_trades$year <- sapply(strsplit(ema_trades[,"sell_date"],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,"diff"] } )
vioplot(ema_yearly)
barplot(sapply(ema_yearly,prod))
abline(h=1,col="red",lty=2)

```


## Conclusion

Hodling BTC over this period gave a 378 fold return.
The best simple moving average was 18 days, which gave a 1366 fold return (with fees at 0.1%).
The best exponential moving average was 16 days, which gave a return of 1680 fold.
The fee rate strongly affects the fold return, so it is definitely worthwhile to seek lower price exchanges.
Profitability varies strongly by year with bull market most profitable and bear market least profitable.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
