---
title: "Trading ETH based on MA and EMA crosses"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trialing simple moving averages to find the one which performs the best.
We are focusing on ethereum, not bitcoin, in the period 2015 to 2021.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")


```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=2015-09-01&time_end=",mydate,sep="")
download.file(URL,destfile="ethdat.txt")

ethdat <- fromJSON("ethdat.txt")
ethprice <- ethdat$data$quotes
ethprice <- data.frame(ethprice$time_close,ethprice$quote$USD$close,stringsAsFactors=FALSE)
colnames(ethprice) <- c("date","Close")
ethprice$date <- sapply(strsplit(ethprice$date,"T"),"[[",1)

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2015-09-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")

btcdat <- fromJSON("btcdat.txt")
btcprice <- btcdat$data$quotes
btcprice <- data.frame(btcprice$time_close,btcprice$quote$USD$close,stringsAsFactors=FALSE)
colnames(btcprice) <- c("date","Close")
btcprice$date <- sapply(strsplit(btcprice$date,"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- ethprice[c(1,nrow(ethprice)),2][2] 
mystart <- ethprice[c(1,nrow(ethprice)),2][1] 
hodl <- (myend - mystart) / mystart
hodl


```


## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}

sma <- function(btcprice, ethprice, n, fee) {
ma <- SMA(Cl(btcprice),n=n)
price2 <- btcprice
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2$Close <- ethprice$Close
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(prod(trades$diff))
}



```

## Exponential moving average function


```{r,ema}

ema <- function(btcprice, ethprice, n, fee) {
ma <- EMA(Cl(btcprice),n=n)
price2 <- btcprice
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2$Close <- ethprice$Close
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(prod(trades$diff))
}



```

## Backtesting

```{r,backtest}

fee=0.001

res_sma <- c(rep(0,4),sapply(5:200,function(n) { sma(btcprice,ethprice,n,fee) } ) )

res_ema <- c(rep(0,4),sapply(5:200,function(n) { ema(btcprice,ethprice,n,fee) } ) )

```

## Plots

```{r,plots,fig.height=8,fig.width=8}

plot(res_sma,pch=19,xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(res_sma,pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(res_ema,pch=19,xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(res_ema,pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

```

## Optimisation

```{r,best_sma,fig.width=10,fig.height=8}

best_sma = which(res_sma==max(res_sma))
best_sma

ma <- SMA(Cl(price),n=best_sma)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
trades 
prod(trades$diff)
SMA_HEADER=paste(best_sma,"day moving average")

plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")
sma_trades <- trades

```

```{r,best_ema,fig.width=10,fig.height=8}

best_ema = which(res_ema==max(res_ema))
best_ema

ma <- EMA(Cl(price),n=best_sma)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
trades
nrow(trades)
prod(trades$diff)
EMA_HEADER=paste(best_ema,"day exponential moving average")

plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")
ema_trades <- trades

```

## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_sma_fees <- sapply(fee_levels,function(x) { sma(price,best_sma,fee=x) } )
plot(fee_levels,res_sma_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

res_ema_fees <- sapply(fee_levels,function(x) { ema(price,best_ema,fee=x) } )
plot(fee_levels,res_ema_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades$year <- sapply(strsplit(sma_trades[,3],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,5] } )
vioplot(sma_yearly)
barplot(sapply(sma_yearly,prod))
abline(h=1,col="red",lty=2)

ema_trades$year <- sapply(strsplit(ema_trades[,3],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,5] } )
vioplot(ema_yearly)
barplot(sapply(ema_yearly,prod))
abline(h=1,col="red",lty=2)

```


## Conclusion

Hodling ETH over this period gave a 521 fold return.
The best simple moving average was 34 days, which gave a 20970 fold return (with fees at 0.1%).
The best exponential moving average was 17 days, which gave a return of 10492 fold.
The fee rate affects the fold return a bit less than BTC. 
Still, it is definitely worthwhile to seek lower price exchanges.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
