---
title: "Trading based on MA and EMA crosses"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

```

## Get data

Obtaining historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="ethdat.txt")

dat <- fromJSON("ethdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}

sma <- function(price,n,fee) {
ma <- SMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}


```

## Exponential moving average function


```{r,ema}

ema <- function(price,n,fee) {
ma <- EMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(c(prod(trades$diff),nrow(trades)))
}

```

## Backtesting

```{r,backtest}

fee=0.001

myrange <- 3:200

res <- sapply(myrange,function(n) { sma(price,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_sma <- res

res <- sapply(myrange,function(n) { ema(price,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_ema <- res

```

## Plots

```{r,plots,fig.height=8,fig.width=8}

plot(x=myrange, y=res_sma$returns, pch=19,
  xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_sma$returns, pch=19,
  xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_ema$returns, pch=19,
  xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_ema$returns, pch=19,
  xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(res_sma$mean,res_sma$returns, main="SMA",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$mean,res_sma$returns , cex=1, labels=rownames(res_sma))
abline(h=hodl,col="red")

plot(res_ema$mean,res_ema$returns, main="EMA",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$mean,res_ema$returns , cex=1, labels=rownames(res_ema))
abline(h=hodl,col="red")

plot(res_sma$trades,res_sma$returns, main="SMA",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$trades,res_sma$returns , cex=1, labels=rownames(res_sma))
abline(h=hodl,col="red")

plot(res_ema$trades,res_ema$returns, main="SMA",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$trades,res_ema$returns , cex=1, labels=rownames(res_ema))
abline(h=hodl,col="red")


```

## Optimisation

```{r,best_sma,fig.width=10,fig.height=8}

best_sma = which(res_sma==max(res_sma))
best_sma
ma <- SMA(Cl(price),n=best_sma)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]  
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
sma_trades <- cbind(buy,sell)
nrow(sma_trades)
sma_trades$diff <- ( ( sma_trades$sell_price - sma_trades$buy_price ) / sma_trades$buy_price ) +1 - (2 * fee )
sma_trades 
prod(sma_trades$diff)
SMA_HEADER=paste(best_sma,"day moving average")

plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```


```{r,sma43,fig.width=10,fig.height=8}

best_sma = 43
best_sma
ma <- SMA(Cl(price),n=best_sma)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
sma_trades <- cbind(buy,sell)
nrow(sma_trades)
sma_trades$diff <- ( ( sma_trades$sell_price - sma_trades$buy_price ) / sma_trades$buy_price ) +1 - (2 * fee )
sma_trades
prod(sma_trades$diff)
SMA_HEADER=paste(best_sma,"day moving average")

plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```




```{r,best_ema,fig.width=10,fig.height=8}

best_ema = which(res_ema==max(res_ema))
best_ema
ma <- EMA(Cl(price),n=best_ema)
price2 <- price
price2$ma <- ma
price2$higher <- price2$Close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
ema_trades <- cbind(buy,sell)
ema_trades$diff <- ( ( ema_trades$sell_price - ema_trades$buy_price ) / ema_trades$buy_price ) +1 - ( 2 * fee )
ema_trades
nrow(ema_trades)
prod(ema_trades$diff)
EMA_HEADER=paste(best_ema,"day exponential moving average")

plot(price$Close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```




## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_sma_fees <- sapply(fee_levels,function(x) { sma(price,best_sma,fee=x) } )[1,]
plot(fee_levels,res_sma_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

res_ema_fees <- sapply(fee_levels,function(x) { ema(price,best_ema,fee=x) } )[1,]
plot(fee_levels,res_ema_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades$year <- sapply(strsplit(sma_trades[,3],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,5] } )
vioplot(sma_yearly)

yr <- sapply(sma_yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

ema_trades$year <- sapply(strsplit(ema_trades[,3],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,5] } )
vioplot(ema_yearly)

yr <- sapply(ema_yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
