---
title: "Trading based on MA and EMA crosses"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

## Intro

Trading off moving averages, exponenial moving averages and crosses between these is a popular method for timing entry and exit of positions.
In this blog post I will be trilaing simple moving averages to find the one which performs the best.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("gplots")
library("vioplot")

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

if ( ! file.exists("btcdat.txt") ) {

  download.file(URL,destfile="btcdat.txt")

}

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close)
colnames(price) <- c("date","close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl


```


## Moving average function

Assuming 0.1% trading fee, as per binance current rate.

```{r,sma}

sma <- function(price,n,fee,margin) {
ma <- SMA(Cl(price),n=n) * ( 1 + margin )
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- apply(price2,1,function(x) {
  as.numeric(as.logical(x[4])) - as.numeric(as.logical(x[5]))
} )
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(prod(trades$diff))
}


```

## Exponential moving average function


```{r,ema}

ema <- function(price, n, fee, margin ) {
ma <- EMA(Cl(price),n=n) * (1 + margin)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- apply(price2,1,function(x) {
  as.numeric(as.logical(x[4])) - as.numeric(as.logical(x[5]))
} )
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(prod(trades$diff))
}

```

## Backtesting

Previously, we showed the profitability of various SMAs and EMAs.
The 11 day SMA gave a return of 1704 fold with 187 trades.
Here, I did the same analysis with one difference.
The MA is multiplied by a margin factor between -0.1and 0.1.

```{r,backtest,fig.height=8,fig.width=8}

fee=0.002

intervals <- c(rep(0,5),6:400)

margins <- c(-0.15 , -0.125 , -0.1, -0.075 , -0.05, -0.02, -0.015, -0.01, 0, 0.01 , 0.015, 0.02, 0.05, 0.075 , 0.1, 0.125, 0.15)

res_sma <- lapply(margins, function(margin) {
  c(0,0,sapply(3:200,function(n) { sma(price,n,fee,margin) } ) )
} )

res_sma <- t(as.data.frame(do.call(rbind, res_sma)))
colnames(res_sma) <- margins
rownames(res_sma) <- 1:nrow(res_sma)

sma_MAX = max(res_sma)
sma_INT = names(which(apply(res_sma,1,function(x) {length(which(x==sma_MAX))})>0))
sma_MARG = names(which(apply(res_sma,2,function(x) {length(which(x==sma_MAX))})>0))

sma_MAX
sma_INT
sma_MARG

colfunc <- colorRampPalette(c("white", "red"))
heatmap.2(res_sma[1:50,],col=colfunc(25),trace="none",scale="none", dendrogram="none",Rowv=FALSE,Colv=FALSE)

res_ema <- lapply(margins, function(margin) {
  c(0,0,sapply(3:200,function(n) { ema(price,n,fee,margin) } ) )
} )

res_ema <- t(as.data.frame(do.call(rbind, res_ema)))
colnames(res_ema) <- margins
rownames(res_ema) <- 1:nrow(res_ema)

ema_MAX = max(res_ema)
ema_INT = names(which(apply(res_ema,1,function(x) {length(which(x==ema_MAX))})>0))
ema_MARG = names(which(apply(res_ema,2,function(x) {length(which(x==ema_MAX))})>0))

ema_MAX
ema_INT
ema_MARG

heatmap.2(res_ema[1:50,],col=colfunc(25),trace="none",scale="none", dendrogram="none",Rowv=FALSE,Colv=FALSE)

```

## Plots

```{r,plots,fig.height=8,fig.width=8}

plot(res_sma[,which(colnames(res_sma) %in% sma_MARG)],
  pch=19,xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(res_sma[,which(colnames(res_sma) %in% sma_MARG)],
  pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(res_ema[,which(colnames(res_ema) %in% ema_MARG)],
  pch=19,xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(res_ema[,which(colnames(res_ema) %in% ema_MARG)],
  pch=19,xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

```

## Optimisation

```{r,best_sma,fig.width=10,fig.height=8}

ma <- SMA(Cl(price),n=as.numeric(sma_INT)) * ( 1 + as.numeric(sma_MARG ) )
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- apply(price2,1,function(x) {
  as.numeric(as.logical(x[4])) - as.numeric(as.logical(x[5]))
} )
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]  
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
sma_trades <- cbind(buy,sell)
nrow(sma_trades)
sma_trades$diff <- ( ( sma_trades$sell_price - sma_trades$buy_price ) / sma_trades$buy_price ) +1 - (2 * fee )
sma_trades 
prod(sma_trades$diff)
SMA_HEADER=paste(sma_INT,"day moving average.", sma_MARG , "margin")

plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```

```{r,best_ema,fig.width=10,fig.height=8}

best_ema = which(res_ema==max(res_ema))
best_ema
ma <- EMA(Cl(price),n=as.numeric(ema_INT)) * ( 1 + as.numeric(ema_MARG ) )
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- apply(price2,1,function(x) {
  as.numeric(as.logical(x[4])) - as.numeric(as.logical(x[5]))
} )
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
ema_trades <- cbind(buy,sell)
ema_trades$diff <- ( ( ema_trades$sell_price - ema_trades$buy_price ) / ema_trades$buy_price ) +1 - ( 2 * fee )
ema_trades
nrow(ema_trades)
prod(ema_trades$diff)
EMA_HEADER=paste(ema_INT,"day exponential moving average.", ema_MARG , "margin")

plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```

## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_sma_fees <- sapply(fee_levels,function(x) { sma(price,as.numeric(sma_INT),fee=x,margin=as.numeric(sma_MARG)) } )
plot(fee_levels,res_sma_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

res_ema_fees <- sapply(fee_levels,function(x) { ema(price,as.numeric(ema_INT),fee=x,margin=as.numeric(ema_MARG)) } )
plot(fee_levels,res_ema_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades$year <- sapply(strsplit(sma_trades[,3],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,5] } )
sma_yearly <- sma_yearly[which(lapply(sma_yearly,length)>0)]
vioplot(sma_yearly)
barplot(sapply(sma_yearly,prod))
abline(h=1,col="red",lty=2)

ema_trades$year <- sapply(strsplit(ema_trades[,3],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,5] } )
ema_yearly <- ema_yearly[which(lapply(ema_yearly,length)>0)]
vioplot(ema_yearly)
barplot(sapply(ema_yearly,prod))
abline(h=1,col="red",lty=2)

```

## Conclusion

Hodling BTC over this period gave a 378 fold return.
The best simple moving average was 18 days, which gave a 1366 fold return (with fees at 0.1%).
The best exponential moving average was 16 days, which gave a return of 1680 fold.
The fee rate strongly affects the fold return, so it is definitely worthwhile to seek lower price exchanges.
Naturally, bull market years are more profitable and bear market less profitable.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
