---
title: "Pivot points for BTC and ETH"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 15
    fig_height: 12
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/pivots.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

From Tradingview:

>A pivot point is a technical analysis indicator, or calculations, used to determine the overall trend of the market over different time frames. The pivot point itself is simply the average of the intraday high and low, and the closing price from the previous trading day. On the subsequent day, trading above the pivot point is thought to indicate ongoing bullish sentiment, while trading below the pivot point indicates bearish sentiment.

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("lubridate")
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
})

Sys.Date()

```

Reminder: this is not financial advice.

## Pivot point indicator

Also called Central Pivot Range (CPR).

central pivot point (pp) =  (prev_low + prev_high + prev_close) / 3

Resistance points:

r1 = 2  * pp - prev_low

r2 = pp + ( prev_h - prev_low )

r3 = pp + 2 * ( prev_high  - prev_low )

support points

s1 = 2 * pp - prev_high 

s2 = pp - ( prev_high  - prev_low )

s3 = pp - 2 * ( prev_high  - prev_low )

In this report, the pivot points are shown with dashed lines.
Pivot line is black.
S1 line is red.
S2 line is orange.
R1 line is blue.
R2 line is gren.

```{r,piv_fun}

# this function is more for analysing tabular data
pivots_weekly <- function(price) {
  weeks <- paste(year(price$date),week(price$date))
  wl <- lapply(X=unique(weeks) , FUN=function(w) {
    rows <- which(weeks ==w )
    wdat <- price[rows,]
    hi <- max(wdat$high)
    lo <- min(wdat$low)
    cl <- wdat[nrow(wdat),"close"]
    d <- wdat[nrow(wdat),"date"]
    wagg <- c("date"=d,"high"=hi,"low"=lo,"close"=cl)
  }  )
  wmx <- as.data.frame(do.call(rbind,wl))
  wmx$date <- ymd(wmx$date)
  wmx$high <- as.numeric(wmx$high)
  wmx$low <- as.numeric(wmx$low)
  wmx$close <- as.numeric(wmx$close)
  hlc2 <- HLC(wmx)
  pp <- rowMeans(hlc2)
  r1 <- 2  * pp - hlc2$low
  r2 <- pp + ( hlc2$high - hlc2$low )
  r3 <- pp + 2 * ( hlc2$high  - hlc2$low )
  s1 <- 2 * pp - hlc2$high
  s2 <- pp - ( hlc2$high  - hlc2$low )
  s3 <- pp - 2 * ( hlc2$high  - hlc2$low )
  weeks2 <- weeks[which(weeks != weeks[1])]
  pp <- rep(pp[1:(length(pp)-1)],table(weeks2))
  r1 <- rep(r1[1:(length(r1)-1)],table(weeks2))
  r2 <- rep(r2[1:(length(r2)-1)],table(weeks2))
  r3 <- rep(r3[1:(length(r3)-1)],table(weeks2))
  s1 <- rep(s1[1:(length(s1)-1)],table(weeks2))
  s2 <- rep(s2[1:(length(s2)-1)],table(weeks2))
  s3 <- rep(s3[1:(length(s3)-1)],table(weeks2))
  pivots <- as.data.frame(cbind(pp,r1,r2,r3,s1,s2,s3)) 
  deduct_start <- ( nrow(price) - nrow(pivots) ) + 1 
  price2 <- price[deduct_start:nrow(price),]
  res <- cbind(price2,pivots)
  return(res)
}
#pivw <- pivots_weekly(price)

# this function is more for analysing tabular data
pivots_monthly <- function(price) {
  months <- paste(year(price$date),month(price$date))
  wl <- lapply(X=unique(months) , FUN=function(m) {
    rows <- which(months == m )
    mdat <- price[rows,]
    hi <- max(mdat$high)
    lo <- min(mdat$low)
    cl <- mdat[nrow(mdat),"close"]
    d <- mdat[nrow(mdat),"date"]
    magg <- c("date"=d,"high"=hi,"low"=lo,"close"=cl)
  }  )
  wmx <- as.data.frame(do.call(rbind,wl))
  wmx$date <- ymd(wmx$date)
  wmx$high <- as.numeric(wmx$high)
  wmx$low <- as.numeric(wmx$low)
  wmx$close <- as.numeric(wmx$close)
  hlc2 <- HLC(wmx)
  pp <- rowMeans(hlc2)
  r1 <- 2  * pp - hlc2$low
  r2 <- pp + ( hlc2$high - hlc2$low )
  r3 <- pp + 2 * ( hlc2$high  - hlc2$low )
  s1 <- 2 * pp - hlc2$high
  s2 <- pp - ( hlc2$high  - hlc2$low )
  s3 <- pp - 2 * ( hlc2$high  - hlc2$low )
  months2 <- months[which(months != months[1])]
  pp <- rep(pp[1:(length(pp)-1)],table(months2))
  r1 <- rep(r1[1:(length(r1)-1)],table(months2))
  r2 <- rep(r2[1:(length(r2)-1)],table(months2))
  r3 <- rep(r3[1:(length(r3)-1)],table(months2))
  s1 <- rep(s1[1:(length(s1)-1)],table(months2))
  s2 <- rep(s2[1:(length(s2)-1)],table(months2))
  s3 <- rep(s3[1:(length(s3)-1)],table(months2))
  pivots <- as.data.frame(cbind(pp,r1,r2,r3,s1,s2,s3))
  deduct_start <- ( nrow(price) - nrow(pivots) ) + 1
  price2 <- price[deduct_start:nrow(price),]
  res <- cbind(price2,pivots)
  return(res)
}
#pivm <- pivots_monthly(price)

# this function is more for analysing tabular data
pivots_quarterly <- function(price) {
  quarters <- paste(year(price$date),quarter(price$date))
  wl <- lapply(X=unique(quarters) , FUN=function(m) {
    rows <- which(quarters == m )
    mdat <- price[rows,]
    hi <- max(mdat$high)
    lo <- min(mdat$low)
    cl <- mdat[nrow(mdat),"close"]
    d <- mdat[nrow(mdat),"date"]
    magg <- c("date"=d,"high"=hi,"low"=lo,"close"=cl)
  }  )
  wmx <- as.data.frame(do.call(rbind,wl))
  wmx$date <- ymd(wmx$date)
  wmx$high <- as.numeric(wmx$high)
  wmx$low <- as.numeric(wmx$low)
  wmx$close <- as.numeric(wmx$close)
  hlc2 <- HLC(wmx)
  pp <- rowMeans(hlc2)
  r1 <- 2  * pp - hlc2$low
  r2 <- pp + ( hlc2$high - hlc2$low )
  r3 <- pp + 2 * ( hlc2$high  - hlc2$low )
  s1 <- 2 * pp - hlc2$high
  s2 <- pp - ( hlc2$high  - hlc2$low )
  s3 <- pp - 2 * ( hlc2$high  - hlc2$low )
  quarters2 <- quarters[which(quarters != quarters[1])]
  pp <- rep(pp[1:(length(pp)-1)],table(quarters2))
  r1 <- rep(r1[1:(length(r1)-1)],table(quarters2))
  r2 <- rep(r2[1:(length(r2)-1)],table(quarters2))
  r3 <- rep(r3[1:(length(r3)-1)],table(quarters2))
  s1 <- rep(s1[1:(length(s1)-1)],table(quarters2))
  s2 <- rep(s2[1:(length(s2)-1)],table(quarters2))
  s3 <- rep(s3[1:(length(s3)-1)],table(quarters2))
  pivots <- as.data.frame(cbind(pp,r1,r2,r3,s1,s2,s3))
  deduct_start <- ( nrow(price) - nrow(pivots) ) + 1
  price2 <- price[deduct_start:nrow(price),]
  res <- cbind(price2,pivots)
  return(res)
}
#pivm <- pivots_quarterly(price)

# this function is more for analysing tabular data
pivots_yearly <- function(price) {
  years <- year(price$date)
  wl <- lapply(X=unique(years) , FUN=function(m) {
    rows <- which(years == m )
    mdat <- price[rows,]
    hi <- max(mdat$high)
    lo <- min(mdat$low)
    cl <- mdat[nrow(mdat),"close"]
    d <- mdat[nrow(mdat),"date"]
    magg <- c("date"=d,"high"=hi,"low"=lo,"close"=cl)
  }  )
  wmx <- as.data.frame(do.call(rbind,wl))
  wmx$date <- ymd(wmx$date)
  wmx$high <- as.numeric(wmx$high)
  wmx$low <- as.numeric(wmx$low)
  wmx$close <- as.numeric(wmx$close)
  hlc2 <- HLC(wmx)
  pp <- rowMeans(hlc2)
  r1 <- 2  * pp - hlc2$low
  r2 <- pp + ( hlc2$high - hlc2$low )
  r3 <- pp + 2 * ( hlc2$high  - hlc2$low )
  s1 <- 2 * pp - hlc2$high
  s2 <- pp - ( hlc2$high  - hlc2$low )
  s3 <- pp - 2 * ( hlc2$high  - hlc2$low )
  years2 <- years[which(years != years[1])]
  pp <- rep(pp[1:(length(pp)-1)],table(years2))
  r1 <- rep(r1[1:(length(r1)-1)],table(years2))
  r2 <- rep(r2[1:(length(r2)-1)],table(years2))
  r3 <- rep(r3[1:(length(r3)-1)],table(years2))
  s1 <- rep(s1[1:(length(s1)-1)],table(years2))
  s2 <- rep(s2[1:(length(s2)-1)],table(years2))
  s3 <- rep(s3[1:(length(s3)-1)],table(years2))
  pivots <- as.data.frame(cbind(pp,r1,r2,r3,s1,s2,s3))
  deduct_start <- ( nrow(price) - nrow(pivots) ) + 1
  price2 <- price[deduct_start:nrow(price),]
  res <- cbind(price2,pivots)
  return(res)
}
#pivm <- pivots_yearly(price)

# this function is more for plotting the pivots nicely
pivots_weekly2 <- function(price) {
  weeks <- paste(year(price$date),week(price$date))
  res <- lapply(X=unique(weeks) , FUN=function(m) {
    rows <- which(weeks == m )
    mdat <- price[rows,]
    hi <- max(mdat$high)
    lo <- min(mdat$low)
    cl <- mdat[nrow(mdat),"close"]
    pp <- mean(c(hi, lo, cl))
    r1 <- ( 2  * pp ) - lo
    r2 <- pp + ( hi - lo )
    r3 <- pp + ( 2 * ( hi - lo ) )
    s1 <- ( 2 * pp ) - hi
    s2 <- pp - ( hi  - lo )
    s3 <- pp - ( 2 * ( hi - lo ) )
    d0 <- mdat[1,"date"]
    d1 <- as.Date(d0)+7
    d2 <- d1+6
    magg <- list("date1"=d1, "date2"=d2, "high"=hi,"low"=lo,"close"=cl ,
      "pp"=pp, "r1"=r1, "r2"=r2, "r3"=r3, "s1"=s1, "s2"=s2, "s3"=s3)
  }  )
  res
}

# this function is more for plotting the pivots nicely
pivots_monthly2 <- function(price) {
  months <- paste(year(price$date),month(price$date))
  res <- lapply(X=unique(months) , FUN=function(m) {
    rows <- which(months == m )
    mdat <- price[rows,]
    hi <- max(mdat$high)
    lo <- min(mdat$low)
    cl <- mdat[nrow(mdat),"close"]
    pp <- mean(c(hi, lo, cl))
    r1 <- ( 2  * pp ) - lo
    r2 <- pp + ( hi - lo )
    r3 <- pp + ( 2 * ( hi - lo ) )
    s1 <- ( 2 * pp ) - hi
    s2 <- pp - ( hi  - lo )
    s3 <- pp - ( 2 * ( hi - lo ) )
    d <- mdat[nrow(mdat),"date"]
    yr <- year(d)
    mo <- month(d) +1 
    if ( mo > 12 ) { mo=1 ; yr <- yr +1 }
    d1 <- as.Date(paste(yr,mo,"01",sep="-"))
    yr <- year(d)
    mo <- month(d) +2
    if ( mo > 13 ) { mo=2 ; yr <- yr +1 }
    if ( mo > 12 ) { mo=1 ; yr <- yr +1 }
    d2 <- as.Date(paste(yr,mo,"01",sep="-"))-1
    magg <- list("date1"=d1, "date2"=d2, "high"=hi,"low"=lo,"close"=cl ,
      "pp"=pp, "r1"=r1, "r2"=r2, "r3"=r3, "s1"=s1, "s2"=s2, "s3"=s3)
  }  )
  res
}


# this function is more for plotting the pivots nicely
pivots_quarterly2 <- function(price) {
  quarters <- paste(year(price$date),quarter(price$date))
  res <- lapply(X=unique(quarters) , FUN=function(m) {
    rows <- which(quarters == m )
    mdat <- price[rows,]
    hi <- max(mdat$high)
    lo <- min(mdat$low)
    cl <- mdat[nrow(mdat),"close"]
    pp <- mean(c(hi, lo, cl))
    r1 <- ( 2  * pp ) - lo
    r2 <- pp + ( hi - lo )
    r3 <- pp + ( 2 * ( hi - lo ) )
    s1 <- ( 2 * pp ) - hi
    s2 <- pp - ( hi  - lo )
    s3 <- pp - ( 2 * ( hi - lo ) )
    d <- as.Date(mdat[1,"date"])
    y1 <- year(d)
    q <- quarter(d)
    m0 <- (q*3)-3
    m1 <- (q*3) + 1
    if ( m1 > 12 ) { m1=1 ; y1=y1+1 }
    d1 <- as.Date(paste(y1,m1,"01",sep="-"))
    m2 <- m1 + 3
    y2 <- y1
    if ( m2 > 12 ) { m2=1 ; y2=y1+1 }
    d2 <- as.Date(paste(y2,m2,"01",sep="-"))-1
    magg <- list("date1"=d1, "date2"=d2, "high"=hi,"low"=lo,"close"=cl ,
      "pp"=pp, "r1"=r1, "r2"=r2, "r3"=r3, "s1"=s1, "s2"=s2, "s3"=s3)
  }  )
  res
}

# this function is more for plotting the pivots nicely
pivots_yearly2 <- function(price) {
  years <- year(price$date)
  res <- lapply(X=unique(years) , FUN=function(m) {
    rows <- which(years == m )
    mdat <- price[rows,]
    hi <- max(mdat$high)
    lo <- min(mdat$low)
    cl <- mdat[nrow(mdat),"close"]
    pp <- mean(c(hi, lo, cl))
    r1 <- ( 2  * pp ) - lo
    r2 <- pp + ( hi - lo )
    r3 <- pp + ( 2 * ( hi - lo ) )
    s1 <- ( 2 * pp ) - hi
    s2 <- pp - ( hi  - lo )
    s3 <- pp - ( 2 * ( hi - lo ) )
    d1 <- as.Date(paste(m+1,"-01-01",sep=""))
    d2 <- as.Date(paste(m+1,"-12-31",sep=""))
    magg <- list("date1"=d1, "date2"=d2, "high"=hi,"low"=lo,"close"=cl ,
      "pp"=pp, "r1"=r1, "r2"=r2, "r3"=r3, "s1"=s1, "s2"=s2, "s3"=s3)
  }  )
  res
}

```

## Fetch BTC data

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")

dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high,
  price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Weekly Pivot points for BTC

```{r,pivw1}

price2 <- price
pivml <- pivots_weekly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,50)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Weekly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Now zoom in on the recent action (past 200 days).

You can see that an estimate of next week's values are shown despite the week not being finished yet.

```{r,pivw2}

price2 <- tail(price,200)
pivml <- pivots_weekly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,50)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Weekly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivw3}

piv <- pivots_weekly(price2)
tail(piv) %>% kbl(caption="Recent price action and weekly pivots") %>%  kable_styling()

```

## Monthly Pivot points for BTC

```{r,pivm1}

price2 <- price
pivml <- pivots_monthly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,50)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Monthly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Now zoom in on the recent action (past 400 days).

```{r,pivm2}

price2 <- tail(price,400)
pivml <- pivots_monthly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,50)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Monthly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivm3}

piv <- pivots_monthly(price)
tail(piv) %>% kbl(caption="Recent price action and monthly pivots") %>%  kable_styling()

```

## Quarterly Pivot points for BTC

```{r,pivq1}

price2 <- price
pivml <- pivots_quarterly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,50)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Quarterly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Now zoom in on the recent action (past 400 days).

```{r,pivq2}

price2 <- tail(price,400)
pivml <- pivots_quarterly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,50)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Quarterly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivq3}

piv <- pivots_quarterly(price)
tail(piv) %>% kbl(caption="Recent price action and quarterly pivots") %>%  kable_styling()

```

## Yearly Pivot points for BTC

```{r,pivy1}

price2 <- price
pivml <- pivots_yearly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,50)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Yearly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivy2}

piv <- pivots_yearly(price)
tail(piv) %>% kbl(caption="Recent price action and yearly pivots") %>%  kable_styling()

```

## Fetch ETH data

Obtaining ETH historical data (daily) from CoinMarketCap.com from 2015 to present.

```{r,getdata_eth}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")

dat <- fromJSON("ethdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high,
  price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Weekly Pivot points for ETH

```{r,pivw1e}

price2 <- price
pivml <- pivots_weekly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,.1)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Weekly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Now zoom in on the recent action (past 200 days).

You can see that an estimate of next week's values are shown despite the week not being finished yet.

```{r,pivw2e}

price2 <- tail(price,200)
pivml <- pivots_weekly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,.1)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Weekly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivw3e}

piv <- pivots_weekly(price2)
tail(piv) %>% kbl(caption="Recent price action and weekly pivots") %>%  kable_styling()

```

## Monthly Pivot points for ETH

```{r,pivm1e}

price2 <- price
pivml <- pivots_monthly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,.1)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Monthly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Now zoom in on the recent action (past 400 days).

```{r,pivm2e}

price2 <- tail(price,400)
pivml <- pivots_monthly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,.1)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Monthly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivm3e}

piv <- pivots_monthly(price)
tail(piv) %>% kbl(caption="Recent price action and monthly pivots") %>%  kable_styling()

```

## Quarterly Pivot points for ETH

```{r,pivq1e}

price2 <- price
pivml <- pivots_quarterly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,.1)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Quarterly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Now zoom in on the recent action (past 400 days).

```{r,pivq2e}

price2 <- tail(price,400)
pivml <- pivots_quarterly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,.1)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Quarterly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivq3e}

piv <- pivots_quarterly(price)
tail(piv) %>% kbl(caption="Recent price action and quarterly pivots") %>%  kable_styling()

```

## Yearly Pivot points for ETH

```{r,pivy1e}

price2 <- price
pivml <- pivots_yearly2(price2)

start_date <- pivml[[1]]$date1
end_date <- pivml[[length(pivml)]]$date2
ymax <- max( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",7) ) )
ymin <- min( c( sapply(pivml,"[[",5) , sapply(pivml,"[[",10) ) )
ymin <- max(ymin,.1)

plot( c(start_date,end_date) ,  c(ymin,ymax)  ,type="l",log="y",lwd=0,
  xlab="Date",ylab="price (USD)",main="Yearly pivot points standard")
grid()
lines(as.Date(price2$date) , price2$close , col="black")

x <- lapply(pivml, function(x) {
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[6] , x[6] ) , col="black") #pp
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[10] , x[10] ) , col="red") #s1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[11] , x[11] ) , col="orange") #s2
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[7] , x[7] ) , col="blue") #r1
  lines(c(as.Date(x[[1]]),as.Date(x[[2]]) ), c( x[8] , x[8] ) , col="darkgreen") #r2
} )

```

Current values

```{r,pivy2e}

piv <- pivots_yearly(price)
tail(piv) %>% kbl(caption="Recent price action and yearly pivots") %>%  kable_styling()

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
