---
title: "A BTC risk index for measuring the temperature of the market"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 10
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/risk_index.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

Prominent YouTuber [Benjamin Cowan](https://www.youtube.com/channel/UCRvqjQPSeaWn-uEx-w0XOIg) 
regularly discusses the BTC risk metric on his channel.
It is based on a ratio between a relatively short moving average (50 day) and a longer one (50 week),
although people implement this in different ways.

For some background information read [here](https://www.reddit.com/r/Bitcoin/comments/mgr5kq/im_searching_for_the_glorious_risk_metric_formula/).

A strategy used by some Bitcoiners is to accumulate when the risk metric is low and sell 
when the risk metric is high.

Here I'm going to calculate these moving averages and use them to calculate my own risk index.

This document is designed to be updated daily, so revisit whenever you need.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library("anytime")
  library("forecast")
})

```

Reminder: this is not financial advice.

## Get data

Historical data pre 2021 are from https://www.investing.com/indices/investing.com-btc-usd-historical-data
while the recent data are from http://coinmarketcap.com/

```{r,historicaldata}

daily <- read.csv("BitcoinHistoricalData_daily.csv",stringsAsFactors=FALSE)
daily <- daily[nrow(daily):1,]
daily <- apply(daily,2,function(x) { gsub(",","",x) } )

daily <- as.data.frame(daily,stringsAsFactors=FALSE)

daily$Price <- as.numeric(daily$Price)
daily$Open <- as.numeric(daily$Open)
daily$High <- as.numeric(daily$High)
daily$Low <- as.numeric(daily$Low)

daily$Date <- anydate(daily$Date)

daily <- daily[,c("Date","High","Low","Open","Price")]
colnames(daily) <- c("Date","High","Low","Open","Close")
daily <- daily[grep("2021",daily$Date,invert=TRUE),]

```

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2021-01-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat_daily.txt")
btcdat <- fromJSON("btcdat_daily.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$open, price$quote$USD$close, stringsAsFactors=FALSE)
colnames(price) <- c("Date","High","Low","Open","Close")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)

price$Date <- anydate(price$Date)

daily <- rbind(daily,price)

```

## Calculate moving averages, ratio and risk index

The first step is to calculate the 50d and 50w MAs as well as the ratio.

The MA ratio oscillates but the peaks are getting lower over time.

The upper line probably follows a logarithmic decay model but I will cheat and use a linear approximation.

The key dates I'm using as the tops are:

2013-04-10 266.00

2013-12-04 1237.6

2017-12-17 19065.70

The risk index can be calculated between 0 and 1 by the distance between the "floor" and what is considered the current top.

```{r,ma1}

daily$MA50d <- SMA(daily$Close,n=50)
daily$MA50w <- SMA(daily$Close,n=350)

daily$ratio <- daily$MA50d / daily$MA50w

tail(daily) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

ratio<-daily[which(!is.na(daily$ratio)),"ratio"]
rat_min <- min(ratio)

# tops
topdates <- as.Date(c("2013-04-10","2013-12-04","2017-12-17"))
tops <- daily[which(daily$Date %in% topdates),]

mylm <- lm(tops$ratio ~ tops$Date)

coeffs <- coefficients(mylm)

# generate points
x <- seq(from=daily$Date[1],
  to=daily$Date[nrow(daily)],
  length.out=nrow(daily) )

y <- predict(mylm,newdata=list(x=x), interval="confidence")

today_topval <- coeffs[1] + coeffs[2]*as.numeric(Sys.Date())

daily$topvals <- coeffs[1] + coeffs[2]*as.numeric(daily$Date)

daily$floor <- rat_min

daily$risk_index <- ( daily$ratio - daily$floor ) / daily$topvals

daily$risk_index <- (daily$ratio - daily$floor) / (daily$topvals - daily$floor)

```

## Charts

Now that everything is calculated, plot the results.

```{r,plot1}

plot(daily$Date,daily$Close,
 xlab="", ylab="USD", main="BTC price" ,log="y",pch=19,cex=0.1)
mtext("blue=50d MA, res=50w MA")
grid()
lines(daily$Date,daily$MA50w, col="blue",lwd=2)
lines(daily$Date,daily$MA50d, col="red",lwd=2)

plot(daily$Date,daily$ratio,ylim=c(0.4,4),
 xlab="", ylab="ratio", main="50d MA / 50w MA" ,pch=19,cex=0.1)
mtext("blue line=floor,red line=top")
grid()
abline(h=rat_min,col="blue",lwd=2)
abline(mylm, col="red",lwd=2)
#points(Sys.Date(),today_topval,cex=2)

plot(daily$Date,daily$risk_index, ylim=c(0,1.2),
 xlab="", ylab="Index", main="Risk index" ,pch=19,cex=0.1)
grid()
abline(h=1,col="red",lwd=2)

```

## Indicator performance in 2021

Let's take a look at how this indicator performed in 2021.

It is impossible to precisely time the market top, and based on this chart it would be best to scale out gradually when the index
exceeds 1.05.

If we had to pick a single day to dump, that would be the day when the index dips below 1.05 after a strong rally.
This date is indicated on the chart below

```{r,2021test}

daily2021 <- daily[grep("2021",daily$Date),]
daily2021 <- daily2021[which(daily2021$risk_index>0.95),]
peak <- daily2021[tail(which(daily2021$risk_index>1.05),1)+1,]


par(mfrow=c(2,1))
plot(daily2021$Date,daily2021$Close,type="l",xlab="",ylab="USD",main="2021 peak price")
grid()
abline(v=peak$Date,lty=2)

plot(daily2021$Date,daily2021$risk_index,type="l",xlab="",ylab="Index",main="BTC risk")
grid()
abline(v=peak$Date,lty=2)

```

Here is the data for that day.

```{r,2021peak}

peak

```

## Current BTC risk index

Now we can see the figures for the past 10 days to see how hot/cold the market is getting.

```{r,tbl1}

tail(daily,10) %>% kbl() %>% kable_paper("hover", full_width = F)

```

As you can see, according to this definition, 

This is just the data and indicators based on the 2021 peak, it may not turn out like this in future.
As the old saying goes, "past performance does not predict future results", however the market does seem to have repeating
boom and bust cycles.

This is just one indicator and should not be used in isolation.
There are other indicators of market tops I've listed below:

* https://charts.woobull.com/

* https://www.lookintobitcoin.com/charts/

* https://stats.buybitcoinworldwide.com/long-term-power-law/

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
