---
title: "A BTC risk index for measuring the temperature of the market"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 8
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/risk_index.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.


XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

Prominent YouTuber [Benjamin Cowan](https://www.youtube.com/channel/UCRvqjQPSeaWn-uEx-w0XOIg) 
regularly discusses the BTC risk metric on his channel.
It is based on a ratio between a relatively short moving average (50 day) and a longer one (50 week),
although people implement this in different ways.

For some background information read [here](https://www.reddit.com/r/Bitcoin/comments/mgr5kq/im_searching_for_the_glorious_risk_metric_formula/).

A strategy used by some Bitcoiners is to accumulate when the risk metric is low and sell 
when the risk metric is high.

Here I'm going to calculate these moving averages and use them to calculate my own risk index.

This document is designed to be updated daily, so revisit whenever you need.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library("anytime")
  library("forecast")
  library("plotly")
})

```

Reminder: this is not financial advice.

## Get data

Historical data pre 2021 are from https://www.investing.com/indices/investing.com-btc-usd-historical-data
while the recent data are from http://coinmarketcap.com/

```{r,historicaldata}

daily <- read.csv("BitcoinHistoricalData_daily.csv",stringsAsFactors=FALSE)
daily <- daily[nrow(daily):1,]
daily <- apply(daily,2,function(x) { gsub(",","",x) } )

daily <- as.data.frame(daily,stringsAsFactors=FALSE)

daily$Price <- as.numeric(daily$Price)
daily$Open <- as.numeric(daily$Open)
daily$High <- as.numeric(daily$High)
daily$Low <- as.numeric(daily$Low)

daily$Date <- anydate(daily$Date)

daily <- daily[,c("Date","High","Low","Open","Price")]
colnames(daily) <- c("Date","High","Low","Open","Close")
daily <- daily[grep("2021",daily$Date,invert=TRUE),]

```

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2021-01-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat_daily.txt")
btcdat <- fromJSON("btcdat_daily.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$open, price$quote$USD$close, stringsAsFactors=FALSE)
colnames(price) <- c("Date","High","Low","Open","Close")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)

price$Date <- anydate(price$Date)

daily <- rbind(daily,price)

```

## Calculate moving averages, ratio and risk index

The first step is to calculate the 50d and 50w MAs as well as the ratio.

The MA ratio oscillates but the peaks are getting lower over time.

The upper line probably follows a logarithmic decay model but I will cheat and use a linear approximation.

The key dates I'm using as the tops are:

2013-04-10 266.00

2013-12-04 1237.6

2017-12-17 19065.70

The risk index can be calculated between 0 and 1 by the distance between the "floor" and what is considered the current top.

```{r,ma1}

daily$MA50d <- SMA(daily$Close,n=50)
daily$MA50w <- SMA(daily$Close,n=350)

daily$ratio <- daily$MA50d / daily$MA50w

tail(daily) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

ratio<-daily[which(!is.na(daily$ratio)),"ratio"]
rat_min <- min(ratio)

# tops
topdates <- as.Date(c("2013-04-10","2013-12-04","2017-12-17"))
tops <- daily[which(daily$Date %in% topdates),]

mylm <- lm(tops$ratio ~ tops$Date)

coeffs <- coefficients(mylm)

# generate points
x <- seq(from=daily$Date[1],
  to=daily$Date[nrow(daily)],
  length.out=nrow(daily) )

y <- predict(mylm,newdata=list(x=x), interval="confidence")

today_topval <- coeffs[1] + coeffs[2]*as.numeric(Sys.Date())

daily$topvals <- coeffs[1] + coeffs[2]*as.numeric(daily$Date)

daily$floor <- rat_min

daily$risk_index <- ( daily$ratio - daily$floor ) / daily$topvals

daily$risk_index <- (daily$ratio - daily$floor) / (daily$topvals - daily$floor)

```

## Charts

Here's an interactive chart of price action and MA50d and MA50w.
You can hover over points to see their details as well zoom in by selecting a rectangle.

```{r,plotly1,warning=FALSE}

pal <- c("blue", "red")
fig <- plot_ly(
  daily, type = 'scatter', mode= "markers",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>MA50d:', round(MA50d,0.1),
    '$<br>MA50w:', round(MA50w,0.1),  
    '$<br>ratio:', signif(ratio,3),
    '$<br>Risk Index:', signif(risk_index,3) ),
  colors = pal,
  color = ~risk_index  ) %>%
  add_trace(x = ~Date, y = ~Close, name= "price" ) %>%
  layout(yaxis = list(type="log"),
  xaxis = list(range = c(head(daily,1)$Date,tail(daily,1)$Date)))
fig <- fig %>% add_trace(type='scatter',x= ~Date, y = ~MA50d,
  name = 'MA50d', mode = 'lines', line = list(color = 'orange', width = 4))
fig <- fig %>% add_trace(type='scatter',x= ~Date, y = ~MA50w,
  name = 'MA50w', mode = 'lines', line = list(color = 'black', width = 4))
fig <- fig %>% layout(title = "BTC price history with MA50d and MA50w",
         xaxis = list(title = "Date"),
         yaxis = list(title = "USD price"))
fig

```

This plot shows how the ratio peaks are diminishing over time.

```{r,plot1}

plot(daily$Date,daily$ratio,ylim=c(0.4,4),
 xlab="", ylab="ratio", main="50d MA / 50w MA" ,pch=19,cex=0.1)
mtext("blue line=floor,red line=top")
grid()
abline(h=rat_min,col="blue",lwd=2)
abline(mylm, col="red",lwd=2)

```


```{r,plotly2,warning=FALSE}

pal <- c("blue", "red")
fig <- plot_ly(
  daily, type = 'scatter', mode= "markers",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>MA50d:', round(MA50d,0.1),
    '$<br>MA50w:', round(MA50w,0.1),
    '$<br>ratio:', signif(ratio,3),
    '$<br>Risk Index:', signif(risk_index,3) ),
  colors = pal,
  color = ~risk_index , marker = list(size = 10) ) %>%
  add_trace(x = ~Date, y = ~risk_index, name= "risk index" ) %>%
  layout( yaxis = list(range = c(0,1.2))     ,  
    xaxis = list(range = c(head(daily,1)$Date,tail(daily,1)$Date)))
fig <- fig %>% layout(title = "BTC risk index",
         xaxis = list(title = "Date"),
         yaxis = list(title = "Index"))
fig

```

## Current BTC risk index values

Now we can see the figures for the past 10 days to see how hot/cold the market is getting.

```{r,tbl1}

tail(daily) %>% kbl() %>% kable_paper("hover", full_width = F)

```

## Final notes

This is just a model of the historical data, it may not turn out like this in future.
As the old saying goes, "past performance does not predict future results", 
however the market does seem to have repeating boom and bust cycles.

This is just one indicator and should not be used in isolation.
There are other indicators of market tops I've listed below:

* https://charts.woobull.com/

* https://www.lookintobitcoin.com/charts/

* https://stats.buybitcoinworldwide.com/long-term-power-law/

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
