---
title: "Establishing a BTC risk metric"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 10
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/risk_metric.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

Prominent YouTuber [Benjamin Cowan](https://www.youtube.com/channel/UCRvqjQPSeaWn-uEx-w0XOIg) 
regularly discusses the BTC risk metric on his channel.
It is based on a ratio between a relatively short moving average (50 day) and a longer one (50 week),
although people implement this in different ways.

A strategy used by some Bitcoiners is to accumulate when the risk metric is low and sell 
when the risk metric is high.

For some background information read [here](https://www.reddit.com/r/Bitcoin/comments/mgr5kq/im_searching_for_the_glorious_risk_metric_formula/).

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library("anytime")
  library("forecast")
})

```

Reminder: this is not financial advice.

## Get data

Here I will be analysing the weekly trading range as a proxy for volatility.
Historical data pre 2021 are from https://www.investing.com/indices/investing.com-btc-usd-historical-data
while the recent data are from http://coinmarketcap.com/

This document is designed to be updated daily.

```{r,historicaldata}

daily <- read.csv("BitcoinHistoricalData_daily.csv",stringsAsFactors=FALSE)
daily <- daily[nrow(daily):1,]
daily <- apply(daily,2,function(x) { gsub(",","",x) } )

daily <- as.data.frame(daily,stringsAsFactors=FALSE)

daily$Price <- as.numeric(daily$Price)
daily$Open <- as.numeric(daily$Open)
daily$High <- as.numeric(daily$High)
daily$Low <- as.numeric(daily$Low)

daily$Date <- anydate(daily$Date)

daily <- daily[,c("Date","High","Low","Open","Price")]
colnames(daily) <- c("Date","High","Low","Open","Close")
daily <- daily[grep("2021",daily$Date,invert=TRUE),]

```

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2021-01-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat_daily.txt")
btcdat <- fromJSON("btcdat_daily.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$open, price$quote$USD$close, stringsAsFactors=FALSE)
colnames(price) <- c("Date","High","Low","Open","Close")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)

price$Date <- anydate(price$Date)

daily <- rbind(daily,price)

```

## Calculate moving averages, ratio and risk index

The first step is to calculate the 50d and 50w MAs as well as the ratio.

The MA ratio oscillates but the peaks are getting lower over time.

The upper line probably follows a logarithmic decay model but I will cheat and use a linear approximation.

The key dates I'm using as the tops are:

2013-04-10 266.00

2013-12-04 1237.6

2017-12-17 19065.70

The risk index can be calculated between 0 and 1 by the distance between the "floor" and what is considered the current top.

```{r,ma1}

daily$MA50d <- SMA(daily$Close,n=50)
daily$MA50w <- SMA(daily$Close,n=350)

daily$ratio <- daily$MA50d / daily$MA50w

tail(daily) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

ratio<-daily[which(!is.na(daily$ratio)),"ratio"]
rat_min <- min(ratio)

# tops
topdates <- as.Date(c("2013-04-10","2013-12-04","2017-12-17"))
tops <- daily[which(daily$Date %in% topdates),]

mylm <- lm(tops$ratio ~ tops$Date)

coeffs <- coefficients(mylm)

# generate points
x <- seq(from=daily$Date[1],
  to=daily$Date[nrow(daily)],
  length.out=nrow(daily) )

y <- predict(mylm,newdata=list(x=x), interval="confidence")

today_topval <- coeffs[1] + coeffs[2]*as.numeric(Sys.Date())

daily$topvals <- coeffs[1] + coeffs[2]*as.numeric(daily$Date)

daily$floor <- rat_min

daily$risk_index <- ( daily$ratio - daily$floor ) / daily$topvals

daily$risk_index <- (daily$ratio - daily$floor) / (daily$topvals - daily$floor)

```

## Charts

Now that everything is calculated, plot the results.

```{r,plot1}

plot(daily$Date,daily$Close,
 xlab="", ylab="USD", main="BTC price" ,log="y",pch=19,cex=0.1)
mtext("blue=50d MA, res=50w MA")
grid()
lines(daily$Date,daily$MA50w, col="blue",lwd=2)
lines(daily$Date,daily$MA50d, col="red",lwd=2)

plot(daily$Date,daily$ratio,ylim=c(0.4,4),
 xlab="", ylab="ratio", main="50d MA / 50w MA" ,pch=19,cex=0.1)
mtext("blue line=floor,red line=top")
grid()
abline(h=rat_min,col="blue",lwd=2)
abline(mylm, col="red",lwd=2)
#points(Sys.Date(),today_topval,cex=2)

plot(daily$Date,daily$risk_index, ylim=c(0,1.2),
 xlab="", ylab="Index", main="Risk index" ,pch=19,cex=0.1)
grid()
abline(h=1,col="red",lwd=2)

```

You can see that the peaks cross the red line and continue to go above it.
Why is that you ask?
Well the 50d MA is still a lagging indicator and the model I made is generated from the actual top days, not the day when the 50D MA peaked.
If you wait for the index to hit 1.1, you'll likely miss the party by 50 days!

Now we can see the figures for the past 10 days to see how hot/cold the market is getting.

```{r,tbl1}

tail(daily,10) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
