---
title: "Watching the weekly RSI for BTC, ETH and XMR"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 8
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/rsi_watch.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.


XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

In this blog post I will be watching the weekly, biweekly and monthly RSI for BTC and ETH which can be 
used to spot good buying or selling opportunities.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("parallel")
  library("kableExtra")
  library("anytime")
  library("plotly")
  library("gridExtra")
})

```

Reminder: this is not financial advice.

## Get BTC data

Obtaining BTC historical data (weekly).
Historical data pre 2021 are from https://www.investing.com/indices/investing.com-btc-usd-historical-data
while the recent data are from http://coinmarketcap.com/

```{r,historicaldata1}

weekly <- read.csv("BitcoinHistoricalData_weekly.csv",stringsAsFactors=FALSE)
weekly <- weekly[nrow(weekly):1,]
weekly <- apply(weekly,2,function(x) { gsub(",","",x) } )
weekly <- as.data.frame(weekly,stringsAsFactors=FALSE)
weekly$Close <- as.numeric(weekly$Price)
weekly$Date <- anydate(weekly$Date)
weekly <- weekly[,c("Date","Close")]
weekly <- weekly[grep("2021",weekly$Date,invert=TRUE),]
weekly <- weekly[grep("2022",weekly$Date,invert=TRUE),]

```

```{r,getdata1}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=weekly&time_start=2021-01-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat_weekly.txt")
btcdat <- fromJSON("btcdat_weekly.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$close ,stringsAsFactors=FALSE)
colnames(price) <- c("Date","Close")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)
price <- rbind(weekly,price)

```

## RSI on weekly and biweekly

These can be used to spot market cycle tops and bottoms.

Firstly to the weekly RSI.

```{r,rsi1btc}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

```

These charts are interactive, so you can hover, zoom, pan, etc.

```{r,rsi1btc2}
hline <- function(y = 0, color = "black") {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = color, dash="dot")
  )
}

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=35,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>% 
  layout(title = list(text = "BTC weekly RSI"))

fig

```

Biweekly RSI.

```{r,rsi2btc}

if ( nrow(price) %% 2 == 0 ) {
  idx <- seq(2,nrow(price),2)
} else {
  idx <- seq(1,nrow(price),2)
}

price <- price[idx,]

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=40,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "BTC biweekly RSI"))

fig

```

## Get monthly data

Obtaining BTC historical data (monthly).
Historical data pre 2021 are from https://www.investing.com/indices/investing.com-btc-usd-historical-data
while the recent data are from http://coinmarketcap.com/

```{r,historicaldata2}

monthly <- read.csv("BitcoinHistoricalData_monthly.csv",stringsAsFactors=FALSE)
monthly <- monthly[nrow(monthly):1,]
monthly <- apply(monthly,2,function(x) { gsub(",","",x) } )
monthly <- as.data.frame(monthly,stringsAsFactors=FALSE)
monthly$Close <- as.numeric(monthly$Price)

y <- paste("20",sapply(strsplit(monthly$Date," "),"[[",2),sep="")
m <- sapply(strsplit(monthly$Date," "),"[[",1)
monthly$Date <- anydate(paste(y,m))

monthly <- monthly[,c("Date","Close")]
monthly <- monthly[grep("2021",monthly$Date,invert=TRUE),]
monthly <- monthly[grep("2022",monthly$Date,invert=TRUE),]

```

```{r,getdata2}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=monthly&time_start=2020-12-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat_monthly.txt")
btcdat <- fromJSON("btcdat_monthly.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$close ,stringsAsFactors=FALSE)
colnames(price) <- c("Date","Close")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)
price$Date <- anydate(price$Date)
price <- rbind(monthly,price)

```

## RSI on the monthly

These can be used to spot market cycle tops.

```{r,rsi3btc}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=45,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "BTC monthly RSI"))

fig

```

## Get ETH data

Obtaining ETH historical data (daily).

```{r,getdataeth}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=weekly&time_start=2015-08-01&time_end=",mydate,sep="")
download.file(URL,destfile="ethdat.txt")
dat <- fromJSON("ethdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$close ,stringsAsFactors=FALSE)
colnames(price) <- c("Date","Close")
price$Date <- as.Date(sapply(strsplit(as.character(price$Date),"T"),"[[",1))


```

## RSI on weekly and biweekly

These can be used to spot market cycle tops.

Firstly to the weekly RSI.

```{r,rsi1eth}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

```

These charts are interactive, so you can hover, zoom, pan, etc.  

```{r,rsi1eth2}

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=30,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "ETH weekly RSI"))

fig

```

Biweekly RSI.

```{r,rsi2eth}

if ( nrow(price) %% 2 == 0 ) {
  idx <- seq(2,nrow(price),2)
} else {
  idx <- seq(1,nrow(price),2)
}

price <- price[idx,]

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=40,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "ETH biweekly RSI"))

fig

```

## Get monthly data

```{r,getdata2eth}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=monthly&time_start=2015-08-01&time_end=",mydate,sep="")
download.file(URL,destfile="ethdat.txt")
dat <- fromJSON("ethdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$close ,stringsAsFactors=FALSE)
colnames(price) <- c("Date","Close")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)
price$Date <- anydate(price$Date)


```

## RSI on the monthly

These can be used to spot market cycle tops.

```{r,rsi3eth}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=45,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "ETH monthly RSI"))

fig

```

## Get XMR data

Obtaining XMR historical data (daily).

```{r,getdataxmr}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=XMR&convert=USD&interval=weekly&time_start=2015-08-01&time_end=",mydate,sep="")
download.file(URL,destfile="xmrdat.txt")
dat <- fromJSON("xmrdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$close ,stringsAsFactors=FALSE)
colnames(price) <- c("Date","Close")
price$Date <- as.Date(sapply(strsplit(as.character(price$Date),"T"),"[[",1))


```

## RSI on weekly and biweekly

These can be used to spot market cycle tops.

Firstly to the weekly RSI.

```{r,rsi1xmr}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

```

These charts are interactive, so you can hover, zoom, pan, etc.  

```{r,rsi1xmr2}

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=30,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "XMR weekly RSI"))

fig

```

Biweekly RSI.

```{r,rsi2xmr}

if ( nrow(price) %% 2 == 0 ) {
  idx <- seq(2,nrow(price),2)
} else {
  idx <- seq(1,nrow(price),2)
}

price <- price[idx,]

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=40,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "XMR biweekly RSI"))

fig

```

## Get monthly data

```{r,getdata2xmr}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=XMR&convert=USD&interval=monthly&time_start=2015-08-01&time_end=",mydate,sep="")
download.file(URL,destfile="xmrdat.txt")
dat <- fromJSON("xmrdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$close ,stringsAsFactors=FALSE)
colnames(price) <- c("Date","Close")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)
price$Date <- anydate(price$Date)


```

## RSI on the monthly

These can be used to spot market cycle tops.

```{r,rsi3xmr}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

fig1 <- plot_ly(price, x= ~Date, y= ~Close,
  type="scatter", mode="lines", name="log USD price",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(xaxis = list(title="Date"),yaxis = list(type="log",title="log USD price"))

fig2 <- plot_ly(price, x= ~Date, y= ~rsi,
  type="scatter", mode="lines", name="RSI",
  text = ~paste("Price: ", round(Close,0.1),
    '$<br>RSI:', round(rsi,0.1) )) %>%
  layout(shapes = list(hline(y=90,color="red"),hline(y=45,color="blue") ) ,
    xaxis=list(title="Date"),yaxis=list("RSI"))

fig <- subplot(fig1, fig2, nrows = 2, shareX=TRUE) %>%
  layout(title = list(text = "XMR monthly RSI"))

fig

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
