---
title: "Watching the weekly RSI"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/ma_trading_weekly.Rmd

If you find this resource useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

In this blog post I will be watching the weekly, biweekly and monthly RSI which can be used to spot 
good buying or selling opportunities.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("parallel")
  library("kableExtra")
})

```

Reminder: this is not financial advice.

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=weekly&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")
dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## RSI on weekly and biweekly

These can be used to spot market cycle tops.

Firstly to the weekly RSI.

```{r,rsi1,fig.height=10,fig.width=8}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

par(mfrow=c(2,1))
plot(price$Close ~ as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price$rsi ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="RSI",ylim=c(20,100),main="weekly RSI")
grid()
abline(h=90,col="red")
abline(h=30,col="black")

```

Biweekly RSI.

```{r,rsi2,fig.height=10,fig.width=8}

if ( nrow(price) %% 2 == 0 ) {
  idx <- seq(2,nrow(price),2)
} else {
  idx <- seq(1,nrow(price),2)
}

price <- price[idx,]

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

par(mfrow=c(2,1))
plot(price$Close ~ as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price$rsi ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="RSI",ylim=c(20,100),main="biweekly RSI")
grid()
abline(h=90,col="red")
abline(h=30,col="black")

```

## Get monthly data

```{r,getdata2}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=monthly&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")
dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","Close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## RSI on the monthly

These can be used to spot market cycle tops.

```{r,rsi3,fig.height=10,fig.width=8}

price$rsi <- RSI(price$Close,n=14)

tail(price) %>% kbl() %>% kable_paper("hover", full_width = F)

par(mfrow=c(2,1))
plot(price$Close ~ as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price$rsi ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="RSI",ylim=c(20,100),main="monthly RSI")
grid()
abline(h=90,col="red")
abline(h=50,col="black")

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
