---
title: "<br>Trading based on RSI"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
---

## Intro

In this blog post I will be trilaing using RSI for entry and exit.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")

knitr::opts_chunk$set(fig.width=8, fig.height=8) 

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

if ( ! file.exists("btcdat.txt") ) {
  download.file(URL,destfile="btcdat.txt")
}

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl


```


## RSI analysis



FEES: Assuming 0.1% trading fee, as per binance current rate.

The idea here is to enter after RSI recovers from a low point and then recovers to positive territory.

Exit when RSI stops increasing


```{r,rsi}

n=14
window=5

rsi <- RSI(price$close,n=14)

change <- unlist(sapply(window:length(rsi) , function(x) {
  return( rsi[x] - rsi[(x-5)]) 
}))

change <- c(rep(NA,nrow(price) - length(change)), change)

price <- data.frame(price,rsi,change,stringsAsFactors=FALSE)

price$low <- change < -20
price$high <- change > 20

par(mfrow=c(2,1))
plot(price$close~ as.Date(price$date),type="l",log="y",xlab="",ylab="USD price")
grid()
plot(price$rsi ~ as.Date(price$date),type="l",xlab="",ylab="RSI")
grid()
abline(h=30,col="red")
abline(h=70,col="red")

```





```{r,rsisma}

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","close")
price$date <- sapply(strsplit(price$date,"T"),"[[",1)


rsisma <- function(price,n_ma,n_rsi,fee) {
rsi <- RSI(Cl(price),n=n_rsi)
rsima <- SMA(rsi,n=n_ma)
price2 <- data.frame(price,rsi,rsima,stringsAsFactors=FALSE)
price2$higher <- price2$rsi > price2$rsima
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(prod(trades$diff))
}


```

## Exponential moving average function


```{r,ema}

rsiema <- function(price,n_ma,n_rsi,fee) {
rsi <- RSI(Cl(price),n=n_rsi)
rsima <- EMA(rsi,n=n_ma)
price2 <- data.frame(price,rsi,rsima,stringsAsFactors=FALSE)
price2$higher <- price2$rsi > price2$rsima
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(prod(trades$diff))
}


```

## Backtesting

```{r,backtest}

fee=0.002
n_rsi=14

res_sma <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=n_rsi,fee) } ) )

res_ema <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=n_rsi,fee) } ) )

```

## Plots

```{r,plots,fig.height=8,fig.width=8}

plot(res_sma,pch=19,xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(res_sma,pch=19,xlim=c(70,120),xlab="moving average interval (days)",ylab="fold return",main="SMA")
abline(h=hodl,col="red")
grid()

plot(res_ema,pch=19,xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

plot(res_ema,pch=19,xlim=c(70,120),xlab="moving average interval (days)",ylab="fold return",main="EMA")
abline(h=hodl,col="red")
grid()

```

## RSI optimisation 1

```{r,rsiopt1}

res_sma_11 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=11,fee) } ) )
res_sma_14 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=14,fee) } ) )
res_sma_17 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=17,fee) } ) )
res_sma_20 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=20,fee) } ) )
max_sma <- max(c(res_sma_11,res_sma_14,res_sma_17,res_sma_20))

res_ema_11 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=11,fee) } ) )
res_ema_14 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=14,fee) } ) )
res_ema_17 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=17,fee) } ) )
res_ema_20 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=20,fee) } ) )
max_ema <- max(c(res_ema_11,res_ema_14,res_ema_17,res_ema_20))

#plot sma
plot(res_sma_11,pch=19,xlab="moving average interval (days)",ylab="fold return",
  main="RSI SMA",ylim=c(0,max_sma),col="blue")
points(res_sma_14,pch=19,col="black")
points(res_sma_17,pch=19,col="red")
points(res_sma_20,pch=19,col="green")
legend("topright", 
  legend = c("11", "14", "17", "20"), 
  col = c("blue","black","red","green"), 
  pch = 19, 
  bty = "y", 
  pt.cex = 2, 
  cex = 1.2, 
  text.col = "black", 
  horiz = F )
abline(h=hodl,col="red")
grid()

#zoom
plot(res_sma_11,pch=19,xlab="moving average interval (days)",ylab="fold return", 
  main="RSI SMA",ylim=c(0,max_sma), xlim=c(70,120),col="blue")
points(res_sma_14,pch=19,col="black")
points(res_sma_17,pch=19,col="red")
points(res_sma_20,pch=19,col="green")
legend("bottomleft",
  legend = c("11", "14", "17", "20"),
  col = c("blue","black","red","green"),
  pch = 19,
  bty = "y", 
  pt.cex = 2, 
  cex = 1.2,
  text.col = "black",
  horiz = F )
abline(h=hodl,col="red")
grid()


#plot ema
plot(res_ema_11,pch=19,xlab="moving average interval (days)",ylab="fold return",
  main="RSI EMA",ylim=c(0,max_ema),col="blue")
points(res_ema_14,pch=19,col="black")
points(res_ema_17,pch=19,col="red")
points(res_ema_20,pch=19,col="green")
legend("topright",
  legend = c("11", "14", "17", "20"),
  col = c("blue","black","red","green"),
  pch = 19,
  bty = "y", 
  pt.cex = 2, 
  cex = 1.2,
  text.col = "black",
  horiz = F )
abline(h=hodl,col="red")
grid()

```

## RSI optimisation 2

```{r,rsiopt2}

res_sma_12 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=12,fee) } ) )
res_sma_13 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=13,fee) } ) )
res_sma_14 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=14,fee) } ) )
res_sma_15 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=15,fee) } ) )
res_sma_16 <- c(0,0,sapply(3:200,function(n) { rsisma(price,n_ma=n,n_rsi=16,fee) } ) )
max_sma <- max(c(res_sma_12,res_sma_13,res_sma_14,res_sma_15,res_sma_16))
max_sma
res_ema_12 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=12,fee) } ) )
res_ema_13 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=13,fee) } ) )
res_ema_14 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=14,fee) } ) )
res_ema_15 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=15,fee) } ) )
res_ema_16 <- c(0,0,sapply(3:200,function(n) { rsiema(price,n_ma=n,n_rsi=16,fee) } ) )
max_ema <- max(c(res_ema_12,res_ema_13,res_ema_14,res_ema_15,res_ema_16))

#plot sma
plot(res_sma_12,pch=19,xlab="moving average interval (days)",ylab="fold return",
  main="RSI SMA",ylim=c(0,max_sma),col="blue")
points(res_sma_13,pch=19,col="black")
points(res_sma_14,pch=19,col="red")
points(res_sma_15,pch=19,col="green")
points(res_sma_16,pch=19,col="gray")

legend("topright",
  legend = c("12", "13", "14", "15", "16"),
  col = c("blue","black","red","green", "gray"),
  pch = 19,
  bty = "y",
  pt.cex = 2,
  cex = 1.2,
  text.col = "black",
  horiz = F )
abline(h=hodl,col="red")
grid()

#zoom
plot(res_sma_12,pch=19,xlab="moving average interval (days)",ylab="fold return",
  main="RSI SMA",ylim=c(0,max_sma),xlim=c(80,120),col="blue")
points(res_sma_13,pch=19,col="black")
points(res_sma_14,pch=19,col="red")
points(res_sma_15,pch=19,col="green")
points(res_sma_16,pch=19,col="gray")

legend("bottomleft",
  legend = c("12", "13", "14", "15", "16"),
  col = c("blue","black","red","green", "gray"),
  pch = 19,
  bty = "y",
  pt.cex = 2,
  cex = 1.2,
  text.col = "black",
  horiz = F )
abline(h=hodl,col="red")
grid()

#plot ema
plot(res_ema_12,pch=19,xlab="moving average interval (days)",ylab="fold return",
  main="RSI EMA",ylim=c(0,max_ema),col="blue")
points(res_ema_13,pch=19,col="black")
points(res_ema_14,pch=19,col="red")
points(res_ema_15,pch=19,col="green")
points(res_ema_16,pch=19,col="gray")

legend("topright",
  legend = c("12", "13", "14", "15", "16"),
  col = c("blue","black","red","green", "gray"),
  pch = 19,
  bty = "y",
  pt.cex = 2,
  cex = 1.2,
  text.col = "black",
  horiz = F )
abline(h=hodl,col="red")
grid()

#zoom
plot(res_ema_12,pch=19,xlab="moving average interval (days)",ylab="fold return",
  main="RSI EMA",ylim=c(0,max_ema),xlim=c(80,120),col="blue")
points(res_ema_13,pch=19,col="black")
points(res_ema_14,pch=19,col="red")
points(res_ema_15,pch=19,col="green")
points(res_ema_16,pch=19,col="gray")

legend("bottomleft",
  legend = c("12", "13", "14", "15", "16"),
  col = c("blue","black","red","green", "gray"),
  pch = 19,
  bty = "y",
  pt.cex = 2,
  cex = 1.2,
  text.col = "black",
  horiz = F )
abline(h=hodl,col="red")
grid()

```



## Show results for best parameters

These data show RSI EMA (14d) is optimum.

Will do same SMA for comparison purposes.

```{r,best_sma,fig.width=10,fig.height=8}

rsisma2 <- function(price,n_ma,n_rsi,fee) {
rsi <- RSI(Cl(price),n=n_rsi)
rsima <- SMA(rsi,n=n_ma)
price2 <- data.frame(price,rsi,rsima,stringsAsFactors=FALSE)
price2$higher <- price2$rsi > price2$rsima
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(trades)
}

rsiema2 <- function(price,n_ma,n_rsi,fee) {
rsi <- RSI(Cl(price),n=n_rsi)
rsima <- EMA(rsi,n=n_ma)
price2 <- data.frame(price,rsi,rsima,stringsAsFactors=FALSE)
price2$higher <- price2$rsi > price2$rsima
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
return(trades)
}


best_sma_14 = which(res_sma_14==max(res_sma_14))
best_sma_14
best_sma_14_trades <- rsisma2(price,n_ma=best_sma_14,n_rsi=14,fee)
prod(prod(best_sma_14_trades$diff))
nrow(best_sma_14_trades)


par(mfrow=c(2,1))
SMA_HEADER=paste(best_sma_14,"day RSI cross simple moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()

rsi <- RSI(Cl(price),n=14)
rsima <- SMA(rsi,n=best_sma_14)

plot(rsi~as.Date(price$date),type="l",
  xlab="Date",ylab="RSI",main=SMA_HEADER)
grid()
lines(rsima~ as.Date(price$date) ,col="red")


best_ema_14 = which(res_ema_14==max(res_ema_14))
best_ema_14
best_ema_14_trades <- rsiema2(price,n_ma=best_ema_14,n_rsi=14,fee)
prod(prod(best_ema_14_trades$diff))
nrow(best_ema_14_trades)


par(mfrow=c(2,1))
EMA_HEADER=paste(best_ema_14,"day RSI cross exponential moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()

rsi <- RSI(Cl(price),n=14)
rsima <- EMA(rsi,n=best_ema_14)

plot(rsi~as.Date(price$date),type="l", 
  xlab="Date",ylab="RSI",main=EMA_HEADER)
grid()
lines(rsima~ as.Date(price$date) ,col="red")

```


## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)
best_sma <- best_sma_14
res_sma_fees <- sapply(fee_levels,function(x) { rsisma(price,best_sma, n_rsi=14 ,fee=x) } )
plot(fee_levels,res_sma_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

best_ema <- best_ema_14
res_ema_fees <- sapply(fee_levels,function(x) { rsiema(price,best_ema,n_rsi=14,fee=x) } )
plot(fee_levels,res_ema_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades <- best_sma_14_trades
sma_trades$year <- sapply(strsplit(sma_trades[,3],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,5] } )
sma_yearly <- sma_yearly[which(lapply(sma_yearly,length)>0)]
vioplot(sma_yearly)

yr <- sapply(sma_yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

ema_trades <- best_ema_14_trades
ema_trades$year <- sapply(strsplit(ema_trades[,3],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,5] } )
ema_yearly <- ema_yearly[which(lapply(ema_yearly,length)>0)]
vioplot(ema_yearly)

yr <- sapply(ema_yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```


## Conclusion

Hodling BTC over this period gave a 378 fold return.
The best simple moving average was 18 days, which gave a 1366 fold return (with fees at 0.1%).
The best exponential moving average was 16 days, which gave a return of 1680 fold.
The fee rate strongly affects the fold return, so it is definitely worthwhile to seek lower price exchanges.
Profitability varies strongly by year with bull market most profitable and bear market least profitable.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
