---
title: "Stablecoin dominance as a tool for timing the market top"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

## Intro

I had an idea that the share of stablecoins could be used to predict the end of a bull market, so I went to [tradingview](https://www.tradingview.com/symbols/CRYPTOCAP-USDT.D/) 
and yes it appears that a jump in stablecoin dominance began in mid 2020, peaking in September and then began dropping, coinciding with the bull run of January 2021.
The concept is that once stablecoin dominance returns to normal levels, the bull market will start to run out of steam due to falling inflows relative to profit takers.

```{r,lib}

library(jsonlite)
library(tidyverse)

```

## Get data

We will be simplifying things by working only with the market cap of the top two stablecoins Tether (USDT) and USD Coin (USDC) against cryptocurrencies Bitcoin (BTC) and Ethereum (ETH).





```{r,getdata}

mydate <- Sys.Date()-1

BTCURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(BTCURL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
head(btcdat$data$quotes)
btc_mc <- btcdat$data$quotes$quote$USD$market_cap

ETHURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(ETHURL,destfile="ethdat.txt")
ethdat <- fromJSON("ethdat.txt")
head(ethdat$data$quotes)
eth_mc <- ethdat$data$quotes$quote$USD$market_cap 

USDTURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=USDT&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(USDTURL,destfile="usdtdat.txt")
usdtdat <- fromJSON("usdtdat.txt")
head(usdtdat$data$quotes)
usdt_mc <- usdtdat$data$quotes$quote$USD$market_cap

USDCURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=USDC&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(USDCURL,destfile="usdcdat.txt")
usdcdat <- fromJSON("usdcdat.txt")
head(usdcdat$data$quotes)
usdc_mc <- usdcdat$data$quotes$quote$USD$market_cap

times_mc <- btcdat$data$quotes$quote$USD$timestamp
mc <- data.frame(times_mc,btc_mc,eth_mc,usdt_mc,usdc_mc)


```

## Now calculate stablecoin dominance


```{r,scdom}

mc$total <- mc$btc_mc + mc$eth_mc + mc$usdt_mc + mc$usdc_mc

mc$sc <- mc$usdt_mc + mc$usdc_mc

mc$scrat <- mc$sc / mc$total

mc$times_mc <- as.Date(sapply(strsplit(mc$times_mc,"T"),"[[",1))

p <- ggplot(mc, aes(x=times_mc, y=scrat)) +
  geom_line() + 
  xlab("") +
  geom_hline(yintercept=0.03, linetype="dashed", color = "red", size=2) +
  ylab("stablecoin dominance")
p


```

## When was maximum stablecoin dominance?

```{r,scmax}

mc[which(mc$scrat == max(mc$scrat) ),]

mc[which(mc$scrat == min(mc$scrat) ),]

```

## Calculate the linear model for stablecoin dominance since the peak

```{r,lm}

top <- mc[which(mc$scrat == max(mc$scrat) ),]

mc_pk <- mc[631:nrow(mc),]

p <- ggplot(mc_pk, aes(x=times_mc, y=scrat)) +
  geom_point() + 
  ylab("stablecoin dominance") +
  xlab("")
p

model <- lm(scrat ~ times_mc, data = mc_pk)

model

```

## When will the peak be reached at the current rate?

Assume the market peaks out when stablecoin dominance reduces to 3%

```{r,timing}

ggplot(mc_pk, aes(x=times_mc, y=scrat)) +
  scale_x_date(limit=c(as.Date("2020-09-01"),as.Date("2021-06-01"))) +
  ylim(0.025, 0.12) +
  geom_point() +
  xlab("") +
  ylab("stablecoin dominance") +
  geom_hline(yintercept=0.03, linetype="dashed", color = "red", size=2) +
  stat_smooth(method = lm,fullrange=TRUE)

```

This chart indicates the current bull run will start to run out of puff at the end of March.


## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
