---
title: "Stablecoin dominance: a tool for timing the market top?"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/stablecoincap.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.


XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

I had an idea that the share of stablecoins could be used to predict the end of a bull market, 
so I went to [tradingview](https://www.tradingview.com/symbols/CRYPTOCAP-USDT.D/) 
and yes it appears that a jump in stablecoin dominance began in mid 2020, peaking in September and then
began dropping, coinciding with the bull run of January 2021.
The concept is that once stablecoin dominance returns to normal levels, the bull market will start to run out of steam due to falling inflows relative to profit takers.

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
library(jsonlite)
library(tidyverse)
})

Sys.Date()

knitr::opts_chunk$set(fig.width=8, fig.height=8) 

```

Reminder: this is not financial advice.

## Get data

We will be simplifying things by working only with the market cap of the top two stablecoins Tether (USDT) and USD Coin (USDC) against cryptocurrencies Bitcoin (BTC) and Ethereum (ETH).

Here is a graph of the market cap of these assets since 2019.

As you can see in the second chart, stablecoin market cap is increasing dramatically.

This represents the phenomenon that stablecoins are becoming an important part of the crypto ecosystem.
This could be driven by utility as a medium of exchange without borders, as a vehicle for earning high interest, or simply a way to avoid currency debasement in vulnerable countries 
like Venezuela, Lebanon and Turkey.
It seems natural that some of this money will find it's way into existing crypto assets once the markey gains momentum again.

```{r,getdata,fig.height=8,fig.width=8}

mydate <- Sys.Date()-1

BTCURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(BTCURL,destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
btc_mc <- btcdat$data$quotes$quote$USD$market_cap

ETHURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(ETHURL,destfile="ethdat.txt")
ethdat <- fromJSON("ethdat.txt")
eth_mc <- ethdat$data$quotes$quote$USD$market_cap 

USDTURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=USDT&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(USDTURL,destfile="usdtdat.txt")
usdtdat <- fromJSON("usdtdat.txt")
usdt_mc <- usdtdat$data$quotes$quote$USD$market_cap

USDCURL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=USDC&convert=USD&interval=daily&time_start=2019-01-01&time_end=",mydate,sep="")
download.file(USDCURL,destfile="usdcdat.txt")
usdcdat <- fromJSON("usdcdat.txt")
usdc_mc <- usdcdat$data$quotes$quote$USD$market_cap

times_mc <- btcdat$data$quotes$quote$USD$timestamp
mc <- data.frame(times_mc,btc_mc,eth_mc,usdt_mc,usdc_mc)

mc$total <- mc$btc_mc + mc$eth_mc + mc$usdt_mc + mc$usdc_mc
mc$sc <- mc$usdt_mc + mc$usdc_mc
mc$scrat <- mc$sc / mc$total
mc$times_mc <- as.character(mc$times_mc)
mc$times_mc <- as.Date(sapply(strsplit(mc$times_mc,"T"),"[[",1))

plot(mc$times_mc,mc$btc_mc,type="l",col="orange",
  ylab="market cap (USD)",main="Major crypto and stable asset market cap")
lines(mc$times_mc,mc$eth_mc,col="blue")
lines(mc$times_mc,mc$usdt_mc,col="green")
lines(mc$times_mc,mc$usdc_mc,col="red")
legend("topleft", legend=c("BTC", "ETH", "USDT", "USDC"),
       col=c("orange", "blue", "green", "red"), lty=1:4)

plot(mc$times_mc,mc$usdt_mc,type="l",col="green",
  ylab="market cap (USD)",main="Major stable asset market cap")
lines(mc$times_mc,mc$usdc_mc,col="red")
legend("topleft", legend=c("USDT", "USDC"),
       col=c("green", "red"), lty=1:4)

```

## Now chart the stablecoin dominance

Next, we can see that the dominance is oscillating in an overall upward direciton.

Previously, we imagined that the stablecoin dominance would reach 3% for a second time, but it bottommed out at 6%.

The upward momentum is very strong, the recent local tops were 7.5% and 10.5%, so it seems logical it will reach 13.5% in the current downturn.
After the dominance starts to dip again, we should be looking at the 9% level for a potential bottom.

```{r,scdom,fig.height=8,fig.width=8}

p <- ggplot(mc, aes(x=times_mc, y=scrat)) +
  geom_line() + 
  xlab("") +
  geom_hline(yintercept=0.03, linetype="dashed", color = "red", size=2) +
  ylab("stablecoin dominance")
p


```

## Notes

Keep in mind that the tragectory of the stablecoin dominance changes can change dramatically, 
especially in a blow-off top when crypto valuations become rapidly overheated.
This indicator is not meant to be used in isolation in calling a top. 
Other supporting indicators could be NUPL, Pi Cycle Top Indicator, 200 Week Moving Average Heatmap and the power law chart. 
These are linked below:

* https://www.lookintobitcoin.com/charts/

* https://stats.buybitcoinworldwide.com/long-term-power-law/

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
