---
title: "Trading based on Directional Movement Index (DMI)"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

## Intro

Here, we're using the TSI as a buy/sell indicator for trading BTC.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")
library("quantmod")
library("TTR")
library("vioplot")
library("kableExtra")

```

## Get data

Obtaining BTC historical data (daily).

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

btcdat <- fromJSON("btcdat.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),4][2] 
mystart <- price[c(1,nrow(price)),4][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Basic look at the indicator

Default n = 14.

```{r,tsi1,fig.height=10,fig.width=8}

TSI <- function(x, n.first = 25, n.second = 13, n.signal = 7) {
  #True Strength Indicator
  #https://school.stockcharts.com/doku.php?id=technical_indicators:true_strength_index
  x <- try.xts(x, error = as.matrix)
  pc <- x - lag.xts(x, na.pad = T) #force lag.xts to get na padding
  dspc <- EMA(EMA(pc, n = n.first), n = n.second)
  dsapc <- EMA(EMA(abs(pc), n = n.first), n = n.second)
  
  tsi <- 100 * (dspc/dsapc)
  signal <- EMA(tsi, n = n.signal)
  r <- cbind(tsi, signal)
  r <- reclass(r, x)
  if (!is.null(dim(r))) colnames(r) <- c("tsi", "signal")
  return(r)
}

addTSI <- newTA(TSI, preFUN = Cl, col = c("blue", "red"))

par(mfrow=c(2,1))
price2 <- tail(price,730)

tsi <- TSI(price2$close , n.first = 25, n.second = 13, n.signal = 7 )

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tsi[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="TSI",main="TSI")
lines(as.Date(price2$date), tsi[,2]  , col="red"  )
grid()

price2 <- tail(price,150)
tsi <- TSI(price2$close, n.first = 25, n.second = 13, n.signal = 7 )
par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tsi[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="TSI",main="TSI")
lines(as.Date(price2$date), tsi[,2]  , col="red"  )
grid()

par(mfrow=c(1,1))


```

## TSI trading function

Assuming 0.1% trading fee, as per binance current rate.

```{r,tsifunc1}

mytsi <- function(price,n1,n2,ns,fee) {
  tsi <- TSI(price$close , n.first = n1, n.second = n2, n.signal = ns )
  price2 <- price
  price2 <- cbind(price2,tsi)
  price2$higher <- price2$tsi > price2$signal
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c(1,2)]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c(1,2)]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(c(prod(trades$diff),nrow(trades)))
}

```


## Backtesting

Set 1 higher time frames.

Best: n1=5, n2=5, nsig=50 --> 2692 returns per trade

```{r,backtest1}

fee=0.001

n1r <- seq(5,100,5)
n2r <- seq(5,100,5)
nsr <- seq(5,100,5)

params <- expand.grid(n1r,n2r,nsr)

colnames(params) <- c("n1","n2","nsignal")

res <- lapply(X=1:nrow(params) , FUN=function(i) {
  n1=as.numeric(params[i,1])
  n2=as.numeric(params[i,2])
  ns=as.numeric(params[i,3])
  mytsi(price,n1=n1,n2=n2,ns=ns,fee=fee)
} )

res <- as.data.frame(do.call(rbind, res))

colnames(res) <- c("returns","trades")

res2 <- cbind(params,res)

res2$returnspertrade <- res2$returns/res2$trades

head(res2[order(-res2$returns),],50) 
#%>% kbl() %>%  kable_paper("hover", full_width = F)

head(res2[order(-res2$returnspertrade),],50) 
#%>% kbl() %>%  kable_paper("hover", full_width = F)

res3 <- subset(res2,trades<200)

head(res3[order(-res3$returns),],50)
#%>% kbl() %>%  kable_paper("hover", full_width = F)

```

## Optimisation1

Long timeframe best

```{r,best1,fig.width=10,fig.height=8}

best_n1 = 5
best_n2 = 5
best_ns = 20

tsi <- TSI(price$close , n.first = best_n1, n.second = best_n2, n.signal = best_ns )
price2 <- price
price2 <- cbind(price2,tsi)
price2$higher <- price2$tsi > price2$signal
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

nrow(trades)
trades
prod(trades$diff)

trades1 <- trades

HEADER="TSI"
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tsi[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="TSI",main="TSI")
lines(as.Date(price$date), tsi[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Backtesting 2

Set 2 smaller time frames.

Best: n1=1, n2=2, nsig=67 --> 2692 returns per trade

```{r,backtest2}

n1r <- seq(1,10,1)
n2r <- seq(2,20,1)
nsr <- seq(2,100,1)

params <- expand.grid(n1r,n2r,nsr)

colnames(params) <- c("n1","n2","nsignal")

res <- lapply(X=1:nrow(params) , FUN=function(i) {
  n1=as.numeric(params[i,1])
  n2=as.numeric(params[i,2])
  ns=as.numeric(params[i,3])
  mytsi(price,n1=n1,n2=n2,ns=ns,fee=fee)
} )

res <- as.data.frame(do.call(rbind, res))

colnames(res) <- c("returns","trades")

res2 <- cbind(params,res)

res2$returnspertrade <- res2$returns/res2$trades

head(res2[order(-res2$returns),],50)
#%>% kbl() %>%  kable_paper("hover", full_width = F)

head(res2[order(-res2$returnspertrade),],50)
#%>% kbl() %>%  kable_paper("hover", full_width = F)

res3 <- subset(res2,trades<250)

head(res3[order(-res3$returns),],50)
#%>% kbl() %>%  kable_paper("hover", full_width = F)

```

## Optimisation 2

```{r,best2,fig.width=10,fig.height=8}

best_n1 = 2
best_n2 = 8
best_ns = 65

tsi <- TSI(price$close , n.first = best_n1, n.second = best_n2, n.signal = best_ns )
price2 <- price
price2 <- cbind(price2,tsi)
price2$higher <- price2$tsi > price2$signal
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

nrow(trades)
trades
prod(trades$diff)

trades2 <- trades

HEADER="TSI"
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tsi[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="TSI",main="TSI")
lines(as.Date(price$date), tsi[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Optimisation 3

```{r,best3,fig.width=10,fig.height=8}

best_n1 = 1
best_n2 = 2
best_ns = 2

tsi <- TSI(price$close , n.first = best_n1, n.second = best_n2, n.signal = best_ns )
price2 <- price
price2 <- cbind(price2,tsi)
price2$higher <- price2$tsi > price2$signal
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c(1,2)]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c(1,2)]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

nrow(trades)
trades
prod(trades$diff)

trades3 <- trades

HEADER="TSI"
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tsi[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="TSI",main="TSI")
lines(as.Date(price$date), tsi[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```


## Effect of fees

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

best_n1 = 5
best_n2 = 5
best_ns = 20

mytsi(price,n1=best_n1,n2=best_n2,ns=best_ns,fee=fee)

res_fees <- sapply(fee_levels,function(x) {
  mytsi(price,n1=best_n1,n2=best_n2,ns=best_ns,fee=x)
} )

plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext("higher time frame")
abline(h=hodl,col="red")
grid()


best_n1 = 2
best_n2 = 8
best_ns = 67

mytsi(price,n1=best_n1,n2=best_n2,ns=best_ns,fee=fee)

res_fees <- sapply(fee_levels,function(x) {
  mytsi(price,n1=best_n1,n2=best_n2,ns=best_ns,fee=x)
} )

plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext("lower time frame low trade number")
abline(h=hodl,col="red")
grid()


best_n1 = 1
best_n2 = 2
best_ns = 2

mytsi(price,n1=best_n1,n2=best_n2,ns=best_ns,fee=fee)

res_fees <- sapply(fee_levels,function(x) {
  mytsi(price,n1=best_n1,n2=best_n2,ns=best_ns,fee=x)
} )

plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext("lower time frame high trade number")
abline(h=hodl,col="red")
grid()



```

## Trades by year

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

trades1$year <- sapply(strsplit(trades1[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades1,year==y)[,5] } )
vioplot(yearly)
yr <- sapply(yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

trades2$year <- sapply(strsplit(trades2[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades2,year==y)[,5] } )
vioplot(yearly)
yr <- sapply(yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

trades3$year <- sapply(strsplit(trades3[,3],"-"),"[[",1)
yearly <- sapply(years,function(y) { x <- subset(trades3,year==y)[,5] } )
vioplot(yearly)
yr <- sapply(yearly,prod)
barplot(yr)
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```

## Conclusion

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
