---
title: "Ultimate Ocillator for BTC and ETH"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 8
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/ultimate_oscillator.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

From Investopedia:

>The Ultimate Oscillator is a range-bound indicator with a value that fluctuates between 0 and 100. Similar to the Relative Strength Index (RSI), levels below 30 are deemed to be oversold, and levels above 70 are deemed to be overbought. Trading signals are generated when the price moves in the opposite direction as the indicator, and are based on a three-step method.

>Larry Williams developed the Ultimate Oscillator in 1976 and published it in Stocks & Commodities Magazine in 1985. With many momentum oscillators correlating too heavily to near-term price movements, Williams developed the Ultimate Oscillator to incorporate multiple timeframes to smooth out the indicator's movements and provide a more reliable indicator of momentum, with fewer false divergences.

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
})

Sys.Date()

```

Reminder: this is not financial advice.

## Ultimate Oscillator for BTC

Obtaining BTC historical data (daily) from CoinMarketCap.com from June 2013 to present.

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat.txt")

dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, 
  price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

Here n = c(7, 14, 28) and wts = c(4, 2, 1).

Complete history.

```{r,uo1}

price2 <- price
uo <- ultimateOscillator(HLC(price2),n=c(7,14,28),wts=c(4,2,1))

par(mfrow=c(2,1))

plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="Price")
grid()

plot(uo ~ as.Date(price2$date),type="l",xlab="Date",main="Ultimate Oscillator")

par(mfrow=c(1,1))

```

Last 400 days.

```{r,uo2}

price2 <- tail(price, 400)
uo <- tail(uo, 400)

par(mfrow=c(2,1))

plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="Price")
grid()

plot(uo ~ as.Date(price2$date),type="l",xlab="Date",main="Ultimate Oscillator")

par(mfrow=c(1,1))

```

## UO Longer period

Here n = c(14, 28, 48) and wts = c(1, 1, 1).

Complete history.

```{r,uo3}

price2 <- price
uo <- ultimateOscillator(HLC(price2),n=c(14,28,48),wts=c(1,1,1))

par(mfrow=c(2,1))

plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="Price")
grid()

plot(uo ~ as.Date(price2$date),type="l",xlab="Date",main="Ultimate Oscillator")

par(mfrow=c(1,1))

```

With smoothing.

```{r,uo4}

price2 <- price
uo <- ultimateOscillator(log(HLC(price2)),n=c(28,48,96),wts=c(1,1,1))
sm1 <- SMA(ultimateOscillator(log(HLC(price2)),n=c(14,28,48),wts=c(1,1,1)),100)
sm2 <- SMA(ultimateOscillator(log(HLC(price2)),n=c(14,28,48),wts=c(1,1,1)),200)

par(mfrow=c(2,1))

plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="Price")
grid()

plot(uo ~ as.Date(price2$date),type="l",xlab="Date",main="Ultimate Oscillator")
lines( sm1 ~ as.Date(price2$date) , col="blue"   )
lines( sm2 ~ as.Date(price2$date) , col="red"   )
par(mfrow=c(1,1))

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
