---
title: "Does bitcoin volatility decrease over time?"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/dmi_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

As a market grows, it is thought that volatility reduces. Lets see if this is the case with bitcoin.

Historical data from https://www.investing.com/indices/investing.com-btc-usd-historical-data

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library(anytime)
})

```

Reminder: this is not financial advice.

## Get data

We will be simplifying things by working only with the market cap of the top two stablecoins Tether (USDT) and USD Coin (USDC) against cryptocurrencies Bitcoin (BTC) and Ethereum (ETH).

Data from https://www.investing.com/indices/investing.com-btc-usd-historical-data

The data is in CSV format and the dates need to be cleaned a bit.

```{r,getdata}

daily <- read.csv("BitcoinHistoricalData.csv")
daily <- daily[nrow(daily):1,]
daily <- apply(daily,2,function(x) { gsub(",","",x) } )

weekly <- read.csv("BitcoinHistoricalData_weekly.csv")
weekly <- weekly[nrow(weekly):1,]
weekly <- apply(weekly,2,function(x) { gsub(",","",x) } )

weekly <- as.data.frame(weekly)

weekly$Price <- as.numeric(weekly$Price)
weekly$Open <- as.numeric(weekly$Open)
weekly$High <- as.numeric(weekly$High)
weekly$Low <- as.numeric(weekly$Low)

weekly$Date <- anydate(weekly$Date)

```

## Volatility on the weekly

Plotting the weekly trading range as a volatility metric.

```{r,weekly1}

weekly$vix <- ( weekly$High - weekly$Low ) / weekly$Open * 100

plot(weekly$Date,weekly$vix)

mylm <- lm(weekly$vix ~ weekly$Date)

summary(mylm)

abline(mylm,col="red",lwd=2)

```

Now try log regression


```{r,logreg1}

model <- lm(weekly$vix ~ log(as.numeric(weekly$Date)))

summary(model)

plot(weekly$Date,weekly$vix,xlim=c(weekly$Date[1],weekly$Date[nrow(weekly)]+1000))

# generate points
x <- seq(from=as.numeric(weekly$Date)[1],
  to=as.numeric(weekly$Date)[nrow(weekly)]+1000,
  length.out=nrow(weekly) )

y <- predict(model,newdata=list(x=x), interval="confidence")

matlines(x,y, lwd=2)

```

With a trimmed y axis.

```{r,logreg2}

model <- lm(weekly$vix ~ log(as.numeric(weekly$Date)))

summary(model)

plot(weekly$Date,weekly$vix,
  xlim=c(weekly$Date[1],weekly$Date[nrow(weekly)]+1000),
  ylim=c(0,50)  )

# generate points
x <- seq(from=as.numeric(weekly$Date)[1],
  to=as.numeric(weekly$Date)[nrow(weekly)]+1000,
  length.out=nrow(weekly) )

y <- predict(model,newdata=list(x=x), interval="confidence")

matlines(x,y, lwd=2)

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
