---
title: "Does volatility decrease with increasing market cap?"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

## Intro

As a market grows, it is thought that volatility reduces. Lets see if this is the case with bitcoin.

```{r,lib}

library("jsonlite")
library("tidyverse")
library("runner")

```

## Get data

We will be simplifying things by working only with the market cap of the top two stablecoins Tether (USDT) and USD Coin (USDC) against cryptocurrencies Bitcoin (BTC) and Ethereum (ETH).

```{r,getdata}
download.file("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=2021-02-19",destfile="btcdat.txt")
btcdat <- fromJSON("btcdat.txt")
head(btcdat$data$quotes)

btc_mc <- btcdat$data$quotes$quote$USD$market_cap

times_mc <- btcdat$data$quotes$quote$USD$timestamp

percent_moves <- (btcdat$data$quotes$quote$USD$close - btcdat$data$quotes$quote$USD$open)/btcdat$data$quotes$quote$USD$open * 100

mc <- data.frame(times_mc, btc_mc, percent_moves, abs(percent_moves) )

mc$times_mc <- as.Date(sapply(strsplit(mc$times_mc,"T"),"[[",1))

```

## Calculate correlation

```{r,correl}

model <- lm(mc$abs.percent_moves. ~ times_mc, data = mc)
summary(model)

cor.test(as.numeric(mc$times_mc),mc$abs.percent_moves.)

plot(mc$times_mc,mc$abs.percent_moves.,pch=19,cex=0.5,ylab="absolute daily % change")
lines(as.numeric(mc$times_mc), predict(model), col = 'red',lwd=3)

plot(mc$times_mc,mc$abs.percent_moves.,pch=19,cex=0.5,ylab="absolute daily % change",ylim=c(2.4,3))
lines(as.numeric(mc$times_mc), predict(model), col = 'red',lwd=3)

```

This chart indicates that BTC volatility has reduced over time from daily swings of 2.9 percent per day to 2.4%.
This is still very high volatility as compared to stocks, which rarely have daily changes greater than 2-3%.
Statistical test suggests the association is 98% certain it is a real effect.
Bull/bear markets after halvings looks to cause a massive spike in volatility which tapers off after.

## Now plot by market cap

Put the market cap on the x axis.

```{r,vol2}

mc$log_btc_mc <- log(mc$btc_mc)

model <- lm(abs.percent_moves. ~ log_btc_mc, data = mc)
summary(model)

cor.test(mc$log_btc_mc,mc$abs.percent_moves.)

plot(mc$log_btc_mc ,mc$abs.percent_moves.,pch=19,cex=0.5,ylab="absolute daily % change",xlab="log10 mkt cap")
lines(mc$log_btc_mc, predict(model), col = 'red',lwd=3)

plot(mc$log_btc_mc ,mc$abs.percent_moves.,pch=19,cex=0.5,ylab="absolute daily % change",
  xlab="log10 mkt cap",ylim=c(2.3,3))

lines(mc$log_btc_mc, predict(model), col = 'red',lwd=3)

```

This chart shows that volatility actually increases with market cap, which might be due to BTC entering 
current bull market.

## Conclusion

This analysis shows there is a weak trend for reducing volatility over time.
Volatility increases with market cap but that may simply be due to entering recent all time highs.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
