---
title: "Does bitcoin volatility decrease over time?"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/volatility_size.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: bc1qgsecwc3e7r0g7zch8wum04a8hk74tfy488m2nv

XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

As a market grows, it is thought that volatility reduces. Lets see if this is the case with bitcoin.

Historical data from https://www.investing.com/indices/investing.com-btc-usd-historical-data

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library(anytime)
})

```

Reminder: this is not financial advice.

## Get data

Here I will be analysing the weekly trading range as a proxy for volatility.
Data are from https://www.investing.com/indices/investing.com-btc-usd-historical-data

The data is in CSV format and the dates need to be cleaned a bit.

```{r,historicaldata}

weekly <- read.csv("BitcoinHistoricalData_weekly.csv",stringsAsFactors=FALSE)
weekly <- weekly[nrow(weekly):1,]
weekly <- apply(weekly,2,function(x) { gsub(",","",x) } )

weekly <- as.data.frame(weekly,stringsAsFactors=FALSE)

weekly$Price <- as.numeric(weekly$Price)
weekly$Open <- as.numeric(weekly$Open)
weekly$High <- as.numeric(weekly$High)
weekly$Low <- as.numeric(weekly$Low)

weekly$Date <- anydate(weekly$Date)

weekly <- weekly[,c("Date","High","Low","Open")]
weekly <- weekly[grep("2021",weekly$Date,invert=TRUE),]

```

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=weekly&time_start=2021-01-01&time_end=",mydate,sep="")
download.file(URL,destfile="btcdat_weekly.txt")
btcdat <- fromJSON("btcdat_weekly.txt")
price <- btcdat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$open,stringsAsFactors=FALSE)
colnames(price) <- c("Date","High","Low","Open")
price$Date <- sapply(strsplit(as.character(price$Date),"T"),"[[",1)

price$Date <- anydate(price$Date)

weekly <- rbind(weekly,price)

```

## Volatility on the weekly

Plotting the weekly trading range as a volatility metric.

```{r,weekly1}

weekly$vix <- ( weekly$High - weekly$Low ) / weekly$Open * 100

plot(weekly$Date,weekly$vix,
 xlab="", ylab="VIX", main="weekly trading range / opening price" )

mylm <- lm(weekly$vix ~ weekly$Date)

abline(mylm,col="red",lwd=2)

```

Hmm, this looks okay but we expect volatility to taper off, not just linearly go to zero.

Now try log regression


```{r,logreg1}

model <- lm(weekly$vix ~ log(as.numeric(weekly$Date)))

plot(weekly$Date,weekly$vix,xlim=c(weekly$Date[1],weekly$Date[nrow(weekly)]+1000),
 xlab="", ylab="VIX", main="weekly trading range / opening price" )

# generate points
x <- seq(from=as.numeric(weekly$Date)[1],
  to=as.numeric(weekly$Date)[nrow(weekly)]+1000,
  length.out=nrow(weekly) )

y <- predict(model,newdata=list(x=x), interval="confidence")

matlines(x,y, lwd=2)

summary(model)

```

With a trimmed y axis.

```{r,logreg2}

plot(weekly$Date,weekly$vix,
  xlim=c(weekly$Date[1],weekly$Date[nrow(weekly)]+1000),
  ylim=c(0,50) ,
  xlab="", ylab="VIX", main="weekly trading range / opening price" )
grid()


# generate frecast points
x <- seq(from=as.numeric(weekly$Date)[1],
  to=as.numeric(weekly$Date)[nrow(weekly)]+1000,
  length.out=nrow(weekly) )

y <- predict(model,newdata=list(x=x), interval="confidence")

matlines(x,y, lwd=2)

```

It is quite interesting that the logarithmic regression yielded a more or less linear curve,
indicating that volatility is decreasing in a constant fashion.

Also interesting to see the increasing volatility at the the peak bubble years incl 2013, 2017, 2020, and lower volatility in 2016 and in 2021 
which could be possible described as reaccumulation phases with historically low volatility.

```{r,logreg3}

plot(weekly$Date,weekly$vix,
  xlim=c(weekly$Date[1],weekly$Date[nrow(weekly)]+1000),
  ylim=c(0,50) ,col="gray", pch=19,
  xlab="", ylab="VIX", main="weekly trading range / opening price" )
grid()
mtext("red line = 20 wk moving average")
lines(SMA(weekly$vix,n=20) ~ weekly$Date ,col="red",lwd=3)

```

The chart definitely shows that Q3/Q4 2021 have printed very low volatility, is this a precursor to a bull run, just like 2016?

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
