---
title: "Daily trade alerts for XRP"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/MDZanalytics/cryptoblog/blob/main/alerts_XRP.Rmd

## Intro

In the backtesting analysis, I have established five different profitable strategies for timing entry/exits.
Click on the hyperlinks below to see how these signals were identified.

1. Simple moving averages

2. Moving average crosses

3. Directional movement index (DMI)

4. True strength index (TSI) **BEST**

5. Stochastic oscillator (stoch)

6. Double RSI

This script will examine the price action and these trade signals just before the daily close.
If the signal is TRUE, it is bullish and if FALSE it is bearish.
When there is a switch from bearish to bullish it is a buy signal and when it switches from bullish to bearish it is a sell signal.
This HTML is updated just before the daily close each day, but push notifications for any trade signals are also sent by push bullet.

To receive these signals in real-time, subscribe to the following channels - it's free for a limited time!

https://www.pushbullet.com/channel?tag=xrp_signal

<a class="pushbullet-subscribe-widget" data-channel="xrp_signal" data-widget="button" data-size="small"></a>
<script type="text/javascript">(function(){var a=document.createElement('script');a.type='text/javascript';a.async=true;a.src='https://widget.pushbullet.com/embed.js';var b=document.getElementsByTagName('script')[0];b.parentNode.insertBefore(a,b);})();</script>

If you need some help to design your own trading signals/strategies, I am happy to help for a fee.
email me at mark.ziemann{αt}gmail.com for any enquiries/suggestions/feedback.

This report is distributed for FREE under the [MIT licence](https://github.com/MDZanalytics/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

Reminder: this analysis is not financial advice.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("RPushbullet")
  library("kableExtra")
})

TESTING = FALSE

MYNOTE = NULL

```

```{r,time}

#system("DATE=$(date -u) ; sed \"s/MYTIME/${DATE}/\" index0.html > index1.html")

mydate <- Sys.time()
attr(mydate, "tzone") <- "UTC"
mydate <- as.Date(mydate)

```

## Get XRP parameters

```{r,params}

params <- read.table("https://mdz-analytics.com/altcoins/XRP/XRP_dat.txt", header=TRUE)
params %>% kbl(caption="optimised backtested parameters") %>%  kable_paper("hover", full_width = F)

```

## Get XRP price data

Obtaining XRP price data (daily) for the last 300 days.

```{r,xrp_data1}

URL="https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=XRP&convert=USD&interval=daily&count=300"
download.file(URL,destfile="xrpdat.txt")
xrpdat <- fromJSON("xrpdat.txt")
price <- xrpdat$data$quotes
price <- data.frame(  as.Date(price$time_close) , price$quote$USD$high, price$quote$USD$low, price$quote$USD$close)
colnames(price) <- c("date","high","low","close")
#tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Now gather hourly price data for the last 23 hours.

```{r,xrp_data2}

URL="https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=XRP&convert=USD&interval=hourly&time_period=hourly&count=11"
download.file(URL,destfile="xrpdat.txt")
xrpdat <- fromJSON("xrpdat.txt")
price2 <- xrpdat$data$quotes
price2 <- data.frame( as.Date(price2$time_close) , price2$quote$USD$high, price2$quote$USD$low, price2$quote$USD$close,stringsAsFactors=FALSE)
colnames(price2) <- c("date","high","low","close")
#tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)

high <- max(price2[,2])
low <-  min(price2[,3])
close <- price2[nrow(price2),4]

df <- data.frame(date=mydate,high=high,low=low,close=close,stringsAsFactors=FALSE)
price <- rbind(price,df)
tail(price) %>% kbl() %>%  kable_paper("hover", full_width = F)

if ( TESTING == TRUE) {
  price[nrow(price),"Close"] <- price[nrow(price),"Close"] * 2
}

```

## XRP SMA indicator

Now to determine whether XRP has crossed the SMA line.

```{r, xrp_sma}

price2 <- price

n <- as.numeric(params[params$indicator == "SMA",2])

price2$ma <- SMA(Cl(price2),n=n)
price2$signal <- price2$close > price2$ma
tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)
today_signal <- price2[nrow(price2),"signal"]
previous_signal <- price2[(nrow(price2)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    xrp_signal="BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    xrp_signal="SELL"
  }
} else {
xrp_signal <- "NONE"
}

#if (today_signal==TRUE) {
# system("sed -i 's/XRP_SMA_DAILY/BULLISH/' index1.html")
# system("sed -i 's/XRP_SMA_COL/green/' index1.html")
#} else {
# system("sed -i 's/XRP_SMA_DAILY/BEARISH/' index1.html")
# system("sed -i 's/XRP_SMA_COL/red/' index1.html")
#}

message(paste("XRP SMA signal:",xrp_signal))

```

```{r, xrp_sma_plot,fig.height=8,fig.width=8}

plot(price2$close~as.Date(price2$date),type="l",
  xlab="Date",ylab="price (USD)",main="XRP w SMA")
grid()
lines(price2$ma~ as.Date(price2$date) ,col="red")

```

```{r,xrp_sma_push}

if ( xrp_signal != "NONE") {
  MYNOTE <- paste("The XRP SMA signal is", xrp_signal,".")
}

```

## XRP SMA cross indicator

The SMA cross is a reasonably good strategy.

```{r, xrp_cross}

price2 <- price

n <- as.numeric(unlist(strsplit(params[params$indicator == "SMAcross",2],",")))
n1 <- n[1]
n2 <- n[2]

price2$ma1 <- SMA(Cl(price2),n=n1)
price2$ma2 <- SMA(Cl(price2),n=n2)

price2$signal <- price2$ma2 > price2$ma1
tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)
today_signal <- price2[nrow(price2),"signal"]
previous_signal <- price2[(nrow(price2)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    xrp_signal="BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    xrp_signal="SELL"
  }
} else {
xrp_signal <- "NONE"
}

#if (today_signal==TRUE) {
# system("sed -i 's/XRP_SMA_CROSS_DAILY/BULLISH/' index1.html")
# system("sed -i 's/XRP_SMA_CROSS_COL/green/' index1.html")
#} else {
# system("sed -i 's/XRP_SMA_CROSS_DAILY/BEARISH/' index1.html")
# system("sed -i 's/XRP_SMA_CROSS_COL/red/' index1.html")
#}

message(paste("XRP SMA cross signal:",xrp_signal))

```


```{r, xrp_cross_plot,fig.height=8,fig.width=8}

plot(price2$close~as.Date(price2$date),type="l",
  xlab="Date",ylab="price (USD)",main="XRP w SMA cross")
grid()
lines(price2$ma1~ as.Date(price2$date) ,col="red")
lines(price2$ma2~ as.Date(price2$date) ,col="blue")

```

```{r,xrp_cross_push}

if ( xrp_signal != "NONE") {
  MYNOTE <- paste(MYNOTE, "The XRP SMA cross signal is", xrp_signal,".")
}

```

## XRP DMI indicator

Directional movement indicator is a good approach to identify trend changes.

```{r,xrp_dmi,fig.height=10,fig.width=8}

n <- as.numeric(params[params$indicator == "DMI",2])

dmi.adx <- ADX(price[,c("high","low","close")],n=n)
price2 <- cbind(price,dmi.adx)
price2$signal <- dmi.adx[,1] > dmi.adx[,2]
tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)

today_signal <- price2[nrow(price2),"signal"]
previous_signal <- price2[(nrow(price2)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    xrp_signal="DMI BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    xrp_signal="DMI SELL"
  }
} else {
  xrp_signal <- "NONE"
}

#if (today_signal==TRUE) {
# system("sed -i 's/XRP_DMI_DAILY/BULLISH/' index1.html")
# system("sed -i 's/XRP_DMI_COL/green/' index1.html")
#} else {
# system("sed -i 's/XRP_DMI_DAILY/BEARISH/' index1.html")
# system("sed -i 's/XRP_DMI_COL/red/' index1.html")
#}

message(paste("XRP DMI signal:",xrp_signal))

```

```{r,xrp_dmi_plot,fig.height=10,fig.width=8}

price2 <- as.data.frame(price2)
par(mfrow=c(2,1))
plot(price2$close~as.Date(price2$date),type="l",
  xlab="Date",ylab="price (USD)",main="XRP USD price")
grid()
plot(price2$DIp~as.Date(price2$date),type="l", col="blue",
  xlab="Date",ylab="price (USD)",main="XRP DMI")
grid()
lines(price2$DIn ~ as.Date(price2$date) ,col="red")

```

```{r,xrp_dmi_push}

if ( xrp_signal != "NONE") {
  MYNOTE <- paste(MYNOTE,"XRP is flashing a ", xrp_signal,".")
}

```

## XRP TSI indicator

True Strength Indicator is one of the more profitable approaches.
Its success lies in how it changes rapidly when momentum shifts.
For most coins, this indicator performs "best" in backtesting analysis, which means it
should carry relatively more weight than other indicators.

```{r,xrp_tsi,fig.height=10,fig.width=8}

n <- as.numeric(unlist(strsplit(params[params$indicator == "TSI",2],",")))
n1 <- n[1]
n2 <- n[2]
ns <- n[3]

TSI <- function(x, n.first = 25, n.second = 13, n.signal = 7) {
  #True Strength Indicator
  #https://school.stockcharts.com/doku.php?id=technical_indicators:true_strength_index
  x <- try.xts(x, error = as.matrix)
  pc <- x - lag.xts(x, na.pad = T) #force lag.xts to get na padding
  dspc <- EMA(EMA(pc, n = n.first), n = n.second)
  dsapc <- EMA(EMA(abs(pc), n = n.first), n = n.second)

  tsi <- 100 * (dspc/dsapc)
  signal <- EMA(tsi, n = n.signal)
  r <- cbind(tsi, signal)
  r <- reclass(r, x)
  if (!is.null(dim(r))) colnames(r) <- c("tsi", "signal")
  return(r)
}

tsi <- TSI(price$close , n.first = n1, n.second = n2, n.signal = ns )
colnames(tsi) <- c("tsi","sig")

price2 <- cbind(price,tsi)
price2$signal <- tsi[,1] > tsi[,2]
tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)

today_signal <- price2[nrow(price2),"signal"]
previous_signal <- price2[(nrow(price2)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    xrp_signal="TSI BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    xrp_signal="TSI SELL"
  }
} else {
  xrp_signal <- "NONE"
}

#if (today_signal==TRUE) {
# system("sed -i 's/XRP_TSI_DAILY/BULLISH/' index1.html")
# system("sed -i 's/XRP_TSI_COL/green/' index1.html")
#} else {
# system("sed -i 's/XRP_TSI_DAILY/BEARISH/' index1.html")
# system("sed -i 's/XRP_TSI_COL/red/' index1.html")
#}

message(paste("XRP TSI signal:",xrp_signal))

```

```{r,xrp_tsi_plot,fig.height=10,fig.width=8}

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tsi[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="TSI",main="TSI")
lines(as.Date(price2$date), tsi[,2]  , col="red"  )
grid()

par(mfrow=c(1,1))

```

```{r,xrp_tsi_push}

if ( xrp_signal != "NONE") {
  MYNOTE <- paste(MYNOTE,"XRP is flashing a ", xrp_signal,".")
}

```

## XRP stochastic oscillator indicator

Similar to the TSI, the stoch can pinpoint early changes in momentum, however it may give many
false positives.

```{r,xrp_stoch,fig.height=10,fig.width=8}

n <- as.numeric(unlist(strsplit(params[params$indicator == "stoch",2],",")))
n1 <- n[1]
n2 <- n[2]
n3 <- n[3]

sto <- stoch(HLC(price), nFastK=n1 , nFastD=n2 , nSlowD=n2 , bounded = TRUE, smooth=n3)

price2 <- cbind(price,sto)
price2$signal <- price2$fastK > price2$fastD

tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)

today_signal <- price2[nrow(price2),"signal"]
previous_signal <- price2[(nrow(price2)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    xrp_signal="stoch BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    xrp_signal="stoch SELL"
  }
} else {
  xrp_signal <- "NONE"
}

#if (today_signal==TRUE) {
# system("sed -i 's/XRP_STOCH_DAILY/BULLISH/' index1.html")
# system("sed -i 's/XRP_STOCH_COL/green/' index1.html")
#} else {
# system("sed -i 's/XRP_STOCH_DAILY/BEARISH/' index1.html")
# system("sed -i 's/XRP_STOCH_COL/red/' index1.html")
#}

message(paste("XRP stoch signal:",xrp_signal))

```

```{r,xrp_stoch_plot,fig.height=10,fig.width=8}

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$fastK ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="index",main="stochastic oscillator")
lines(as.Date(price2$date), price2$fastD  , col="red"  )
grid()

par(mfrow=c(1,1))

```

```{r,xrp_stoch_push}

if ( xrp_signal != "NONE") {
  MYNOTE <- paste(MYNOTE,"XRP is flashing a ", xrp_signal,".")
}

```

## XRP double RSI indicator

Double RSI is simply two RSI lines.

```{r,xrp_rsi1,fig.height=10,fig.width=8}

n <- as.numeric(unlist(strsplit(params[params$indicator == "RSI2",2],",")))
n1 <- n[1]
n2 <- n[2]

rsi1 <- RSI(price$close,n=n1,maType=EMA)
rsi2 <- RSI(price$close,n=n2,maType=EMA)

price2 <- cbind(price,rsi1,rsi2)
price2$signal <- price2$rsi1 > price2$rsi2

tail(price2) %>% kbl() %>%  kable_paper("hover", full_width = F)

today_signal <- price2[nrow(price2),"signal"]
previous_signal <- price2[(nrow(price2)-1),"signal"]
if (today_signal != previous_signal ) {
  if ( (today_signal == TRUE ) & (previous_signal == FALSE ) ) {
    xrp_signal="RSI2 BUY"
  }
  if ( (today_signal == FALSE ) & (previous_signal == TRUE ) ) {
    xrp_signal="RSI2 SELL"
  }
} else {
  xrp_signal <- "NONE"
}

#if (today_signal==TRUE) {
# system("sed -i 's/XRP_RSI_DAILY/BULLISH/' index1.html")
# system("sed -i 's/XRP_RSI_COL/green/' index1.html")
#} else {
# system("sed -i 's/XRP_RSI_DAILY/BEARISH/' index1.html")
# system("sed -i 's/XRP_RSI_COL/red/' index1.html")
#}

message(paste("XRP RSI signal:",xrp_signal))

```

```{r,xrp_rsi_plot,fig.height=10,fig.width=8}

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$rsi1 ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="index",main="double RSI")
lines(as.Date(price2$date), price2$rsi2  , col="red"  )
grid()

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$rsi1/price2$rsi2 ~ as.Date(price2$date),type="l",
  xlab="Date",ylab="RSI ratio",main="double RSI")
grid()
abline(h=1,lty=2,lwd=2)
par(mfrow=c(1,1))

```

```{r,xrp_rsi_push}

if ( xrp_signal != "NONE") {
  MYNOTE <- paste(MYNOTE,"XRP is flashing a ", xrp_signal,".")
}

```

## Send a push notification.

```{r,push}

MYNOTE

if ( !is.null(MYNOTE) ) {
  MYNOTE <- paste(MYNOTE, ". Visit https://mdz-analytics.com/altcoins/XRP/alerts_XRP.html for the details")
  pbPost("note", "Crypto Alert", MYNOTE ,channel="xrp_signal")
}

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
