---
title: "Systematic analysis coin market indicator performance"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/coin_script_indicators.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

Previously I performed backtesting for major altcoins.
Now it is time to look at which indicator performed best across all coins.

```{r,lib0}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
})

```

Reminder: this is not financial advice.

## Get data

Here I'm looking at the backtest results for the coins examined.
I want to know which indicator gives the best,most reliable overall performance.

```{r,data1}

files <- list.files(".",pattern="_dat.txt")

l <- lapply(files,function(f) {
  df <- read.table(f,header=TRUE)
  name <- sapply(strsplit(f,"_"),"[[",1)
  df$coin <- name
  return(df)
})

df <- do.call(rbind,l)


sma <- subset(df,indicator=="SMA")[,"xhodl"]
ema <- subset(df,indicator=="EMA")[,"xhodl"]
smacross <- subset(df,indicator=="SMAcross")[,"xhodl"]
emacross <- subset(df,indicator=="EMAcross")[,"xhodl"]
dmi <- subset(df,indicator=="DMI")[,"xhodl"]
tsi <- subset(df,indicator=="TSI")[,"xhodl"]
stoch <- subset(df,indicator=="stoch")[,"xhodl"]
rsi2 <- subset(df,indicator=="RSI2")[,"xhodl"]

l2 <- list("SMA"=sma,"EMA"=ema,"SMAcross"=smacross,"EMAcross"=emacross,"DMI"=dmi,
  "TSI"=tsi,"stoch"=stoch,"RSI2"=rsi2)


l2 <- l2[order(unlist(lapply(l2,median)))]

par(mar=c(5,7,5,1))

vioplot(l2,main="total returns for backtested indicators for all coins",horizontal=TRUE,las=1) ; grid()

vioplot(l2,ylim=c(-10,40), horizontal=TRUE,las=1,
  main="total returns for backtested indicators for all coins")
grid()

boxplot(l2, ylim=c(-10,40),horizontal=TRUE,las=1,main="total ROI by indicator") ; grid()

barplot(unlist(lapply(l2,median)),horiz=TRUE,las=1,main="median total ROI") ; grid()

```

At writing, the stoch performed best, followed by SMAcross

## Conclusion

Interestingly the stoch performed best in this analysis (2022/07/20), followed by SMAcross and EMAcross, DMI and TSI.

## Session information

```{r,sessionInfo0}

sessionInfo()

```

