---
title: "Trading based on MA and EMAs"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/ma_trading_xx.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

Trading off simple moving averages (SMA), exponenial moving averages (EMA) and crosses between these is 
a popular method for timing entry and exit of positions.
In this blog post I will be exploring MA strategies on the daily timeframe to find the ones which 
perform best.

This page is designed to be updated daily - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
})

Sys.Date()

```

Reminder: this is not financial advice.

## Get data

Obtaining historical data (daily) from CoinMarketCap.com

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=TICKER&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
#download.file(URL,destfile="dat.txt")
dat <- fromJSON("dat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close,price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

It's important to calculate the returns for simply holding the asset rather than actively trading.

The number shown here is the fold return on investment since listing on coinmarketcap.com

```{r,hodl}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Moving average function

Here are the functions for trading based on SMA and EMAs, buying when the price breaks through the
MA line and sell when it falls below the MA line. All trades are executed at the end of the daily session (UTC midnight).

```{r,functions}

sma <- function(price,n,fee) {
ma <- SMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if (nrow(price2)>0) {
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(c(prod(trades$diff),nrow(trades)))
} else {
  return(c(0,0))
}
}

ema <- function(price,n,fee) {
ma <- EMA(Cl(price),n=n)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if (nrow(price2)>0) {
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(c(prod(trades$diff),nrow(trades)))
} else {
  return(c(0,0))
}
}

```

## Backtesting

Let's look back in time and trial SMA and EMA strategies with an interval from 3 to 200 days.

Let's assume a 0.1% trading fee, as per binance current rate.


```{r,backtest}

fee=0.001

myrange <- 3:200

res <- sapply(myrange,function(n) { sma(price,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_sma <- res

res <- sapply(myrange,function(n) { ema(price,n,fee) } )
colnames(res) <- myrange
res <- t(res)
colnames(res) <- c("returns","trades")
res <- as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res_ema <- res

```

## Plots

Now let's look at which of these approaches works best.

```{r,plots1,fig.height=8,fig.width=8}

plot(x=myrange, y=res_sma$returns, pch=19,
  xlab="moving average interval (days)",ylab="fold return",main="daily SMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_sma$returns, pch=19,
  xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="daily SMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_ema$returns, pch=19,
  xlab="moving average interval (days)",ylab="fold return",main="daily EMA")
abline(h=hodl,col="red")
grid()

plot(x=myrange,y=res_ema$returns, pch=19,
  xlim=c(1,50),xlab="moving average interval (days)",ylab="fold return",main="daily EMA")
abline(h=hodl,col="red")
grid()

```

Some of those gains look very high, but the time frames are quite short.
Could it mean that there are a lot of trades each with very small gains?

```{r,plots2,fig.height=8,fig.width=8}

plot(res_sma$mean,res_sma$returns, main="daily SMA",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$mean,res_sma$returns , cex=1, labels=rownames(res_sma))
abline(h=hodl,col="red")

head(res_sma[order(-res_sma$mean),],20) %>% kbl(caption="Highest ROI SMA strategy") %>%  kable_styling()

plot(res_ema$mean,res_ema$returns, main="daily EMA",
  xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$mean,res_ema$returns , cex=1, labels=rownames(res_ema))
abline(h=hodl,col="red")

head(res_ema[order(-res_ema$mean),],20) %>% kbl(caption="Highest ROI EMA strategy") %>%  kable_styling()

```

Now we'll have a look at how many trades are required for these approaches.
Ideally we don't want to be making trades every few days.

```{r,plots3,fig.height=8,fig.width=8}

plot(res_sma$trades,res_sma$returns, main="daily SMA",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_sma$trades,res_sma$returns , cex=1, labels=rownames(res_sma))
abline(h=hodl,col="red")

plot(res_ema$trades,res_ema$returns, main="daily EMA",
  xlab="no. trades",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res_ema$trades,res_ema$returns , cex=1, labels=rownames(res_ema))
abline(h=hodl,col="red")

```

One good approach is to set a minimum mean ROI for a strategy.
I suggest 1.08 which means on average each trade should give 8% upside.
We then select the strategy with mean ROI>1.08 and the overall highest ROI.

```{r,top}

res_sma <- subset(res_sma,mean>1.08)
res_sma <- res_sma[order(-res_sma$returns),]
res_sma$MA <- rownames(res_sma)
head(res_sma,10) %>% kbl(caption="Best performing SMA strategies") %>%  kable_styling()

res_ema <- subset(res_ema,mean>1.08)
res_ema <- res_ema[order(-res_ema$returns),]
res_ema$MA <- rownames(res_ema)
head(res_ema,10) %>% kbl(caption="Best performing EMA strategies") %>%  kable_styling()

```

## Optimisation

Now let's select the two best approaches and show the history of trades.

### Best SMA

```{r,best_sma,fig.width=10,fig.height=8}

best_sma = as.numeric(rownames(head(res_sma[order(-res_sma$returns),],1)))
if( length(best_sma) == 0 ) { best_sma=2 }
message(paste("Best SMA period:",best_sma))
ma <- SMA(Cl(price),n=best_sma)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]  
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
sma_trades <- cbind(buy,sell)
sma_trades$diff <- ( ( sma_trades$sell_price - sma_trades$buy_price ) / sma_trades$buy_price ) +1 - (2 * fee )

message(paste("Number of trades:",nrow(sma_trades)))
message(paste("Total ROI:",prod(sma_trades$diff)))
message(paste("Returns over hodl:",prod(sma_trades$diff)/hodl))

sma_dat <- c("indicator"="SMA",
  "parameter"=best_sma,
  "meanROI"= mean(sma_trades$diff) ,
  "totalROI"= prod(sma_trades$diff) ,
  "ntrades"= nrow(sma_trades) ,
  "ndays"= nrow(price) ,
  "xhodl"= prod(sma_trades$diff)/hodl )

sma_trades  %>% kbl() %>%  kable_styling()
SMA_HEADER=paste(best_sma,"day moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```

### SMA recent action

```{r,sma_recent}

plot(tail(price$close,200)~tail(as.Date(price$date),200),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=SMA_HEADER)
grid()
lines(tail(ma,200) ~ tail(as.Date(price$date),200) ,col="red")

```

### Best EMA

Now have a look at the best EMA option.

```{r,best_ema,fig.width=10,fig.height=8}

best_ema = as.numeric(rownames(head(res_ema[order(-res_ema$returns),],1)))
if( length(best_ema) == 0 ) { best_ema=2 }
message(paste("Best EMA period:",best_ema))
ma <- EMA(Cl(price),n=best_ema)
price2 <- price
price2$ma <- ma
price2$higher <- price2$close > price2$ma
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
ema_trades <- cbind(buy,sell)
ema_trades$diff <- ( ( ema_trades$sell_price - ema_trades$buy_price ) / ema_trades$buy_price ) +1 - ( 2 * fee )

message(paste("Number of trades:",nrow(ema_trades)))
message(paste("Total ROI:",prod(ema_trades$diff)))
message(paste("Returns over hodl:",prod(ema_trades$diff)/hodl))

ema_dat <- c("indicator"="EMA",
  "parameter"=best_ema ,
  "meanROI"= mean(ema_trades$diff) ,
  "totalROI"= prod(ema_trades$diff) ,
  "ntrades"= nrow(ema_trades) ,
  "ndays"= nrow(price) ,
  "xhodl"= prod(ema_trades$diff)/hodl )

ema_trades  %>% kbl() %>%  kable_styling()
EMA_HEADER=paste(best_ema,"day exponential moving average")
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(ma~ as.Date(price$date) ,col="red")

```

### EMA recent action

```{r,ema_recent}

plot(tail(price$close,200)~tail(as.Date(price$date),200),type="l",log="y",
  xlab="Date",ylab="price (USD)",main=EMA_HEADER)
grid()
lines(tail(ma,200) ~ tail(as.Date(price$date),200) ,col="red")

```

## Effect of fees

The fee levels are a serious consideration because they will erode the profits if fees are too high.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_sma_fees <- sapply(fee_levels,function(x) { sma(price,best_sma,fee=x) } )[1,]
plot(fee_levels,res_sma_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(SMA_HEADER)
abline(h=hodl,col="red")
grid()

res_ema_fees <- sapply(fee_levels,function(x) { ema(price,best_ema,fee=x) } )[1,]
plot(fee_levels,res_ema_fees ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(EMA_HEADER)
abline(h=hodl,col="red")
grid()

```

## Trades by year

We know that bitcoin moves in four year cycles. Lets take a look at profits over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

sma_trades$year <- sapply(strsplit(sma_trades[,3],"-"),"[[",1)
sma_yearly <- sapply(years,function(y) { x <- subset(sma_trades,year==y)[,5] } )
sma_yearly <- sma_yearly[which(lapply(sma_yearly,length)>0)]
vioplot(sma_yearly,ylab="trade ROI distribution",main="daily SMA strategy")

yr <- sapply(sma_yearly,prod)
barplot(yr,ylab="yearly fold returns",main="daily SMA strategy")
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

ema_trades$year <- sapply(strsplit(ema_trades[,3],"-"),"[[",1)
ema_yearly <- sapply(years,function(y) { x <- subset(ema_trades,year==y)[,5] } )
ema_yearly <- ema_yearly[which(lapply(ema_yearly,length)>0)]
vioplot(ema_yearly,ylab="trade ROI distribution",main="daily EMA strategy")

yr <- sapply(ema_yearly,prod)
barplot(yr,ylab="yearly fold returns",main="daily EMA strategy")
text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
abline(h=1,col="red",lty=2)

```

## Data export

Saving data: indicator, parameter, meanROI, returns, ntrades, ndays, xhodl

```{r,export}

filename <- paste(TICKER,"_dat.txt",sep="")
write.table(t(sma_dat),file=filename,append=TRUE,row.names=FALSE)
write.table(t(ema_dat),file=filename,append=TRUE,col.names=FALSE,row.names=FALSE)

```
## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
