---
title: "Trading TICKER based on double RSI"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/rsi2_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.

BTC: 3PCBnq7TH37t6uQmVePjpHaHvzFFh72ZbM

LTC: MLzwpLhHE81xB459H15sDN7kBW16kLgHCF

## Intro

There are many indicators of market momentum, one of them is the relative strength index (RSI).
It is often used by traders as an entry or exit signal on its own, but here we are using two 
RSI lines, a shorter period and a longer one.
When the short period line crosses above the longer period one it is a buy signal and when the
short interval line passes under the longer interval line it is a sell signal.
I will be testing different intervals for the short and long RSI lines to see which gives the best results.

This report is time sensitive - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library("parallel")
})

Sys.Date()

CORES = detectCores()

```

Reminder: this is not financial advice.

## Get data

Obtaining historical data (daily) from CoinMarketCap.com.

```{r,getdata}

mydate <- Sys.Date()-1
URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=TICKER&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
#download.file(URL,destfile="dat.txt")
dat <- fromJSON("dat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```

## Calculate hodl returns

It's important to calculate the returns for simply holding the asset rather than actively trading.

The number shown here is the fold return on investment.

```{r,hodl}

myend <- price[c(1,nrow(price)),4][2] 
mystart <- price[c(1,nrow(price)),4][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Basic look at the double RSI indicator

Here we can see the RSI over the past 730 days, using the intervals of 14d and 28d.
The default moving average algorithm is the Welles Wilder EMA.

```{r,rsi2,fig.height=10,fig.width=8}

par(mfrow=c(2,1))
price2 <- tail(price,730)
rsi1 <- RSI(price2$close,n=14,maType=EMA)
rsi2 <- RSI(price2$close,n=28,maType=EMA)

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(rsi1 ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="RSI",main="RSI 14d/28d")
lines(as.Date(price2$date), rsi2  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Double RSI trading function

Here is the function for trading based on RSI crosses, buying when the short term indicator 
goes above the longer term line, and sell when it falls below.

All trades are executed at the end of the daily session (UTC midnight).

```{r,rsi2_func}

rsi <- function(price,n1,n2,maType=EMA,fee=0.001) {
  rsi1 <- RSI(price$close,n=n1,maType=maType)
  rsi2 <- RSI(price$close,n=n2,maType=maType)
  price2 <- price
  price2 <- cbind(price2,rsi1,rsi2)
  price2$higher <- price2$rsi1 > price2$rsi2
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if (nrow(price2)>0) { 
    if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
    if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
    buy <- subset(price2,trade==1)[,c("date","close")]
    colnames(buy) <- c("buy_date","buy_price")
    sell <- subset(price2,trade==-1)[,c("date","close")]
    colnames(sell) <- c("sell_date","sell_price")
    if( nrow(buy) != nrow(sell) ) {
      MIN=min(nrow(buy),nrow(sell))
      buy <- buy[1:MIN,]
      sell <- sell[1:MIN,]
    }
    trades <- cbind(buy,sell)
    trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
    return(c(prod(trades$diff),nrow(trades)))
  } else {
    return(c(0,0))
  }
}

```

## Backtesting

Now let's go back in time to test the performance of the double RSI indicator using different intervals.

Let's assume a 0.1% trading fee, as per binance current rate.

```{r,backtestema}

fee=0.001

params <- expand.grid(4:200,10:200)
colnames(params) <- c("n1","n2")
params <- subset(params,n2>n1)
dim(params)

res <- mclapply(1:nrow(params), function(x) {
  rsi(price,n1=params[x,1],n2=params[x,2],maType=EMA,fee=0.001)
} , mc.cores= CORES)
res <- do.call(rbind,res)
colnames(res) <- c("returns","trades")

res <-  as.data.frame(res)
res$mean <- res$returns^(1/res$trades)
res <- cbind(params,res,stringsAsFactors=FALSE)
#res %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Here is a table of results.

```{r,filt1}

res2 <- res[order(-res$returns),]

head(res2,20) %>% 
  kbl(caption="highest returning strategies") %>%
  kable_paper("hover", full_width = F)

res2 <- res[order(-res$mean),]

head(res2,20) %>%
  kbl(caption="highest mean return strategies") %>%
  kable_paper("hover", full_width = F)

res2 <- res[order(-res$returns),]

res2 <- subset(res2,trades<200)

head(res2,20) %>%
  kbl(caption="highest returning strategies with fewer than 200 trades") %>%
  kable_paper("hover", full_width = F)

res2 <- subset(res2,mean>1.08)

head(res2,20) %>%
  kbl(caption="highest returning strategies with fewer than 200 trades and mean ROI > 1.1") %>%
  kable_paper("hover", full_width = F)

```

I also tried SMA but the results were disappointing.

## Plots

Let's examine some of the trends in the data with some charts.

```{r,plots,fig.height=8,fig.width=8}

plot(res$trades,res$returns,log="y", cex=2, pch=19,col="gray",
  xlab="num trades",ylab="fold returns")
text(res$trades,res$returns,labels=res$period,cex=1.5)
abline(h=hodl,col="red")
grid()

plot(res$mean,res$returns,xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res$mean,res$returns , cex=1, labels=res$period)
abline(h=hodl,col="red")

plot(res$mean,res$returns,log="y",xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res$mean,res$returns , cex=1, labels=res$period)
abline(h=hodl,col="red")

if (length(res2$trades)>0 ) {

plot(res2$trades,res2$returns,log="y", cex=2, pch=19,col="gray",
  ,xlab="num trades",ylab="fold returns")
text(res$trades,res$returns,labels=res$period,cex=1.5)
abline(h=hodl,col="red")
grid()

plot(res2$mean,res2$returns,log="y",xlab="mean ROI",ylab="total ROI",cex=2,pch=19,col="gray")
grid()
text(res$mean,res$returns , cex=1, labels=res$period)
abline(h=hodl,col="red")

}

```

Based on this we can do some filtering for mean>1.08 and returns>2000.


```{r,filt2}

top <- subset(res,mean>1.08 )
top <- head(top,100)
top %>%
  kbl(caption="optimal strategies with high mean ROI and high returns") %>%
  kable_paper("hover", full_width = F)


```

## Best RSI

This is good indicator because mean gains are >1.08 and number of trades is less than 200.

```{r,best,fig.width=10,fig.height=8}

best_n1 = top[1,1]
best_n2 = top[1,2]
if(is.na(best_n1)) { best_n1 = 3 }
if(is.na(best_n2)) { best_n2 = 3 }
 
message(paste("Selected RSI periods:",best_n1,best_n2))

rsi1 <- RSI(price$close,n=best_n1,maType=EMA)
rsi2 <- RSI(price$close,n=best_n2,maType=EMA)
price2 <- price
price2 <- cbind(price2,rsi1,rsi2)
price2$higher <- price2$rsi1 > price2$rsi2
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( nrow(price2)>0 ) {
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))
}

rsi2_dat <- c("indicator"="RSI2",
  "parameter"= paste(best_n1,best_n2,sep=",") ,
  "meanROI"= mean(trades$diff) ,
  "totalROI"= prod(trades$diff) ,
  "ntrades"= nrow(trades) ,
  "ndays"= nrow(price) ,
  "xhodl"= prod(trades$diff)/hodl )

trades %>% kbl() %>%  kable_styling()

HEADER=paste("n1:",best_n1,"n2:",best_n2)
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(rsi1 ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="RSI",main=HEADER)
lines(as.Date(price$date), rsi2  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Effect of fees

It is important to factor in the effect of fees because higher exchange fees will erode profits.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_fees <- sapply(fee_levels,function(x) { rsi(price,best_n1,best_n2,fee=x) } )
plot(fee_levels,res_fees[1,] ,pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext(HEADER)
abline(h=hodl,col="red")
grid()


```

## Trades by year

It is good to see whether the strategy gives consistent returns over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

if (nrow(trades)>0) {
  trades$year <- sapply(strsplit(trades[,3],"-"),"[[",1)
  yearly <- sapply(years,function(y) { x <- subset(trades,year==y)[,5] } )
  yearly <- yearly[which(lapply(yearly,length)>0)]
  vioplot(yearly)
  yr <- sapply(yearly,prod)
  barplot(yr)
  text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
  abline(h=1,col="red",lty=2)
}

```

## Recent action

Recent action of the indicator.

```{r,update,fig.height=8,fig.width=8}

n=100
price <- cbind(price,rsi1,rsi2)
price <- tail(price,n)

HEADER=paste("n1:",best_n1,"n2:",best_n2)
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price$rsi1 ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="RSI",main=HEADER)
lines(as.Date(price$date), price$rsi2  , col="red"  )
grid()

tail(price ,20) %>%
  kbl(caption="Last 20 days") %>%
  kable_paper("hover", full_width = F)


```

## Data export

Saving data: indicator, parameter, meanROI, returns, ntrades, ndays, xhodl

```{r,export}

filename <- paste("TICKER","_dat.txt",sep="")
write.table(t(rsi2_dat),file=filename,append=TRUE,row.names=FALSE,col.names=FALSE)

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
