---
title: "Trading TICKER based on the Stochastic indicator"
author: "Mark Ziemann https://mdz-analytics.com"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
theme: cosmo
---

Source: https://github.com/markziemann/cryptoblog/blob/main/stoch_trading.Rmd

This report is distributed for FREE under the [MIT licence](https://github.com/markziemann/cryptoblog/blob/main/LICENSE),
but if you find it useful, consider a small tip.


XMR:4BGrdeAF5qyJQXjzWF4W5uCZF7WuwJU16BfPtgg1WJMnf33jZMtLvoF1jRtZBGpLtz5BQZaLYiBFJJC488anty64FB7SASD

## Intro

Here, we're using the stochastic as a buy/sell indicator for trading.

This report is time sensitive - today's date is shown below.

```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("kableExtra")
  library("parallel")
})

Sys.Date()

CORES=detectCores()

```

Reminder: this is not financial advice.

## Get data

Obtaining historical data (daily) from CoinMarketCap.com.

```{r,getdata}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=TICKER&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")
#download.file(URL,destfile="dat.txt")
dat <- fromJSON("dat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close")
price$date <- sapply(strsplit(as.character(price$date),"T"),"[[",1)

```
## Calculate hodl returns

```{r,hodl}

myend <- price[c(1,nrow(price)),4][2] 
mystart <- price[c(1,nrow(price)),4][1] 
hodl <- (myend - mystart) / mystart
hodl

```

## Basic look at the indicator

Default indicator.
Slightly too fast.

```{r,sto1,fig.height=10,fig.width=8}

price2 <- tail(price,200)
sto <- stoch(HLC(price2), nFastK = 14, nFastD = 3, nSlowD = 3, bounded = TRUE, smooth=1)

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(sto[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="stoch",main="stoch fast")
lines(as.Date(price2$date), sto[,2]  , col="red"  )
grid()

fee = 0.001

sto <- stoch(HLC(price), nFastK=14 , nFastD=3 , nSlowD=3 , bounded = TRUE, smooth=1)
price2 <- price
price2 <- cbind(price2,sto)
price2$higher <- price2$fastK > price2$fastD
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

basic_mean <- mean(trades$diff)
basic_returns <- prod(trades$diff)

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

```

Now with the slower settings.

```{r,sto2,fig.height=10,fig.width=8}

price2 <- tail(price,200)
sto <- stoch(HLC(price2), nFastK = 21, nFastD = 14, nSlowD = 21, bounded = TRUE, smooth=7)

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(sto[,1] ~ as.Date(price2$date),type="l",col="blue",
  xlab="Date",ylab="stoch",main="stoch slow")
lines(as.Date(price2$date), sto[,2]  , col="red"  )
grid()

par(mfrow=c(1,1))

sto <- stoch(HLC(price), nFastK=21 , nFastD=14 , nSlowD=21 , bounded = TRUE, smooth=7)
price2 <- price
price2 <- cbind(price2,sto)
price2$higher <- price2$fastK > price2$fastD
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

```

## Stoch trading function

Assuming 0.1% trading fee, as per binance current rate.


```{r,stofunc1}

mystoch <- function(price,nFastK=14,nFastD=3,nSlowD=3,smooth=1,fee) {
mytry <- try (  sto <- stoch(HLC(price), nFastK , nFastD , nSlowD , bounded = TRUE, smooth=smooth)  ,  silent = TRUE)
if ( typeof(mytry) == "double" ) {
  sto <- stoch(HLC(price), nFastK , nFastD , nSlowD , bounded = TRUE, smooth=smooth)
  price2 <- price
  price2 <- cbind(price2,sto)
  price2$higher <- price2$fastK > price2$fastD
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if (nrow(price2) > 0 ) {
    if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
    if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
    buy <- subset(price2,trade==1)[,c("date","close")]
    colnames(buy) <- c("buy_date","buy_price")
    sell <- subset(price2,trade==-1)[,c("date","close")]
    colnames(sell) <- c("sell_date","sell_price")
    if( nrow(buy) != nrow(sell) ) {
      MIN=min(nrow(buy),nrow(sell))
      buy <- buy[1:MIN,]
      sell <- sell[1:MIN,]
    }
    trades <- cbind(buy,sell)
    trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
    return(c(prod(trades$diff),nrow(trades)))
  } else {
    return(c(0,0))
  }
} else {
  return(c(0,0))
}
}

```

## Backtesting

Trying a range of values for nfastk,nfastd and nslowd

The slowD will be set to the same as fast D.

```{r,backtest1}

fee=0.001

nFastK <- seq(2,50,5)
nFastD <- seq(2,80,5)
smooth <- seq(1,80,5)
params <- expand.grid(nFastK,nFastD,smooth)
colnames(params) <- c("nFastK","nFastD","smooth")
res <- mclapply(X=1:nrow(params) , FUN=function(i) {
  nFastK=as.numeric(params[i,1])
  nFastD=as.numeric(params[i,2])
  nSlowD=as.numeric(params[i,2])
  smooth=as.numeric(params[i,3])
  mystoch(price,nFastK=nFastK,nFastD=nFastD,nSlowD=nSlowD,smooth=smooth,fee=fee)
} , mc.cores = CORES)
res <- as.data.frame(do.call(rbind, res))
colnames(res) <- c("returns","trades")
res2 <- cbind(params,res)
res2$returnspertrade <- res2$returns/res2$trades
res2$mean <- res2$returns^(1/res2$trades)

```

Strategies with best mean trade ROI.
These don't occur frequently, but when they do, it gives a solid return.

```{r,res1}

head(res2[order(-res2$mean),],20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Strategies with the highest returns - these are too frequent and each trade has only small returns.

```{r,res2}

head(res2[order(-res2$returns),],20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Simply exclude any strategies with more than 200 trades and then look at the ones with the highest
returns.

```{r,res3}

res3 <- subset(res2,trades<200)

head(res3[order(-res3$returns),],20) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Now let's have a look at a chart of the total returns and mean ROI of all the strategies tested.
An optimal strategy would be towards the upper right side of the chart.
The default strategy is shown in red.
In the second chart, I'm looking at the strategies with <200 trades since 2013.
Any strategy with mean ROI > 1.1 is considered good. 

```{r,plots1,fig.height=8,fig.width=8}

plot(res2$mean , res2$returns, log="y",xlab="mean ROI", ylab="total returns")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()

plot(res3$mean , res3$returns, log="y",xlab="mean ROI", ylab="total returns, <200 trades")
points(basic_mean,basic_returns,col="red",pch=19,cex=2)
abline(h=hodl,col="red")
grid()

```

Here's the table of results for the strategies with mean trade ROI of >108% and highest overall ROI.

```{r,res4}

res5 <- subset(res2, mean > 1.08 )
res5 <- res5[order(-res5$returns),]
res5 <- head(res5,20)
res5 %>% kbl() %>%  kable_paper("hover", full_width = F)

```

Now that I know which ranges of values work well, I will zoom in and do a more comprehensive
backtesting for the following values nfastk(32:52),nfastd/nslowd(42:82) and smooth (1:5).

```{r,backtest2}

fee=0.001

nFastK <- seq(min(res5$nFastK),max(res5$nFastK),1)
nFastD <- seq(min(res5$nFastD),max(res5$nFastD),1)
smooth <- seq(min(res5$smooth),max(res5$smooth),1)
params <- expand.grid(nFastK,nFastD,smooth)
colnames(params) <- c("nFastK","nFastD","smooth")
res <- mclapply(X=1:nrow(params) , FUN=function(i) {
  nFastK=as.numeric(params[i,1])
  nFastD=as.numeric(params[i,2])
  nSlowD=as.numeric(params[i,2])
  smooth=as.numeric(params[i,3])
  mystoch(price,nFastK=nFastK,nFastD=nFastD,nSlowD=nSlowD,smooth=smooth,fee=fee)
} , mc.cores = CORES)
res <- as.data.frame(do.call(rbind, res))
colnames(res) <- c("returns","trades")
res2 <- cbind(params,res)
res2$returnspertrade <- res2$returns/res2$trades
res2$mean <- res2$returns^(1/res2$trades)

```

Here are the highest returning strategies with fewer than 200 trades.

```{r,res6}

res3 <- subset(res2,trades<200)
res3 <- res3[order(-res3$returns),]

head(res3,10) %>% kbl() %>%  kable_paper("hover", full_width = F)

```

## Optimisation

This one looks to be the best: nFastK=7, nSlowD/nFastD=45, smooth=17

I have tested the best slowD and varying it did not make any improvement.

Below I show the results for this strategy, charts and each individual trade.

```{r,best1,fig.width=10,fig.height=8}

best_nFastK = res3[1,1]
best_nFastD = res3[1,2]
best_smooth = res3[1,3]

message(paste("Best nFastK:",best_nFastK))
message(paste("Best nFastD:",best_nFastD))
message(paste("Best smooth:",best_smooth))

sto <- stoch(HLC(price), nFastK=best_nFastK , nFastD=best_nFastD , nSlowD=best_nFastD , 
  bounded = TRUE, smooth=best_smooth)
price2 <- price
price2 <- cbind(price2,sto)
price2$higher <- price2$fastK > price2$fastD
yesterday <- price2$higher
yesterday <- c(NA,yesterday)
yesterday <- yesterday[1:(length(yesterday)-1)]
price2$yesterday <- yesterday
price2$trade <- price2$higher - price2$yesterday
price2 <- price2[which(price2$trade!=0),]
if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
buy <- subset(price2,trade==1)[,c("date","close")]
colnames(buy) <- c("buy_date","buy_price")
sell <- subset(price2,trade==-1)[,c("date","close")]
colnames(sell) <- c("sell_date","sell_price")
if( nrow(buy) != nrow(sell) ) {
  MIN=min(nrow(buy),nrow(sell))
  buy <- buy[1:MIN,]
  sell <- sell[1:MIN,]
}
trades <- cbind(buy,sell)
trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )

message(paste("Number of trades:",nrow(trades)))
message(paste("Total ROI:",prod(trades$diff)))
message(paste("Returns over hodl:",prod(trades$diff)/hodl))

stoch_dat <- c("indicator"="stoch",
  "parameter"= paste(best_nFastK, best_nFastD , best_smooth, sep=",") ,
  "meanROI"= mean(trades$diff) ,
  "totalROI"= prod(trades$diff) ,
  "ntrades"= nrow(trades) ,
  "ndays"= nrow(price) ,
  "xhodl"= prod(trades$diff)/hodl )

HEADER="STOCH"
par(mfrow=c(2,1))
plot(price$close~as.Date(price$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(sto[,1] ~ as.Date(price$date),type="l",col="blue",
  xlab="Date",ylab="stoch",main="Stoch")
lines(as.Date(price$date), sto[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

trades %>% kbl() %>%  kable_paper("hover", full_width = F)

```

## Recent action

```{r,recentaction,fig.height=8,fig.width=8}

HEADER="STOCH"
par(mfrow=c(2,1))
plot(tail(price$close,200)~tail(as.Date(price$date),200),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(tail(sto[,1],200) ~ tail(as.Date(price$date),200),type="l",col="blue",
  xlab="Date",ylab="stoch",main="Stoch")
lines(as.Date(price$date), sto[,2]  , col="red"  )
grid()
par(mfrow=c(1,1))

```

## Effect of fees

It is important to factor in the effect of fees because higher exchange fees will erode profits.

```{r,fees,fig.width=8,fig.height=7}

fee_levels <- c(0,0.0005,0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,0.01)

res_fees <- sapply(fee_levels,function(x) {
  mystoch(price,nFastK=best_nFastK,nFastD=best_nFastD,nSlowD=best_nFastD,smooth=best_smooth,fee=x)
} )

plot(fee_levels,res_fees[1,] ,log="y",pch=19,xlab="fee proportion",ylab="fold return",main="Effect of fees on returns")
mtext("higher time frame")
abline(h=hodl,col="red")
grid()

```

## Trades by year

It is good to see whether the strategy gives consistent returns over the years.

```{r,yearly}

years <- unique(sapply(strsplit(price[,1],"-"),"[[",1))

if (nrow(trades)>0) {
  trades$year <- sapply(strsplit(trades[,3],"-"),"[[",1)
  yearly <- sapply(years,function(y) { x <- subset(trades,year==y)[,5] } )
  yearly <- yearly[which(lapply(yearly,length)>0)]
  vioplot(yearly)
  yr <- sapply(yearly,prod)
  barplot(yr)
  text(x=(1:length(yr)*1.2)-0.4,y=yr*0.9,labels=signif(yr,2),cex=1.2)
  abline(h=1,col="red",lty=2)
}

```

## Data export

Saving data: indicator, parameter, meanROI, returns, ntrades, ndays, xhodl

```{r,export}

filename <- paste("TICKER","_dat.txt",sep="")
write.table(t(stoch_dat),file=filename,append=TRUE,row.names=FALSE,col.names=FALSE)

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
