---
title: "<br>Trading BTC, ETH and XMR based on an ensemble of MA indicators"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
---

## Intro

In this report I will be looking at whether a combination of MA derived indicators and RSI  works better than
each on their own. 
The theory is that there is a chance that combining them will reduce false positives.

The indicators I will be using are:

1. Optimal simple moving average (SMA).

2. Optimal SMA cross

3. Optimal exponential moving average (EMA)

4. Optimal EMA cross

5. RSI (21d timeframe)



```{r,lib}

suppressPackageStartupMessages({
  library("jsonlite")
  library("tidyverse")
  library("runner")
  library("quantmod")
  library("TTR")
  library("vioplot")
  library("parallel")
  library("kableExtra")
})

knitr::opts_chunk$set(fig.width=8, fig.height=8) 

```

## Get BTC data

Obtaining BTC historical data (daily).

```{r,getdata_btc}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=BTC&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="btcdat.txt")

dat <- fromJSON("btcdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,
  price$quote$USD$volume, stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close","volume")
price$date <- as.Date(sapply(strsplit(as.character(price$date),"T"),"[[",1))

price2 <- price

```

## Fetch BTC optimised parameters

These have been optimised with backtesting.

```{r,params_btc}

params <- read.table("https://mdz-analytics.com/altcoins/BTC/BTC_dat.txt", header=TRUE)

params %>% kbl(caption="optimised backtested parameters") %>%  kable_styling("hover", full_width = F)

```

## Calculate hodl returns

This is the benchmark we use for judging the performance of an indicator.
Any strategy with total returns under this value is considered bad.

```{r,hodl_btc}

myend <- price[c(1,nrow(price)),2][2]
mystart <- price[c(1,nrow(price)),2][1]
hodl <- (myend - mystart) / mystart
hodl

```

### SMA analysis

Fetching the optimal SMA value and comparing it to price action.

```{r,sma_btc}

best_sma = as.numeric(params[which(params$indicator == "SMA"),"parameter"])
ma <- SMA(Cl(price),n=best_sma)
price2$sma <- price2$close > ma

```

### SMA cross analysis

Fetching the optimal SMA cross values.

```{r,sma_cross_btc}

best_sma_cross = params[which(params$indicator == "SMAcross"),"parameter"]

best_sma_cross1 <- as.numeric(unlist(strsplit(best_sma_cross,","))[1])
best_sma_cross2 <- as.numeric(unlist(strsplit(best_sma_cross,","))[2])

sma1 <- SMA(Cl(price),n=best_sma_cross1)
sma2 <- SMA(Cl(price),n=best_sma_cross2)

price2$sma_cross <- sma2>sma1

```

### EMA analysis

Fetching the optimal EMA value and comparing it to price action.

```{r,ema_btc}

best_ema = as.numeric(params[which(params$indicator == "EMA"),"parameter"])
ema <- EMA(Cl(price),n=best_ema)
price2$ema <- price2$close > ema

```

### EMA cross analysis

Fetching the optimal EMA cross values.

```{r,ema_cross_btc}

best_ema_cross = params[which(params$indicator == "EMAcross"),"parameter"]

best_ema_cross1 <- as.numeric(unlist(strsplit(best_ema_cross,","))[1])
best_ema_cross2 <- as.numeric(unlist(strsplit(best_ema_cross,","))[2])

ema1 <- EMA(Cl(price),n=best_ema_cross1)
ema2 <- EMA(Cl(price),n=best_ema_cross2)

price2$ema_cross <- ema2>ema1

```

### RSI analysis

The idea here is to enter when  RSI recovers the 50 level.

Previously we showed that a period of 36 days gave good performance with an entry and exit at the 40 level.

```{r,rsi_btc}

n_rsi <- 36

rsi <- RSI(Cl(price),n=n_rsi)

price2$rsi <- rsi>50

```

## Ensemble trading

Now that the indicators have each been calculated, we can open a trade when X indicators are positive and sell when only Y are positive.

In order to decide on X and Y, we should look at the data first.

```{r,ensembl1_btc}

price2[is.na(price2)] <- 0
price2$sum <- price2$sma + price2$sma_cross + price2$ema + price2$ema_cross + price2$rsi

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$sum ~ as.Date(price2$date),type="l",
  xlab="Date",ylab="positive indicators",main="Indicator set")
grid()

par(mfrow=c(1,1))

plot(price2$close ~ as.Date(price2$date),type="p",log="y", pch=19,cex=0.5,
  col=price2$sum+1,
  xlab="Date",ylab="price (USD)",main="USD price coloured by num indicators")
  mtext("black=0,red=1,green=2,blue=3,red=4,purple=5")
grid()

```

Now that we've calculated the number of indicators that are positive and negative at each timepoint,
we can assess what good entry and exit points are.

```{r,backtest1_btc}

ensembl <- function(price2,threshold,fee) {
  price2$higher <- price2$sum >= threshold
  yesterday <- price2$higher
  yesterday <- c(NA,yesterday)
  yesterday <- yesterday[1:(length(yesterday)-1)]
  price2$yesterday <- yesterday
  price2$trade <- price2$higher - price2$yesterday
  price2 <- price2[which(price2$trade!=0),]
  if ( price2$trade[1] == -1 ) { price2 <- price2[2:nrow(price2),] }
  if ( tail(price2,1)$trade ) { price2 <- price2[1:(nrow(price2)-1),] }
  buy <- subset(price2,trade==1)[,c("date","close")]
  colnames(buy) <- c("buy_date","buy_price")
  sell <- subset(price2,trade==-1)[,c("date","close")]
  colnames(sell) <- c("sell_date","sell_price")
  if( nrow(buy) != nrow(sell) ) {
    MIN=min(nrow(buy),nrow(sell))
    buy <- buy[1:MIN,]
    sell <- sell[1:MIN,]
  }
  trades <- cbind(buy,sell)
  trades$diff <- ( ( trades$sell_price - trades$buy_price ) / trades$buy_price ) +1 - ( 2 * fee )
  return(c(prod(trades$diff),nrow(trades)))
}

myrange=1:5
res <- t( sapply(myrange,function(i) {
  ensembl(price2,threshold=i,fee=0.005)
}) )

rownames(res) <- myrange
xhodl <- t(t(as.numeric(res[, 1]/hodl)))
res <- cbind(res,xhodl)
colnames(res) <- c("totalROI","n_trades","xhodl")
res <- as.data.frame(res)
res$pos_indicator <- myrange
res <- res[,c(4,2,1,3)]
res %>% kbl(caption="backtesting results with minimum postive indicators") %>%  kable_styling("hover", full_width = F)

```

## Get ETH data

Obtaining ETH historical data (daily).

```{r,getdata_eth}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=ETH&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="ethdat.txt")

dat <- fromJSON("ethdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,
  price$quote$USD$volume, stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close","volume")
price$date <- as.Date(sapply(strsplit(as.character(price$date),"T"),"[[",1))

price2 <- price

```

## Fetch ETH optimised parameters

These have been optimised with backtesting.

```{r,params_eth}

params <- read.table("https://mdz-analytics.com/altcoins/ETH/ETH_dat.txt", header=TRUE)

params %>% kbl(caption="optimised backtested parameters") %>%  kable_styling("hover", full_width = F)

```

## Calculate hodl returns

This is the benchmark we use for judging the performance of an indicator. 
Any strategy with total returns under this value is considered bad.

```{r,hodl_eth}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

### SMA analysis

Fetching the optimal SMA value and comparing it to price action.

```{r,sma_eth}

best_sma = as.numeric(params[which(params$indicator == "SMA"),"parameter"])
ma <- SMA(Cl(price),n=best_sma)
price2$sma <- price2$close > ma

```

### SMA cross analysis

Fetching the optimal SMA cross values.

```{r,sma_cross_eth}

best_sma_cross = params[which(params$indicator == "SMAcross"),"parameter"]

best_sma_cross1 <- as.numeric(unlist(strsplit(best_sma_cross,","))[1])
best_sma_cross2 <- as.numeric(unlist(strsplit(best_sma_cross,","))[2])

sma1 <- SMA(Cl(price),n=best_sma_cross1)
sma2 <- SMA(Cl(price),n=best_sma_cross2)

price2$sma_cross <- sma2>sma1

```

### EMA analysis

Fetching the optimal EMA value and comparing it to price action.

```{r,ema_eth}

best_ema = as.numeric(params[which(params$indicator == "EMA"),"parameter"])
ema <- EMA(Cl(price),n=best_ema)
price2$ema <- price2$close > ema

```

### EMA cross analysis

Fetching the optimal EMA cross values.

```{r,ema_cross_eth}

best_ema_cross = params[which(params$indicator == "EMAcross"),"parameter"]

best_ema_cross1 <- as.numeric(unlist(strsplit(best_ema_cross,","))[1])
best_ema_cross2 <- as.numeric(unlist(strsplit(best_ema_cross,","))[2])

ema1 <- EMA(Cl(price),n=best_ema_cross1)
ema2 <- EMA(Cl(price),n=best_ema_cross2)

price2$ema_cross <- ema2>ema1

```

### RSI analysis

The idea here is to enter when  RSI recovers the 50 level.

Previously we showed that a period of 21 days gave good performance with an entry at 49 and exit at 54.

```{r,rsi_eth}

n_rsi <- 21

rsi <- RSI(Cl(price),n=n_rsi)

price2$rsi <- rsi>50

```

## Ensemble trading

Now that the indicators have each been calculated, we can open a trade when X indicators are positive and sell when only Y are positive.

In order to decide on X and Y, we should look at the data first.

```{r,ensembl1_eth}

price2[is.na(price2)] <- 0
price2$sum <- price2$sma + price2$sma_cross + price2$ema + price2$ema_cross + price2$rsi

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$sum ~ as.Date(price2$date),type="l",
  xlab="Date",ylab="positive indicators",main="Indicator set")
grid()

par(mfrow=c(1,1))

plot(price2$close ~ as.Date(price2$date),type="p",log="y", pch=19,cex=0.5,
  col=price2$sum+1,
  xlab="Date",ylab="price (USD)",main="USD price coloured by num indicators")
  mtext("black=0,red=1,green=2,blue=3,red=4,purple=5")
grid()

```

Now that we've calculated the number of indicators that are positive and negative at each timepoint,
we can assess what good entry and exit points are.

```{r,backtest1_eth}

myrange=1:5
res <- t( sapply(myrange,function(i) {
  ensembl(price2,threshold=i,fee=0.005)
}) )

rownames(res) <- myrange
xhodl <- t(t(as.numeric(res[, 1]/hodl)))
res <- cbind(res,xhodl)
colnames(res) <- c("totalROI","n_trades","xhodl")
res <- as.data.frame(res)
res$pos_indicator <- myrange
res <- res[,c(4,2,1,3)]
res %>% kbl(caption="backtesting results with minimum postive indicators") %>%  kable_styling("hover", full_width = F)

```

## Get XMR data

Obtaining XMR historical data (daily).

```{r,getdata_xmr}

mydate <- Sys.Date()-1

URL=paste("https://web-api.coinmarketcap.com/v1/cryptocurrency/ohlcv/historical?symbol=XMR&convert=USD&interval=daily&time_start=2013-06-01&time_end=",mydate,sep="")

download.file(URL,destfile="xmrdat.txt")

dat <- fromJSON("xmrdat.txt")
price <- dat$data$quotes
price <- data.frame(price$time_close, price$quote$USD$high, price$quote$USD$low, price$quote$USD$close,
  price$quote$USD$volume, stringsAsFactors=FALSE)
colnames(price) <- c("date","high","low","close","volume")
price$date <- as.Date(sapply(strsplit(as.character(price$date),"T"),"[[",1))

price2 <- price

```

## Fetch XMR optimised parameters

These have been optimised with backtesting.

```{r,params_xmr}

params <- read.table("https://mdz-analytics.com/altcoins/XMR/XMR_dat.txt", header=TRUE)

params %>% kbl(caption="optimised backtested parameters") %>%  kable_styling("hover", full_width = F)

```

## Calculate hodl returns

This is the benchmark we use for judging the performance of an indicator. 
Any strategy with total returns under this value is considered bad.

```{r,hodl_xmr}

myend <- price[c(1,nrow(price)),2][2] 
mystart <- price[c(1,nrow(price)),2][1] 
hodl <- (myend - mystart) / mystart
hodl

```

### SMA analysis

Fetching the optimal SMA value and comparing it to price action.

```{r,sma_xmr}

best_sma = as.numeric(params[which(params$indicator == "SMA"),"parameter"])
ma <- SMA(Cl(price),n=best_sma)
price2$sma <- price2$close > ma

```

### SMA cross analysis

Fetching the optimal SMA cross values.

```{r,sma_cross_xmr}

best_sma_cross = params[which(params$indicator == "SMAcross"),"parameter"]

best_sma_cross1 <- as.numeric(unlist(strsplit(best_sma_cross,","))[1])
best_sma_cross2 <- as.numeric(unlist(strsplit(best_sma_cross,","))[2])

sma1 <- SMA(Cl(price),n=best_sma_cross1)
sma2 <- SMA(Cl(price),n=best_sma_cross2)

price2$sma_cross <- sma2>sma1

```

### EMA analysis

Fetching the optimal EMA value and comparing it to price action.

```{r,ema_xmr}

best_ema = as.numeric(params[which(params$indicator == "EMA"),"parameter"])
ema <- EMA(Cl(price),n=best_ema)
price2$ema <- price2$close > ema

```

### EMA cross analysis

Fetching the optimal EMA cross values.

```{r,ema_cross_xmr}

best_ema_cross = params[which(params$indicator == "EMAcross"),"parameter"]

best_ema_cross1 <- as.numeric(unlist(strsplit(best_ema_cross,","))[1])
best_ema_cross2 <- as.numeric(unlist(strsplit(best_ema_cross,","))[2])

ema1 <- EMA(Cl(price),n=best_ema_cross1)
ema2 <- EMA(Cl(price),n=best_ema_cross2)

price2$ema_cross <- ema2>ema1

```

### RSI analysis

The idea here is to enter when  RSI recovers the 50 level.

Previously we showed that a period of 21 days gave good performance with an entry at 49 and exit at 54.

```{r,rsi_xmr}

n_rsi <- 21

rsi <- RSI(Cl(price),n=n_rsi)

price2$rsi <- rsi>50

```

## Ensemble trading

Now that the indicators have each been calculated, we can open a trade when X indicators are positive and sell when only Y are positive.

In order to decide on X and Y, we should look at the data first.

```{r,ensembl1_xmr}

price2[is.na(price2)] <- 0
price2$sum <- price2$sma + price2$sma_cross + price2$ema + price2$ema_cross + price2$rsi

par(mfrow=c(2,1))
plot(price2$close ~ as.Date(price2$date),type="l",log="y",
  xlab="Date",ylab="price (USD)",main="USD price")
grid()

plot(price2$sum ~ as.Date(price2$date),type="l",
  xlab="Date",ylab="positive indicators",main="Indicator set")
grid()

par(mfrow=c(1,1))

plot(price2$close ~ as.Date(price2$date),type="p",log="y", pch=19,cex=0.5,
  col=price2$sum+1,
  xlab="Date",ylab="price (USD)",main="USD price coloured by num indicators")
  mtext("black=0,red=1,green=2,blue=3,red=4,purple=5")
grid()

```

Now that we've calculated the number of indicators that are positive and negative at each timepoint,
we can assess what good entry and exit points are.

```{r,backtest1_xmr}

myrange=1:5
res <- t( sapply(myrange,function(i) {
  ensembl(price2,threshold=i,fee=0.005)
}) )

rownames(res) <- myrange
xhodl <- t(t(as.numeric(res[, 1]/hodl)))
res <- cbind(res,xhodl)
colnames(res) <- c("totalROI","n_trades","xhodl")
res <- as.data.frame(res)
res$pos_indicator <- myrange
res <- res[,c(4,2,1,3)]
res %>% kbl(caption="backtesting results with minimum postive indicators") %>%  kable_styling("hover", full_width = F)

```



## Conclusion

This analysis indicates that using a basket of simple indicators tends to work better than the
individual indicators separately.

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,sessioninfo}

sessionInfo()

```

</details>
